---
title: |
  Very little data tampering is needed to manufacture statistically significant results:\
  A Monte Carlo simulation study
author:
  - name: Ian Hussey
    email: ian.hussey@unibe.ch
    affiliations: University of Bern
    attributes:
        corresponding: true
abstract: 'Fabricating a statistically significant result does not require fabricating a dataset. I simulate a dishonest analyst who starts from data generated under the true null hypothesis and alters, deletes, or invents one data point at a time, always the point that most helps, re-running the test after every change until *p* < .05. I formalize this as a family of iterative algorithmically-greedy tampering strategies (dropping cases, switching condition labels, re-pairing values, duplicating cases, fabricating cases, and replacing cases) across three common designs: the two-group mean difference (Student''s t-test), the Pearson correlation, and the 2x2 association (chi-square test). Each strategy runs until significance is reached or further tampering becomes structurally impossible, and the estimand is the effort required: the number of tampered data points, both absolutely and as a proportion of the sample. Two findings recur. First, effort is small in absolute terms and falls steeply in relative terms as samples grow. Median tamper counts rise only slowly with sample size, from a handful of data points at N = 20 to a few dozen at N = 400, so expressed as a share of the sample the effort drops by roughly an order of magnitude. Large studies are therefore proportionally easier to corrupt, not safer. Second, the strategy that is hardest to detect in the two-group design, because it leaves every univariate distribution untouched (condition switching), is also among the cheapest; its correlation analogue (value re-pairing) is cheaper than duplication and costs about as much as unprincipled case deletion, while being invisible to any check of either variable alone. I conclude that fabrication and falsification of results typically requires a worryingly small amount of actual data tampering.'
format:
  elsevier-pdf:
    keep-tex: true
    latex-tinytex: false
    # The elsevier extension defaults to cite-method: natbib, which formats the
    # reference list with elsarticle-harv.bst and ignores csl. Switch to
    # citeproc so the APA 7th CSL below governs both in-text and list style.
    cite-method: citeproc
    # elsarticle.cls loads natbib unconditionally unless given nonatbib, and
    # natbib then errors on the CSLReferences list citeproc emits ("Bibliography
    # not compatible with author-year citations"). nonatbib is the class's own
    # opt-out and is the supported way to pair elsarticle with citeproc.
    classoption:
      - nonatbib
    fig-format: pdf # vector figures embedded in the PDF (no rasterization)
    journal:
      name: psyarxiv
      formatting: preprint
      cite-style: authoryear
execute:
  echo: false
header-includes:
      - \usepackage{threeparttable}
      - \usepackage{pdflscape} # for the landscape page holding Table 2
      - \usepackage{unicode-math}
      - \usepackage[margin=4cm]{geometry} # tighter page margins; gives \textwidth = 386.67pt = 5.35in,
      # which is the fig-width used by every figure chunk so that figures are placed at native size
      - \usepackage{etoolbox}
      - \AtBeginEnvironment{longtable}{\footnotesize} # smaller font inside tables
      - \AtBeginEnvironment{tabular}{\small}
      # double the paragraph indent. AtBeginDocument because the elsevier
      # template sets \parindent after header-includes; 2\parindent doubles
      # whatever it set rather than hardcoding a value.
      - \AtBeginDocument{\setlength{\parindent}{2\parindent}}
bibliography: bibliography.bib
csl: apa.csl # APA Style 7th edition
# Make each in-text citation a hyperlink to its reference-list entry. citeproc
# always emits the \bibitem anchors, but without this nothing points at them.
link-citations: true
# In-text citations use @key syntax, so the reference list is cited-only: an
# entry appears if and only if something cites it, and an unmatched key fails
# loudly at render. Do not reinstate `nocite: @*`, which would silently print
# uncited entries and disable that check.
knitr:
  opts_chunk:
    dev: cairo_pdf # vector device: razor-sharp lines/text at any zoom
    message: false
    warning: false
---

```{r}
#| label: setup
#| include: false

# Every number, figure, and table in this manuscript is computed at render time from
# the simulation caches in ../../data (written by the three research compendia in
# ../../code, which are the canonical source for the simulation and summarising code;
# the helpers below are trimmed copies). Re-rendering after a new simulation run
# updates the entire results section with no manual editing.

library(dplyr)
library(tidyr)
library(tibble)
library(purrr)
library(readr)
library(stringr)
library(forcats)
library(ggplot2)
library(scales)
library(patchwork)
library(knitr)
library(kableExtra)

options(scipen = 999)

# NOTE for editors: knitr chunk options (fig-cap, tbl-cap) are processed by R, and this
# document is rendered in a C locale, which rewrites any non-ASCII character in them as the
# literal text "<U+XXXX>". Keep chunk options ASCII-only -- write "2x2", "+/-", "-" rather
# than the typographic characters. Body prose is handled by pandoc, not R, and may use them
# freely.

data_dir <- "../../data"

# ---- catalogue of simulation results ------------------------------------------------
# One row per (setting, strategy, prevalence). `stem` names the cache written by the
# corresponding optimised compendium in ../../code. Those are the definitive results; the
# reference implementations return identical output but run far more slowly, which
# code/verify_optimised_implementations.R checks.
sources <- tribble(
  ~setting,           ~method,                               ~stem,                      ~prevalence,
  "Mean difference",  "Dropping",                            "dropping_meandiff",         NA_real_,
  "Mean difference",  "Condition switching",                 "swaps_meandiff",            NA_real_,
  "Mean difference",  "Duplication",                         "duplication_meandiff",      NA_real_,
  "Mean difference",  "Fabrication",                         "fabrication_meandiff",      NA_real_,
  "Mean difference",  "Replacement",                         "replacement_meandiff",      NA_real_,
  "Correlation",      "Dropping",                            "dropping_correlation",      NA_real_,
  "Correlation",      "Re-pairing",                          "repairing_correlation",     NA_real_,
  "Correlation",      "Duplication",                         "duplication_correlation",   NA_real_,
  "Correlation",      "Fabrication",                         "fabrication_correlation",   NA_real_,
  "Correlation",      "Replacement",                         "replacement_correlation",   NA_real_,
  "2x2 association",  "Duplication/fabrication/replacement", "fabrication_2x2",           0.20,
  "2x2 association",  "Dropping",                            "dropping_2x2",              0.20,
  "2x2 association",  "Condition switching",                 "switching_2x2",             0.20,
  "2x2 association",  "Duplication/fabrication/replacement", "fabrication_2x2_lowprev",   0.05,
  "2x2 association",  "Dropping",                            "dropping_2x2_lowprev",      0.05,
  "2x2 association",  "Condition switching",                 "switching_2x2_lowprev",     0.05
) |>
  mutate(file = file.path(data_dir, paste0("simulation_iterative_", stem, "_opt_full_k10000.rds")))

missing <- sources$file[!file.exists(sources$file)]
if (length(missing) > 0) {
  stop("Missing simulation caches (run the optimised compendia in ../../code first):\n  ",
       paste(basename(missing), collapse = "\n  "))
}

# ---- assemble one long per-iteration data frame across all settings -----------------
# Each compendium names its columns for its own design, so each is mapped onto the
# common (n_total, tampers_needed, reached_sig, es) schema here.
read_one <- function(setting, method, file, prevalence, ...) {
  d <- read_rds(file)
  n_total <- if ("n" %in% names(d)) d$n else d$n_per_condition * 2
  # the compendia also record the effect size of each tampered dataset; it is not read
  # here, as this article reports only the effort required to manufacture significance
  tibble(setting = setting, method = method, prevalence = prevalence, n_total = n_total,
         tampers_needed = d$tampers_needed, reached_sig = d$reached_sig)
}

raw <- sources |>
  pmap_dfr(read_one) |>
  mutate(setting = factor(setting, levels = c("Mean difference", "Correlation", "2x2 association")))

# ---- Monte Carlo standard errors (trimmed copies of the compendia helpers) ---------

mcse_mean <- function(x){
  x <- x[!is.na(x)]
  if (length(x) < 2) return(NA_real_)
  sd(x) / sqrt(length(x))
}

mcse_proportion <- function(p, k){
  if (is.na(p) || k < 1) return(NA_real_)
  sqrt(p * (1 - p) / k)
}

mcse_median <- function(x, n_boot = 1000, seed = 42){
  x <- x[!is.na(x)]
  n <- length(x)
  if (n < 2) return(NA_real_)
  old_seed <- if (exists(".Random.seed", envir = .GlobalEnv)) get(".Random.seed", envir = .GlobalEnv) else NULL
  set.seed(seed)
  boot_medians <- replicate(n_boot, median(x[sample.int(n, n, replace = TRUE)]))
  if (!is.null(old_seed)) assign(".Random.seed", old_seed, envir = .GlobalEnv)
  if (!all(is.finite(boot_medians))) return(NA_real_) # a censored median has no estimable MCSE
  sd(boot_medians)
}

q_censored <- function(x, probs){
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_real_)
  unname(quantile(x, probs = probs, names = FALSE, type = 7))
}

safe_mean <- function(x){ x <- x[is.finite(x)]; if (length(x) == 0) NA_real_ else mean(x) }
safe_sd   <- function(x){ x <- x[is.finite(x)]; if (length(x) < 2)  NA_real_ else sd(x) }

# ---- censoring-aware summaries ------------------------------------------------------
# An iteration that never attained significance (tampering ran to a structural limit
# without crossing alpha) has no finite tamper count; coding it Inf ranks it above every
# finite count, so percentiles among the finite counts are exact and percentiles among
# the censored cases return Inf (reported as "never"). See the compendia for rationale.

tampers <- raw |>
  mutate(tampers_censored = ifelse(reached_sig, tampers_needed, Inf)) |>
  group_by(setting, method, prevalence, n_total) |>
  summarize(k = n(),
            never = mean(!reached_sig),
            never_mcse = mcse_proportion(never, k),
            prop_gt_10 = mean(tampers_censored > 0.10 * n_total),
            prop_gt_10_mcse = mcse_proportion(prop_gt_10, k),
            prop_gt_25 = mean(tampers_censored > 0.25 * n_total),
            prop_gt_25_mcse = mcse_proportion(prop_gt_25, k),
            median = q_censored(tampers_censored, 0.5),
            median_mcse = mcse_median(tampers_censored),
            iqr_lower = q_censored(tampers_censored, 0.25),
            iqr_upper = q_censored(tampers_censored, 0.75),
            pi_lower = q_censored(tampers_censored, 0.025),
            pi_upper = q_censored(tampers_censored, 0.975),
            mean_reached = safe_mean(tampers_needed[reached_sig]),
            mean_reached_mcse = mcse_mean(tampers_needed[reached_sig]),
            .groups = "drop") |>
  mutate(prop_median = median / n_total,
         prop_median_mcse = median_mcse / n_total,
         prop_iqr_lower = iqr_lower / n_total,
         prop_iqr_upper = iqr_upper / n_total,
         prop_pi_lower = pi_lower / n_total,
         prop_pi_upper = pi_upper / n_total)

# ---- strategy display order: computed from the data, not hardcoded ------------------
# Ordered by impact within each setting: ascending mean (over sample sizes) of the
# censoring-aware median proportion of the sample tampered. Strategies whose median is
# censored at some N sort by the finite cells they have.
method_order <- tampers |>
  group_by(setting, method) |>
  summarize(o = mean(prop_median[is.finite(prop_median)]), .groups = "drop") |>
  arrange(setting, o)

order_methods <- function(df, setting_name) {
  lv <- method_order |> filter(setting == setting_name) |> pull(method)
  df |> mutate(method = factor(method, levels = lv))
}

# ---- shared plotting pieces ---------------------------------------------------------

x_breaks <- c(10, seq(25, 200, by = 25)) * 2

# Strategy is a nominal variable, so the palette is qualitative rather than sequential:
# five muted but clearly separated hues (sienna, olive, ochre, slate, plum) so that
# adjacent categories are told apart by hue and not by lightness, as a viridis ramp
# would require. Ordered so the first three are maximally distinct, since the 2x2
# setting uses only three.
earth_palette <- c("#B4573B", "#7A8C4E", "#3F6B7A", "#D9A441", "#5C4A5E")

# Plot text is set well below the document's 10pt body text so that, at the fixed figure
# width of one text block, more of the figure is panel and less is label.
theme_effort <- function() {
  theme_minimal(base_size = 8) +
    theme(panel.grid.minor = element_blank(),
          legend.key.size = unit(0.8, "lines"),
          plot.margin = margin(2, 2, 2, 2))
}

plot_effort <- function(df, y_count_lab, dodge = 27) {
  # Bars are interquartile ranges, not 95% prediction intervals. One cell (correlation
  # duplication at N = 20, where the duplication pool is only two points) has a 97.5th
  # percentile at nearly nine times the sample size, which would set the y scale for every
  # panel and bury the pattern the figure exists to show. The prediction intervals are in
  # @tbl-results.
  #
  # Where more than a quarter of datasets never attain significance, the upper quartile has
  # no finite value. Simply dropping the bar there reads as a plotting failure rather than
  # as a result, so those bars are instead drawn to the top of the data range and capped
  # with a hollow triangle, the usual convention for an interval that continues beyond the
  # axis. The lower quartile and the median are still drawn wherever they are finite.
  cap_count <- max(df$iqr_upper[is.finite(df$iqr_upper)], na.rm = TRUE) * 1.04
  cap_prop  <- max(df$prop_iqr_upper[is.finite(df$prop_iqr_upper)], na.rm = TRUE) * 1.04
  df <- df |> mutate(censored = !is.finite(iqr_upper),
                     iqr_upper_f = ifelse(censored, cap_count, iqr_upper),
                     prop_iqr_upper_f = ifelse(censored, cap_prop, prop_iqr_upper),
                     median_f = ifelse(is.finite(median), median, NA_real_),
                     prop_median_f = ifelse(is.finite(prop_median), prop_median, NA_real_))
  cens <- df |> filter(censored)
  p1 <- ggplot(df, aes(n_total, median_f, color = method)) +
    geom_point(position = position_dodge(width = dodge), size = 1) +
    geom_linerange(aes(ymin = iqr_lower, ymax = iqr_upper_f), position = position_dodge(width = dodge)) +
    geom_point(data = cens, aes(y = iqr_upper_f), shape = 24, size = 1.1, fill = "white",
               position = position_dodge(width = dodge)) +
    labs(x = "Total sample size (N)", y = y_count_lab, color = "Strategy") +
    scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty(n = 6)) +
    scale_x_continuous(breaks = x_breaks) +
    scale_color_manual(values = earth_palette) +
    theme_effort()
  p2 <- ggplot(df, aes(n_total, prop_median_f, color = method)) +
    geom_point(position = position_dodge(width = dodge), size = 1) +
    geom_linerange(aes(ymin = prop_iqr_lower, ymax = prop_iqr_upper_f), position = position_dodge(width = dodge)) +
    geom_point(data = cens, aes(y = prop_iqr_upper_f), shape = 24, size = 1.1, fill = "white",
               position = position_dodge(width = dodge)) +
    labs(x = "Total sample size (N)", y = "Tampering operations per original observation",
         color = "Strategy") +
    scale_y_continuous(labels = percent_format(), limits = c(0, NA), breaks = breaks_pretty(n = 6)) +
    scale_x_continuous(breaks = x_breaks) +
    scale_color_manual(values = earth_palette) +
    theme_effort()
  p1 + p2 + plot_layout(ncol = 1, guides = "collect")
}

# ---- table cell formatting (MCSE-driven decimal places; Morris et al., 2019) --------

dp_from_mcse <- function(mcse, min_dp = 0, max_dp = 4){
  mcse <- mcse[is.finite(mcse) & mcse > 0]
  if (length(mcse) == 0) return(min_dp)
  min(max_dp, max(min_dp, max(ceiling(-log10(mcse)) + 1)))
}
fmt_num <- function(x, dp){
  ifelse(is.na(x), "NA", ifelse(is.finite(x), formatC(x, format = "f", digits = dp), "never"))
}
mcse_cell <- function(estimate, mcse, min_dp = 0, max_dp = 4){
  dp <- dp_from_mcse(mcse, min_dp = min_dp, max_dp = max_dp)
  ifelse(is.finite(estimate) & is.finite(mcse),
         paste0(fmt_num(estimate, dp), " (", fmt_num(mcse, dp), ")"),
         fmt_num(estimate, dp))
}
interval_cell <- function(lower, upper, dp = 0){
  paste0("[", fmt_num(lower, dp), ", ", fmt_num(upper, dp), "]")
}
fmt_pct <- function(x, dp = 1) paste0(formatC(100 * x, format = "f", digits = dp), "\\%")
fmt_val <- function(x, dp = 0) formatC(x, format = "f", digits = dp)

# ---- key quantities used inline in the prose ---------------------------------------

grab <- function(s, m, n, col, prev = NA) {
  d <- tampers |> filter(setting == s, method == m, n_total == n)
  if (!is.na(prev)) d <- d |> filter(prevalence == prev)
  d[[col]][1]
}

continuous_400 <- tampers |>
  filter(setting != "2x2 association", n_total == 400, is.finite(prop_median))
continuous_20 <- tampers |>
  filter(setting != "2x2 association", n_total == 20, is.finite(prop_median))
# how effort scales: the count grows sub-linearly in N, so the proportion falls
scaling <- tampers |>
  filter(setting != "2x2 association", n_total %in% c(20, 400), is.finite(median)) |>
  select(setting, method, n_total, median, prop_median) |>
  pivot_wider(names_from = n_total, values_from = c(median, prop_median)) |>
  mutate(fold_prop = prop_median_20 / prop_median_400)
gt10_400 <- tampers |> filter(setting != "2x2 association", n_total == 400) |> pull(prop_gt_10)
never_400 <- tampers |> filter(setting != "2x2 association", n_total == 400) |> pull(never)

key <- list(
  # range of the median proportion tampered at the largest N, across the two continuous settings
  prop400_min = min(continuous_400$prop_median),
  prop400_max = max(continuous_400$prop_median),
  prop20_min  = min(continuous_20$prop_median),
  prop20_max  = max(continuous_20$prop_median),
  count20_min = min(continuous_20$median),   count20_max = max(continuous_20$median),
  count400_min = min(continuous_400$median), count400_max = max(continuous_400$median),
  fold_prop_min = min(scaling$fold_prop),    fold_prop_max = max(scaling$fold_prop),
  # cheapest distribution-preserving strategies at N = 100
  swap_median_100   = grab("Mean difference", "Condition switching", 100, "median"),
  repair_median_100 = grab("Correlation", "Re-pairing", 100, "median"),
  # proportion of datasets needing more than 10% of the sample at N = 400, across continuous strategies
  gt10_400_max = max(gt10_400),
  gt10_400_min = min(gt10_400),
  never_400_max = max(never_400),
  # structural impossibility at the rare-outcome extreme
  lowprev_drop_never_20 = grab("2x2 association", "Dropping", 20, "never", prev = 0.05),
  k_iterations = max(tampers$k)
)
```

# Introduction

In November 2019, the pivotal trial of a treatment for ANCA-associated vasculitis was unblinded and its primary analysis run. The result was not significant (*p* = .1025). According to a subsequent letter from the US Food and Drug Administration, unblinded personnel employed by the sponsor then selected nine of the trial's 331 patients for re-adjudication of their remission status, having first confirmed that reclassifying five of them would move the study across the significance threshold. The five were reclassified, the database was re-locked, and the re-run analysis returned *p* = .0132. That analysis was the only one submitted in support of approval, which was granted in 2021. In April 2026 the FDA proposed to withdraw it, concluding that the integrity of the trial had been compromised; the sponsor contests this and a hearing is pending [@us_food_and_drug_administration_center_for_drug_evaluation_and_research_proposal_2026]. Whatever its resolution, one feature of the case is worth dwelling on. Five records out of 331, just 1.5% of the sample, separated a failed trial from a successful one.

How often does it occur that data are simply tampered with until statistical significance is reached? Meta-analysis of the survey literature puts self-admitted research misconduct at 2.9% of researchers (95% CI [2.1, 3.8]), comprising 1.9% for fabrication and 3.3% for falsification, while 15.5% report having witnessed it in others [@xie_prevalence_2021; see also @fanelli_how_2009]. When the National Survey on Research Integrity used a randomized response design to give respondents provable deniability, admitted fabrication rose to 4.3% and falsification to 4.2% [@gopalakrishna_prevalence_2022]. From the other side of the desk, 24% of consulting biostatisticians in a US national survey had been asked, at least once in the preceding five years, to remove or alter data records to better support a research hypothesis [@wang_researcher_2018]. These are not the rates of a vanishing phenomenon, and each is likely a lower bound of the true prevalence. 

Prevalence, though, is only half the question. The other half is how 'effective' such data manipulations actually are. Here, the literature is largely silent. @chambers_seven_2017 set out a "dirty dozen" of ways to guarantee significant results even under the null hypothesis, the second of which is switching participants between conditions. @brown_condition_2020 posted a short simulation of the efficacy of this in a blog post. Both convey the same intuition that a handful of unjustified changes to the data is enough to produce statistically significant results in many cases, but neither is a systematic assessment. The field has no systematic account of how much tampering a given strategy requires, whether that requirement grows with sample size, how the strategies compare with one another, or how large an effect they leave behind.

For the ambiguous end of the continuum of merely 'questionable' research practices, that gap has been filled. @stefan_big_2023 compiled 12 *p*-hacking strategies from the literature, identified the factors controlling each one's severity, and simulated the resulting increase in false-positive rates. Their scope, as they state it, is the "grey area between good practice and outright scientific misconduct". They stop deliberately at its far edge: discussing misreported test statistics, they note that at some point one must ask whether a practice is still questionable or has become outright fraud. As such, the literature says comparably little about the impact of indefensible research practices on false-positive rates. Of course, given a willingness to engage in an arbitrary amount of data falsification or fabrication, by definition one can obtain statistically significant results. Here, I ask a narrower question: how much data tampering, of what form, is typically needed to acquire significant results.  

Quantifying the indefensible practices is worth doing because of how researchers appear to decide what to do. @entradas_shades_2026 find that perceived seriousness is by far the strongest predictor of whether a researcher engages in a questionable practice, ahead of every career and demographic factor they measure. Equally, in @john_measuring_2012, deciding whether to exclude data after seeing its effect on the results was admitted by 38–43% of psychologists and rated 1.61 out of 2 for defensibility, while falsifying data was admitted by 0.6–1.7% and rated 0.16. These results are consistent with the idea of perceived norms shaping research behaviors. That is, if researchers are more likely to engage in practices they understand to be seriously problematic and are less likely to engage in ones that are seen as such. However, the boundary between these plausibly varies between individuals, research communities, and over time, creating what is effectively an Overton Window of questionable versus unacceptable research practices. Many research practices that were regarded as acceptable prior to the invention of the term *p*-hacking are now considered in many research communities to be substantially less acceptable. If perceptions of how problematic a given set of research behaviors are influence the prevalence of those behaviours, then establishing how relatively bad a given set of practices actually are, how little effort they take, and how much bias they inject, likely represents important information to dissuading their practice. 

This perspective is an explicitly social-epistemological one, which understands researcher behaviors as a set of evolving community norms rather than moral acts. Equally, this perspective is concerned only with the undesirable statistical consequences of certain researcher behaviors, but is agnostic to researchers' intent when taking them. With that being said, at a human level, it is easier to imagine a researcher who deletes a small number of data points to 'reveal' an effect they strongly believe is there, lying just under the surface of the data, when it is not (e.g., cases of 'qualitative exclusion strategies': @schiller_addendum_2018) than to imagine scientists simply making up datasets wholesale [although the latter does occur; @verfaellie_case_2011], despite both strategies producing false-positive results. Regardless of whether these two scenarios can be equated in their intent, it is an open question whether they can be equated in their outcomes, as it is generally unclear how relatively robust or fragile most results are to modifications of only a small number of data points [although see @spiess_scientific_2024]. 

Some definition matters are worth considering, insofar as they informed which tampering strategies were simulated. The US government's Office of Research Integrity provides influential definitions of research misconduct and its forms [@steneck_introduction_2007]. According to the ORI definitions, fabrication is "making up data or results and recording or reporting them". This is represented by the the fabrication, duplication and replacement strategies simulated here, which add data points. Falsification is "manipulating research materials, equipment, or processes, or changing or omitting data or results such that the research is not accurately represented in the research record". This is represented by the dropping strategy (which omits observations) and the condition switching and re-pairing strategies (which change labels and pairings without adding or deleting anything). The taxonomy that carries the regulatory and moral weight thus classifies by mechanism: was the data point invented, or was it altered or omitted? Of course, the line is not necessarily clear here, given that an altered data point is equivalent to a removed data point plus a newly added data point. 

In this article I do three things. First, I formalize data tampering as a family of iterative greedy algorithms (dropping, condition switching, re-pairing, duplication, fabrication, and replacement) applied to three of the most common elementary designs: the two-group mean difference (Student's *t*-test), the Pearson correlation, and the 2×2 association (chi-square test). Second, I estimate the effort each strategy requires, with the full design specified under the ADEMP framework [@morris_using_2019; @siepe_simulation_2024] and every estimate accompanied by a Monte Carlo standard error. I invert the design used in previous simulations to understand the impact of *p*-hacking. Where @stefan_big_2023 fix a budget of manipulations and measure the resulting false-positive rate, I fix the goal (statistical significance) and measure the number of tampered data points required to reach it. I take this to be the more ecologically valid framing for deliberate tampering: an analyst who has decided to manufacture a result does not set a budget in advance and stop when it is exhausted, but continues until the *p*-value crosses. It also yields effort directly, in units of edited data points. Because the simulations run without any tampering budget, the complete distribution of required effort is observed, and the proportion of datasets exceeding any given budget is recovered afterwards from the same runs. Third, I show that the effort required is small, that it does not grow with sample size in proportional terms, and that the strategies hardest to detect are among the cheapest.

# Methods

The design is specified under the ADEMP framework (Aims, Data-generating processes, Estimands, Methods, Performance measures) [@morris_using_2019], as operationalized for psychology by @siepe_simulation_2024. The three research compendia containing all simulation code, per-strategy results, and extended rationale are available in the project repository (see *Data and code availability*); this section compresses them.

## Aims

The study is descriptive: there is no closed-form benchmark for the minimum number of tampered points needed to manufacture significance, because that number is a property of a tampering algorithm applied to random data, not an estimator of a population parameter. For each strategy and sample size I aim to estimate (1) how many tampered data points are required to first reach *p* < .05, absolutely and as a proportion of the sample, and how this scales with sample size; (2) the proportion of datasets requiring more than a given share of the sample (post hoc budgets of 10% and 25%), and the proportion for which significance is never attained at all; and (3) how the strategies compare on these measures.

## Data-generating processes

All data are generated under the true null, so every significant result is a false positive produced entirely by the tampering.

- Mean difference. Two independent groups of `n_per_condition` observations each, $Y_{ij} \sim \mathcal{N}(\mu_j, \sigma^2)$ with $\mu_{\text{control}} = \mu_{\text{intervention}} = 0$ and $\sigma = 1$.
- Correlation. A bivariate Normal sample of $n$ observations with $\rho = 0$ and unit variances.
- 2×2 association. Two groups (exposed, not exposed) of `n_per_condition` observations with a binary outcome drawn at the same Bernoulli prevalence in both groups; prevalence is varied (.20 vs .05) to contrast a common with a rare outcome.

Per-group sample sizes are 10 and 25 to 200 in steps of 25 (total $N$ = 20–400; the correlation setting uses the same total $N$s). Fully crossing strategy with sample size (and prevalence) gives 45 conditions in each continuous setting and 54 in the binary setting, each replicated `r format(key$k_iterations, big.mark = ",")` times.

## Estimands

The estimand is `tampers_needed`: the number of tampering operations applied before the test first returns *p* < .05, reported both as a count and as a proportion of the sample. The compendia also record the effect size of each final tampered dataset, but those are not analysed here: this article is concerned with the effort required to manufacture a false positive rather than with the size of the effect that results.

## Methods (tampering strategies)

Every strategy is an iterative greedy procedure: generate a null dataset, test it, and while non-significant, apply one tampering operation targeting the data point most opposed to the desired result, then re-test. @tbl-strategies defines the operations. All strategies push toward a positive effect by convention; tests are two-sided at $\alpha = .05$ (equal-variance *t*-test; Pearson correlation test; chi-square test without continuity correction).

| Strategy | Mean difference | Correlation | 2×2 association |
|:------------------|:----------------------|:----------------------|:----------------------|
| Dropping | Delete the most hostile case (lowest in intervention or highest in control) | Delete the most discordant case (most negative centred cross-product) | Delete a 1 from not-exposed or a 0 from exposed |
| Label switching | Move the most hostile case's *condition label* to the other group, alternating direction to keep groups balanced | *Re-pairing:* swap the *y*-values of two cases so the most discordant case receives the *y* matching its *x*-rank | Move a not-exposed 1 to exposed or an exposed 0 to not-exposed, alternating direction |
| Duplication | Duplicate a case from each group's most favourable 10% | Duplicate one of the most concordant cases | (coincides with fabrication) |
| Fabrication | Add invented cases: control $\sim \mathcal{N}(0, 0.2)$, intervention $\sim \mathcal{N}(5, 0.2)$ | Add a high-leverage concordant pair at $\pm(3, 3)$ with small scatter | Add a 1 to exposed or a 0 to not-exposed |
| Replacement | Replace the most hostile case with an invented one (whichever of the two candidate replacements most lowers *p*) | Replace the most discordant case's *y* with a concordant high-leverage value | (coincides with fabrication) |

: The tampering strategies and their operationalization in each setting. Each operation counts as one tampered data point. For binary data, duplication, fabrication, and replacement collapse onto the same atomic operation (adding a favourable cell) and are treated as a single strategy. {#tbl-strategies}

The label-switching column deserves emphasis because of its forensic properties: condition switching leaves every univariate distribution untouched, and re-pairing preserves both marginal distributions *exactly* (the set of *y*-values is only permuted), so neither is detectable by any univariate check. Swapping *entire* conditions, as opposed to individual cases, is handled analytically rather than simulated: it merely doubles the directional false-positive rate to $\alpha$, adding $\tfrac{1}{2}\alpha$.

No tampering budget is imposed. Each strategy runs until significance is reached or a structural limit is hit, such removing all datat under dropping. A loop guard is included at 10× the sample size as a termination backstop (e.g., a one-step greedy cannot cross the two-sided test's $p \approx 1$ valley when the initial dataset happens to show a strong effect in the *opposite* direction; such iterations, well under 1% of runs, are recorded as never attaining significance). Because the full unbudgeted counts are recorded, the proportion of datasets exceeding *any* tampering budget is computable after the fact; I report budgets of 10% and 25% of the sample.

## Performance measures and Monte Carlo uncertainty

An iteration that never attains significance has no finite tamper count (i.e., is right-censored). All percentile-based summaries (median, 95% prediction interval) are therefore computed over all iterations with the never-significant ones ranked above every finite count; a percentile falling among them is reported as *never* rather than being silently replaced by the survivors' percentile, which would answer a different question ("given that tampering worked, how many edits?") and bias the effort downward. The mean cannot be computed this way at all (the censored counts are unbounded above), so means are reported only as conditional-on-success skew diagnostics. Two intervals are used. The interquartile range is plotted, because one condition (correlation duplication at the smallest sample size) has a 97.5th percentile at nearly nine times the sample and would otherwise set the scale for every panel. The 95% prediction interval is reported in @tbl-results. Effort as a proportion is expressed relative to the *original* sample size, so strategies that append cases, or re-tamper a case already tampered with, can exceed 1.

Every reported estimate carries a Monte Carlo standard error: binomial for proportions, closed-form for means, and nonparametric bootstrap for medians [@morris_using_2019; @siepe_simulation_2024]. In tables, decimal places are set from the MCSE (two significant figures, with the estimate rounded to match) rather than fixed per table. Simulations were parallelized with reproducible L'Ecuyer–CMRG streams, so results are identical regardless of worker count.

# Results

```{r}
#| label: fig-meandiff
#| fig-cap: 'Mean-difference (t-test) setting: the number of tampered data points needed to first reach p < .05 (top) and the same effort as a proportion of the sample (bottom). Points are censoring-aware medians and bars are interquartile ranges. A bar ending in a hollow triangle is open-ended: more than a quarter of those datasets never attained significance, so the upper quartile has no finite value and the bar is drawn to the top of the data range. The lower panel expresses effort relative to the *original* sample size, so strategies that append cases or re-tamper the same case can exceed 1. Strategies are ordered by median proportion tampered. K = 10,000 iterations per condition.'
#| fig-height: 6.2
#| fig-width: 5.35
#| out-width: 100%

tampers |>
  filter(setting == "Mean difference") |>
  order_methods("Mean difference") |>
  plot_effort("Tampered data points to p < .05")
```

```{r}
#| label: fig-correlation
#| fig-cap: 'Correlation setting: tampered data points needed to first reach p < .05 (top) and as a proportion of the sample (bottom). Points are censoring-aware medians and bars are interquartile ranges; a bar ending in a hollow triangle is open-ended, its upper quartile falling among datasets that never attained significance. Effort in the lower panel is relative to the *original* sample size, so strategies that append cases can exceed 1. K = 10,000 iterations per condition.'
#| fig-height: 6.2
#| fig-width: 5.35
#| out-width: 100%

tampers |>
  filter(setting == "Correlation") |>
  order_methods("Correlation") |>
  plot_effort("Tampered data points to p < .05")
```

```{r}
#| label: fig-2x2
#| fig-cap: '2x2 association (chi-square) setting, by outcome prevalence: the number of tampered cells needed to first reach p < .05 (top) and the same effort as a proportion of the sample (bottom). Points are censoring-aware medians and bars are interquartile ranges; a bar ending in a hollow triangle is open-ended, its upper quartile falling among datasets that never attained significance. At prevalence .05 and N = 20, a majority of datasets cannot be pushed to significance at all by dropping, so even the median is not attained and no point is drawn for it. K = 10,000 iterations per condition.'
#| fig-height: 6.4
#| fig-width: 5.35
#| out-width: 100%

d2 <- tampers |>
  filter(setting == "2x2 association") |>
  order_methods("2x2 association") |>
  mutate(prevalence_label = paste0("prevalence = ", prevalence),
         censored = !is.finite(iqr_upper),
         median_f = ifelse(is.finite(median), median, NA_real_),
         prop_median_f = ifelse(is.finite(prop_median), prop_median, NA_real_))
# see plot_effort(): a bar whose upper quartile is never attained is drawn to the top of
# the data range and capped with a hollow triangle rather than being silently dropped
cap_c2 <- max(d2$iqr_upper[is.finite(d2$iqr_upper)], na.rm = TRUE) * 1.04
cap_p2 <- max(d2$prop_iqr_upper[is.finite(d2$prop_iqr_upper)], na.rm = TRUE) * 1.04
d2 <- d2 |> mutate(iqr_upper_f = ifelse(censored, cap_c2, iqr_upper),
                   prop_iqr_upper_f = ifelse(censored, cap_p2, prop_iqr_upper))
cens2 <- d2 |> filter(censored)

# the same two panels as the other settings -- count on top, proportion below -- each
# split by outcome prevalence
p_count <- ggplot(d2, aes(n_total, median_f, color = method)) +
  geom_point(position = position_dodge(width = 27), size = 1) +
  geom_linerange(aes(ymin = iqr_lower, ymax = iqr_upper_f), position = position_dodge(width = 27)) +
  geom_point(data = cens2, aes(y = iqr_upper_f), shape = 24, size = 1.1, fill = "white",
             position = position_dodge(width = 27)) +
  facet_wrap(~ prevalence_label) +
  labs(x = "Total sample size (N)", y = "Tampered cells to p < .05", color = "Strategy") +
  scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty(n = 6)) +
  scale_x_continuous(breaks = x_breaks) +
  scale_color_manual(values = earth_palette) +
  theme_effort()

p_prop <- ggplot(d2, aes(n_total, prop_median_f, color = method)) +
  geom_point(position = position_dodge(width = 27), size = 1) +
  geom_linerange(aes(ymin = prop_iqr_lower, ymax = prop_iqr_upper_f), position = position_dodge(width = 27)) +
  geom_point(data = cens2, aes(y = prop_iqr_upper_f), shape = 24, size = 1.1, fill = "white",
             position = position_dodge(width = 27)) +
  facet_wrap(~ prevalence_label) +
  labs(x = "Total sample size (N)", y = "Tampering operations per original observation",
       color = "Strategy") +
  scale_y_continuous(labels = percent_format(), limits = c(0, NA), breaks = breaks_pretty(n = 6)) +
  scale_x_continuous(breaks = x_breaks) +
  scale_color_manual(values = earth_palette) +
  theme_effort()

p_count + p_prop + plot_layout(ncol = 1, guides = "collect") & theme(legend.position = "bottom")
```

## Effort is small, and falls as a share of the sample as studies grow

Across both continuous settings the absolute number of tampered data points needed to manufacture significance grows with sample size, but only slowly: median counts run from `r fmt_val(key$count20_min)`–`r fmt_val(key$count20_max)` points at $N = 20$ to `r fmt_val(key$count400_min)`–`r fmt_val(key$count400_max)` at $N = 400$, a twentyfold increase in sample size met by roughly a threefold increase in effort (@fig-meandiff, @fig-correlation). Expressed as a share of the sample, effort therefore falls: from `r fmt_pct(key$prop20_min)`–`r fmt_pct(key$prop20_max)` at $N = 20$ to `r fmt_pct(key$prop400_min)`–`r fmt_pct(key$prop400_max)` at $N = 400$, a reduction of between `r fmt_val(key$fold_prop_min, 1)`- and `r fmt_val(key$fold_prop_max, 1)`-fold depending on the strategy. Manufacturing a false positive is not merely no harder in a large study than in a small one; proportionally it is considerably easier. The intuition that a large sample is robust to a few altered observations is exactly backwards. Framed as a budget: at $N = 400$, the proportion of datasets that would require more than 10% of the sample to be tampered is at most `r fmt_pct(key$gt10_400_max)` for the least efficient continuous-setting strategy, and for the cheapest it is `r fmt_pct(key$gt10_400_min)`. Only `r fmt_pct(key$never_400_max)` of datasets resist the least effective strategy altogether at that sample size.

The 2×2 setting is the partial exception that illustrates the mechanism (@fig-2x2). With a rare outcome (prevalence .05) and small samples there is simply not enough signal-bearing data to tamper with: at $N = 20$, `r fmt_pct(key$lowprev_drop_never_20, 0)` of datasets cannot be pushed to significance by dropping at all as the strategy exhausts the table's hostile cells first. Where tampering is structurally possible, rare outcomes are cheap to tip, because each fabricated or relabelled case moves the test statistic disproportionately.

## The stealthiest strategies are among the cheapest

Strategies in every figure are ordered by their median proportion tampered, computed from the data at render time. The forensically critical result is where the label-switching family sits in that ordering. In the two-group design, condition switching is the second cheapest of the five strategies (median `r key$swap_median_100` switched labels at $N = 100$) despite leaving every univariate distribution untouched. Its correlation analogue, re-pairing (median `r key$repair_median_100` re-paired values at $N = 100$), is not the cheapest in its setting. High-leverage replacement and fabrication are cheaper, but it costs about the same as unprincipled case deletion and less than duplication, while preserving both marginal distributions exactly. Neither manipulation is detectable by any check of a variable in isolation; only the joint structure is disturbed. Efficiency and detectability are therefore not in tension for the would-be fraudster: in each setting, at least one of the cheapest strategies is also among the least visible.

```{=latex}
\clearpage
\begin{landscape}
```

```{r}
#| label: tbl-results
#| tbl-cap: "Tampering effort at a representative sample size (total N = 100), by setting and strategy. Values in parentheses are +/-1 Monte Carlo standard error. Never is the proportion of datasets for which significance was never attained, tampering having run to a structural limit. The two budget columns are post hoc tampering budgets: the proportion of datasets needing more than that share of the sample tampered, never-significant datasets exceeding every budget by definition. The median is computed over all iterations with never-significant ones ranked above every finite count, and is also given as a proportion of the sample (Prop.). PI is the 95% prediction interval, the 2.5th-97.5th percentile of the tamper counts across iterations; the figures show interquartile ranges instead, for the reason given in the Methods. D/f/r denotes the duplication, fabrication and replacement strategies, which coincide for binary data."

tampers |>
  filter(n_total == 100) |>
  arrange(setting, prevalence, method) |>
  transmute(Setting = case_when(
              # ASCII "x" deliberately: this string is built in R, and the rendering
              # session runs under a C locale, which converts a non-ASCII character in an
              # R string literal into the literal text "<U+00D7>" on its way into the
              # .tex. Prose elsewhere may use a true multiplication sign because that
              # passes through pandoc rather than R.
              setting == "2x2 association" ~ paste0("2x2, prev. ", prevalence),
              TRUE ~ as.character(setting)),
            Strategy = str_replace(as.character(method),
                                   "Duplication/fabrication/replacement", "D/f/r"),
            # proportions at three decimals and counts at one keep the table inside the
            # text block; the compendia report every column at full MCSE-driven precision
            Never = mcse_cell(never, never_mcse, min_dp = 3, max_dp = 3),
            `> 10%` = mcse_cell(prop_gt_10, prop_gt_10_mcse, min_dp = 3, max_dp = 3),
            `> 25%` = mcse_cell(prop_gt_25, prop_gt_25_mcse, min_dp = 3, max_dp = 3),
            Median = mcse_cell(median, median_mcse, max_dp = 1),
            `95% PI` = interval_cell(pi_lower, pi_upper),
            Prop. = mcse_cell(prop_median, prop_median_mcse, min_dp = 3, max_dp = 3)) |>
  kable(booktabs = TRUE, linesep = "", align = c("l", "l", rep("r", 6))) |>
  kable_styling(latex_options = "hold_position", font_size = 9) |>
  collapse_rows(columns = 1, latex_hline = "major", valign = "top")
```

```{=latex}
\end{landscape}
\clearpage
```

# Discussion

The distance from a null dataset to a statistically significant one is short and cheap. The median effort is a small percentage of the sample across strategies and settings, and for most strategies a minority of datasets require more than 10% of the sample to be tampered. Where tampering fails outright it is because it is structurally impossible rather than merely expensive. Specifically, the rare-outcome 2×2 cells, where too few hostile cases exist to move the table at all. The share of the sample that must be altered falls as samples grow, by roughly an order of magnitude across the range examined, so the intuition that a large study is protected by its size is not merely optimistic but inverted; this extends the finding for *p*-hacking that larger samples alone do not reduce its success rate [@stefan_big_2023]. The strategies that preserve univariate distributions (condition switching and re-pairing) are among the most effective. Further research is necessary to understand the relative statistical detectability of each of these strategies. 

Adding invented cases, deleting inconvenient ones, and relabelling existing ones all require a comparable number of edits. The one clause of the definitions that unifies them is a consequence criterion rather than a mechanism one, such that the research is not accurately represented in the research record, and every strategy here satisfies it identically. Each produces a result that did not exist before.

The practical asymmetry, then, is not in the acts but in how they are regarded. Excluding inconvenient participants is admitted by 38–43% of researchers and rated as close to defensible; falsifying data is admitted by under 2% and rated as indefensible [@john_measuring_2012]. When respondents are given provable deniability, admitted fabrication and falsification converge to indistinguishable rates [@gopalakrishna_prevalence_2022] — evidence that the gap is substantially one of reporting rather than of behaviour. And since perceived seriousness is the strongest available predictor of engagement [@entradas_shades_2026], a seriousness ranking that is mis-set on the dimension these results measure is not merely an intellectual inconsistency: it is a live cause of the behaviour. Deleting a hostile case and inventing a favourable one are, in effort and in damage to the evidence, the same act.

The ethics of reporting the results of this simulation study were given considerable thought. Although the simulations were designed to make a continuum visible rather than to provide a recipe, in principle, this article documents knowledge that could be weaponised by bad actors. At the same time, the tampering strategies simulated here have been publicly described for years [@chambers_seven_2017; @brown_condition_2020]; what is added here is quantification. This quantification could help change future scientific behavior, by raising awareness around how impactful even a very small number of data tampers can be.

The strategies employed in this simulation are deliberately simple and greedy. Real, detection-aware fraud could plausibly trade efficiency for plausibility. However, there are of course many cases of detection-unaware fraud caught because of impossible or implausible results and data, and the prevalence of undetected, relatively expert research fraud is unknown by its very nature. Conversely, several naive strategies here leave potentially detectable signatures (e.g., truncated distributions after dropping; leverage outliers after fabrication; shifted marginal prevalence in the 2×2). Future research is needed to examine such detectability via statistical means and modern forensic tools [@brown_grim_2017; @heathers_recovering_2018; @wilkinson_inspect-sr_2025]. However, the other distribution-preserving strategies imply that relatively statistically invisible tampering is cheap and effective. Auditable chain-of-custody for data, such as those employed in medical clinical trials that allowed the the US FDA to detect abnormalities in the trial, as discussed in the introduction, would be needed to prevent or detect such forms of tampering [@us_food_and_drug_administration_center_for_drug_evaluation_and_research_proposal_2026]. Infrastructure for auditable chain-of-custody is rare in other fields.  

<!-- A degenerate case in correlation duplication at very small N. With $n = 20$ the duplication pool is only two points. Duplicating those two points repeatedly lets them dominate the sample, and two things follow: the effort required becomes extreme (a 97.5th percentile at nearly nine times the original sample size, against a median of 0.45), and the resulting configuration can be strongly negatively correlated, which a two-sided test accepts as significant even though the manufactured effect points the wrong way. This is a property of the specified procedure rather than of its implementation — the reference and optimized implementations agree exactly here — but it is a pathology, and it is confined to the smallest sample size. Requiring a minimum pool size, or a stopping rule that demands the effect be in the intended direction, would remove it. -->
<!-- Algorithmic greedy myopia is part of the procedure. One-step greedy occasionally stalls — most notably when the initial dataset shows a strong effect opposite to the tampering direction, where a two-sided test's $p \approx 1$ valley cannot be crossed one step at a time. These rare failures are recorded as never attaining significance and are a true property of naive tampering, not an artefact; a less myopic tamperer would do better, so the effort estimates here are, if anything, conservative upper bounds. -->

# Data and code availability {.unnumbered}

All simulation code and results, including extended ADEMP specifications, per-condition tables, and Monte Carlo standard errors for every estimate, are available at <https://github.com/ianhussey/simulation-impact-of-data-tampering>.

# References {.unnumbered}

::: {#refs}
:::
