---
title: |
  Little big lies:\
  How little data tampering is needed to manufacture statistically significant results
author:
  - name: Ian Hussey
    email: ian.hussey@unibe.ch
    affiliations:
      - ref: unibe
    attributes:
        corresponding: true
affiliations:
  - id: unibe
    name: University of Bern
# TODO update the qualitative quantities in the abstract ("a few per cent", "a minority")
# against the rendered numbers once the final HPC run has landed; the abstract is the
# one place in this document where values cannot be computed at render time.
abstract: 'Fabricating a statistically significant result does not require fabricating a dataset. We simulate a dishonest analyst who starts from data generated under the true null hypothesis and alters, deletes, or invents one data point at a time — always the point that most helps — re-running the test after every change until *p* < .05. We formalize this as a family of iterative greedy tampering strategies (dropping cases, switching condition labels, re-pairing values, duplicating cases, fabricating cases, and replacing cases) across three common designs: the two-group mean difference (Student''s t-test), the Pearson correlation, and the 2x2 association (chi-square test). Each strategy runs until significance is reached or tampering becomes structurally impossible, and the estimand is the effort required: the number of tampered data points, both absolutely and as a proportion of the sample. Three findings recur. First, effort is small — median tamper counts correspond to a few per cent of the sample at realistic sample sizes — and the proportion of the sample that must be tampered does not grow with sample size, so large studies are no safer proportionally than small ones. Second, the strategies that are hardest to detect, because they preserve every univariate distribution (condition switching; value re-pairing), are among the cheapest. Third, because the greedy procedure stops at the first crossing, the manufactured effect sizes cluster just above the critical effect size for the sample size at hand: large enough to publish, too small to look suspicious. The practical distinction between questionable research practices and fabrication is a difference in keystrokes, not in effort or in the damage done to the evidence.'
format:
  elsevier-pdf:
    keep-tex: true
    latex-tinytex: false
    # The elsevier extension defaults to cite-method: natbib, which formats the
    # reference list with elsarticle-harv.bst and ignores csl. Switch to
    # citeproc so the APA 7th CSL below governs both in-text and list style.
    cite-method: citeproc
    # elsarticle.cls loads natbib unconditionally unless given nonatbib, and
    # natbib then errors on the CSLReferences list citeproc emits ("Bibliography
    # not compatible with author-year citations"). nonatbib is the class's own
    # opt-out and is the supported way to pair elsarticle with citeproc.
    classoption:
      - nonatbib
    fig-format: pdf # vector figures embedded in the PDF (no rasterization)
    journal:
      name: psyarxiv
      formatting: preprint
      cite-style: authoryear
execute:
  echo: false
header-includes:
      - \usepackage{threeparttable}
      - \usepackage{unicode-math}
      - \usepackage[margin=4cm]{geometry} # tighter page margins
      - \usepackage{etoolbox}
      - \AtBeginEnvironment{longtable}{\footnotesize} # smaller font inside tables
      - \AtBeginEnvironment{tabular}{\small}
      # double the paragraph indent. AtBeginDocument because the elsevier
      # template sets \parindent after header-includes; 2\parindent doubles
      # whatever it set rather than hardcoding a value.
      - \AtBeginDocument{\setlength{\parindent}{2\parindent}}
bibliography: bibliography.bib
csl: apa.csl # APA Style 7th edition
# Make each in-text citation a hyperlink to its reference-list entry. citeproc
# always emits the \bibitem anchors, but without this nothing points at them.
link-citations: true
# In-text citations use @key syntax, so the reference list is cited-only: an
# entry appears if and only if something cites it, and an unmatched key fails
# loudly at render. Do not reinstate `nocite: @*`, which would silently print
# uncited entries and disable that check.
knitr:
  opts_chunk:
    dev: cairo_pdf # vector device: razor-sharp lines/text at any zoom
    message: false
    warning: false
---

<!-- TODO after the final HPC run lands:
     (1) re-read the computed strategy orderings (they are derived from the data at
         render time below, so figures/tables self-update, but the prose claims about
         WHICH strategies are cheapest should be checked against them);
     (2) update the abstract's qualitative quantities;
     (3) complete the Spiess et al. citation (dfbeta/influential-observation
         fragility literature) or drop the sentence that would cite it. -->

```{r}
#| label: setup
#| include: false

# Every number, figure, and table in this manuscript is computed at render time from
# the simulation caches in ../../data (written by the three research compendia in
# ../../code, which are the canonical source for the simulation and summarising code;
# the helpers below are trimmed copies). Re-rendering after a new simulation run
# updates the entire results section with no manual editing.

library(dplyr)
library(tidyr)
library(tibble)
library(purrr)
library(readr)
library(stringr)
library(forcats)
library(ggplot2)
library(scales)
library(patchwork)
library(knitr)
library(kableExtra)

options(scipen = 999)

data_dir <- "../../data"

# fail loudly and informatively if the full-run caches are not all present
# (e.g., when rendering before an HPC run has finished and been copied down)
expected_caches <- c(
  paste0("simulation_iterative_", c("dropping", "swaps", "duplication", "fabrication", "replacement"), "_meandiff_full_k10000.rds"),
  paste0("simulation_iterative_", c("dropping", "repairing", "duplication", "fabrication", "replacement"), "_correlation_full_k10000.rds"),
  paste0("simulation_iterative_", c("fabrication", "dropping", "switching"), "_2x2_full_k10000.rds"),
  paste0("simulation_iterative_", c("fabrication", "dropping", "switching"), "_2x2_lowprev_full_k10000.rds")
)
missing_caches <- expected_caches[!file.exists(file.path(data_dir, expected_caches))]
if (length(missing_caches) > 0) {
  stop("Missing simulation caches (run the compendia in ../../code first):\n  ",
       paste(missing_caches, collapse = "\n  "))
}

# ---- assemble one long per-iteration data frame across all settings ----------------

read_meandiff <- function(stem, method) {
  read_rds(file.path(data_dir, paste0("simulation_iterative_", stem, "_meandiff_full_k10000.rds"))) |>
    transmute(setting = "Mean difference", method = method, prevalence = NA_real_,
              n_total = n_per_condition * 2, tampers_needed, reached_sig, es = hedges_g)
}
read_corr <- function(stem, method) {
  read_rds(file.path(data_dir, paste0("simulation_iterative_", stem, "_correlation_full_k10000.rds"))) |>
    transmute(setting = "Correlation", method = method, prevalence = NA_real_,
              n_total = n, tampers_needed, reached_sig, es = pearson_r)
}
read_2x2 <- function(stem, method, lowprev = FALSE) {
  f <- paste0("simulation_iterative_", stem, "_2x2", if (lowprev) "_lowprev", "_full_k10000.rds")
  read_rds(file.path(data_dir, f)) |>
    transmute(setting = "2x2 association", method = method, prevalence,
              n_total = n_per_condition * 2, tampers_needed, reached_sig, es = NA_real_)
}

raw <- bind_rows(
  read_meandiff("dropping",    "Dropping"),
  read_meandiff("swaps",       "Condition switching"),
  read_meandiff("duplication", "Duplication"),
  read_meandiff("fabrication", "Fabrication"),
  read_meandiff("replacement", "Replacement"),
  read_corr("dropping",    "Dropping"),
  read_corr("repairing",   "Re-pairing"),
  read_corr("duplication", "Duplication"),
  read_corr("fabrication", "Fabrication"),
  read_corr("replacement", "Replacement"),
  read_2x2("fabrication", "Duplication/fabrication/replacement"),
  read_2x2("dropping",    "Dropping"),
  read_2x2("switching",   "Condition switching"),
  read_2x2("fabrication", "Duplication/fabrication/replacement", lowprev = TRUE),
  read_2x2("dropping",    "Dropping",                            lowprev = TRUE),
  read_2x2("switching",   "Condition switching",                 lowprev = TRUE)
) |>
  mutate(setting = factor(setting, levels = c("Mean difference", "Correlation", "2x2 association")))

# ---- Monte Carlo standard errors (trimmed copies of the compendia helpers) ---------

mcse_mean <- function(x){
  x <- x[!is.na(x)]
  if (length(x) < 2) return(NA_real_)
  sd(x) / sqrt(length(x))
}

mcse_proportion <- function(p, k){
  if (is.na(p) || k < 1) return(NA_real_)
  sqrt(p * (1 - p) / k)
}

mcse_median <- function(x, n_boot = 1000, seed = 42){
  x <- x[!is.na(x)]
  n <- length(x)
  if (n < 2) return(NA_real_)
  old_seed <- if (exists(".Random.seed", envir = .GlobalEnv)) get(".Random.seed", envir = .GlobalEnv) else NULL
  set.seed(seed)
  boot_medians <- replicate(n_boot, median(x[sample.int(n, n, replace = TRUE)]))
  if (!is.null(old_seed)) assign(".Random.seed", old_seed, envir = .GlobalEnv)
  if (!all(is.finite(boot_medians))) return(NA_real_) # a censored median has no estimable MCSE
  sd(boot_medians)
}

q_censored <- function(x, probs){
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_real_)
  unname(quantile(x, probs = probs, names = FALSE, type = 7))
}

safe_mean <- function(x){ x <- x[is.finite(x)]; if (length(x) == 0) NA_real_ else mean(x) }
safe_sd   <- function(x){ x <- x[is.finite(x)]; if (length(x) < 2)  NA_real_ else sd(x) }

# ---- censoring-aware summaries ------------------------------------------------------
# An iteration that never attained significance (tampering ran to a structural limit
# without crossing alpha) has no finite tamper count; coding it Inf ranks it above every
# finite count, so percentiles among the finite counts are exact and percentiles among
# the censored cases return Inf (reported as "never"). See the compendia for rationale.

tampers <- raw |>
  mutate(tampers_censored = ifelse(reached_sig, tampers_needed, Inf)) |>
  group_by(setting, method, prevalence, n_total) |>
  summarize(k = n(),
            never = mean(!reached_sig),
            never_mcse = mcse_proportion(never, k),
            prop_gt_10 = mean(tampers_censored > 0.10 * n_total),
            prop_gt_10_mcse = mcse_proportion(prop_gt_10, k),
            prop_gt_25 = mean(tampers_censored > 0.25 * n_total),
            prop_gt_25_mcse = mcse_proportion(prop_gt_25, k),
            median = q_censored(tampers_censored, 0.5),
            median_mcse = mcse_median(tampers_censored),
            pi_lower = q_censored(tampers_censored, 0.025),
            pi_upper = q_censored(tampers_censored, 0.975),
            mean_reached = safe_mean(tampers_needed[reached_sig]),
            mean_reached_mcse = mcse_mean(tampers_needed[reached_sig]),
            .groups = "drop") |>
  mutate(prop_median = median / n_total,
         prop_median_mcse = median_mcse / n_total,
         prop_pi_lower = pi_lower / n_total,
         prop_pi_upper = pi_upper / n_total)

bias <- raw |>
  filter(setting != "2x2 association") |>
  group_by(setting, method, n_total) |>
  summarize(es_median = median(es, na.rm = TRUE),
            es_median_mcse = mcse_median(es),
            es_pi_lower = quantile(es, probs = 0.025, na.rm = TRUE),
            es_pi_upper = quantile(es, probs = 0.975, na.rm = TRUE),
            .groups = "drop")

# ---- strategy display order: computed from the data, not hardcoded ------------------
# Ordered by impact within each setting: ascending mean (over sample sizes) of the
# censoring-aware median proportion of the sample tampered. Strategies whose median is
# censored at some N sort by the finite cells they have.
method_order <- tampers |>
  group_by(setting, method) |>
  summarize(o = mean(prop_median[is.finite(prop_median)]), .groups = "drop") |>
  arrange(setting, o)

order_methods <- function(df, setting_name) {
  lv <- method_order |> filter(setting == setting_name) |> pull(method)
  df |> mutate(method = factor(method, levels = lv))
}

# ---- shared plotting pieces ---------------------------------------------------------

x_breaks <- c(10, seq(25, 200, by = 25)) * 2

plot_effort <- function(df, y_count_lab, dodge = 27) {
  # an interval bar is omitted (NA) where its upper percentile falls among the
  # never-significant datasets, since no finite value exists to draw it to
  df <- df |> mutate(pi_upper_f = ifelse(is.finite(pi_upper), pi_upper, NA_real_),
                     prop_pi_upper_f = ifelse(is.finite(prop_pi_upper), prop_pi_upper, NA_real_),
                     median_f = ifelse(is.finite(median), median, NA_real_),
                     prop_median_f = ifelse(is.finite(prop_median), prop_median, NA_real_))
  p1 <- ggplot(df, aes(n_total, median_f, color = method)) +
    geom_point(position = position_dodge(width = dodge), size = 1) +
    geom_linerange(aes(ymin = pi_lower, ymax = pi_upper_f), position = position_dodge(width = dodge)) +
    labs(x = "Total sample size (N)", y = y_count_lab, color = "Strategy") +
    scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty(n = 6)) +
    scale_x_continuous(breaks = x_breaks) +
    scale_color_viridis_d(begin = 0, end = 0.85) +
    theme_linedraw() +
    theme(panel.grid.minor = element_blank())
  p2 <- ggplot(df, aes(n_total, prop_median_f, color = method)) +
    geom_point(position = position_dodge(width = dodge), size = 1) +
    geom_linerange(aes(ymin = prop_pi_lower, ymax = prop_pi_upper_f), position = position_dodge(width = dodge)) +
    labs(x = "Total sample size (N)", y = "Proportion of sample tampered", color = "Strategy") +
    scale_y_continuous(labels = percent_format(), limits = c(0, NA), breaks = breaks_pretty(n = 6)) +
    scale_x_continuous(breaks = x_breaks) +
    scale_color_viridis_d(begin = 0, end = 0.85) +
    theme_linedraw() +
    theme(panel.grid.minor = element_blank())
  p1 + p2 + plot_layout(ncol = 1, guides = "collect")
}

# ---- table cell formatting (MCSE-driven decimal places; Morris et al., 2019) --------

dp_from_mcse <- function(mcse, min_dp = 0, max_dp = 4){
  mcse <- mcse[is.finite(mcse) & mcse > 0]
  if (length(mcse) == 0) return(min_dp)
  min(max_dp, max(min_dp, max(ceiling(-log10(mcse)) + 1)))
}
fmt_num <- function(x, dp){
  ifelse(is.na(x), "NA", ifelse(is.finite(x), formatC(x, format = "f", digits = dp), "never"))
}
mcse_cell <- function(estimate, mcse, min_dp = 0, max_dp = 4){
  dp <- dp_from_mcse(mcse, min_dp = min_dp, max_dp = max_dp)
  ifelse(is.finite(estimate) & is.finite(mcse),
         paste0(fmt_num(estimate, dp), " (", fmt_num(mcse, dp), ")"),
         fmt_num(estimate, dp))
}
interval_cell <- function(lower, upper, dp = 0){
  paste0("[", fmt_num(lower, dp), ", ", fmt_num(upper, dp), "]")
}
fmt_pct <- function(x, dp = 1) paste0(formatC(100 * x, format = "f", digits = dp), "\\%")

# ---- key quantities used inline in the prose ---------------------------------------

grab <- function(s, m, n, col, prev = NA) {
  d <- tampers |> filter(setting == s, method == m, n_total == n)
  if (!is.na(prev)) d <- d |> filter(prevalence == prev)
  d[[col]][1]
}

continuous_400 <- tampers |>
  filter(setting != "2x2 association", n_total == 400, is.finite(prop_median))

key <- list(
  # range of the median proportion tampered at the largest N, across the two continuous settings
  prop400_min = min(continuous_400$prop_median),
  prop400_max = max(continuous_400$prop_median),
  # cheapest distribution-preserving strategies at N = 100
  swap_median_100   = grab("Mean difference", "Condition switching", 100, "median"),
  repair_median_100 = grab("Correlation", "Re-pairing", 100, "median"),
  # proportion of datasets needing more than 10% of the sample at N = 400, across continuous strategies
  gt10_400_max = max(tampers |> filter(setting != "2x2 association", n_total == 400) |> pull(prop_gt_10)),
  gt10_400_min = min(tampers |> filter(setting != "2x2 association", n_total == 400) |> pull(prop_gt_10)),
  never_400_max = max(tampers |> filter(setting != "2x2 association", n_total == 400) |> pull(never)),
  # structural impossibility at the rare-outcome extreme
  lowprev_drop_never_20 = grab("2x2 association", "Dropping", 20, "never", prev = 0.05),
  k_iterations = max(tampers$k)
)
```

# Introduction

Start from a dataset in which the null hypothesis is true by construction. How many data points must a dishonest analyst alter, delete, or invent before a standard significance test returns *p* < .05? This number — the *effort* of manufacturing a false positive — is the estimand of this article. We express it as a count of tampered data points and as a proportion of the sample, because a count is what makes the problem concrete: it is the number of keystrokes between an honest null result and a publishable finding.

The motivating context is the continuum between questionable research practices (QRPs) and outright fabrication. Unprincipled exclusion of inconvenient participants is usually filed under *p*-hacking, is admitted by a substantial fraction of researchers when surveyed [@john_measuring_2012; @simmons_false-positive_2011], and is often treated as common-but-merely-iffy. Typing the *other* condition's label into those same rows is fabrication — rarely admitted [@fanelli_how_2009] and a firing offence. Yet the two acts can differ by nothing more than which cells of the spreadsheet are edited. If the effort required and the damage done to the evidence are also near-identical, then treating the first as common and the second as vanishingly rare risks badly miscalibrating how much fabrication actually occurs.

The specific manipulations simulated here have a public history. @chambers_seven_2017 described condition switching as the second tip in his "dirty dozen" recipe for cheating; @brown_condition_2020 posted a small simulation of that tip showing how few switched labels a *t*-test needs, which is the proximate origin of this project (his simulation is reproduced in our repository). Related work on the fragility of published findings shows that removing a handful of influential observations often flips significance <!-- TODO cite Spiess et al. / the dfbeta influential-observation literature -->, and forensic tools exist to detect some tampering signatures after the fact [@brown_grim_2017; @heathers_recovering_2018; @wilkinson_inspect-sr_2025]. What has been missing is a systematic quantification of the *forward* problem — how much tampering different strategies require — specified and reported to current simulation-study standards.

In this article we therefore do three things. First, we formalize data tampering as a family of iterative greedy algorithms — dropping, condition switching, re-pairing, duplication, fabrication, and replacement — applied to three of the most common elementary designs: the two-group mean difference (Student's *t*-test), the Pearson correlation, and the 2×2 association (chi-square test). Second, we estimate the effort each strategy requires, with the full design specified under the ADEMP framework [@morris_using_2019; @siepe_simulation_2024] and every estimate accompanied by a Monte Carlo standard error. Because the simulations run without any tampering budget — each dataset is tampered until significance is reached or tampering becomes structurally impossible — the complete distribution of required effort is observed, and the proportion of datasets exceeding any budget (we report 10% and 25% of the sample) is recovered afterwards from the same runs. Third, we show that the strategies hardest to detect are among the cheapest, and that the manufactured effect sizes cluster just above the critical effect size — large enough to publish, too small to look suspicious.

# Methods

The design is specified under the ADEMP framework (Aims, Data-generating processes, Estimands, Methods, Performance measures) [@morris_using_2019], as operationalized for psychology by @siepe_simulation_2024. The three research compendia containing all simulation code, per-strategy results, and extended rationale are available in the project repository (see *Data and code availability*); this section compresses them.

## Aims

The study is descriptive: there is no closed-form benchmark for the minimum number of tampered points needed to manufacture significance, because that number is a property of a tampering algorithm applied to random data, not an estimator of a population parameter. For each strategy and sample size we aim to estimate (1) how many tampered data points are required to first reach *p* < .05, absolutely and as a proportion of the sample, and how this scales with sample size; (2) the proportion of datasets requiring more than a given share of the sample (post hoc budgets of 10% and 25%), and the proportion for which significance is never attained at all; (3) the effect-size bias of the final tampered dataset relative to the true effect of zero; and (4) how the strategies compare on these measures.

## Data-generating processes

All data are generated under the true null, so every significant result is a false positive produced entirely by the tampering.

- **Mean difference.** Two independent groups of `n_per_condition` observations each, $Y_{ij} \sim \mathcal{N}(\mu_j, \sigma^2)$ with $\mu_{\text{control}} = \mu_{\text{intervention}} = 0$ and $\sigma = 1$.
- **Correlation.** A bivariate Normal sample of $n$ observations with $\rho = 0$ and unit variances.
- **2×2 association.** Two groups (exposed, not exposed) of `n_per_condition` observations with a binary outcome drawn at the same Bernoulli prevalence in both groups; prevalence is varied (.20 vs .05) to contrast a common with a rare outcome.

Per-group sample sizes are 10 and 25 to 200 in steps of 25 (total $N$ = 20–400; the correlation setting uses the same total $N$s). Fully crossing strategy with sample size (and prevalence) gives 45 conditions in each continuous setting and 54 in the binary setting, each replicated `r format(key$k_iterations, big.mark = ",")` times.

## Estimands

The primary estimand is `tampers_needed`: the number of tampering operations applied before the test first returns *p* < .05. The secondary estimand is the effect size of the final tampered dataset — Hedges' *g* in the mean-difference setting and Pearson's *r* in the correlation setting — which, because the true effect is zero, *is* the bias the tampering introduces; it is reported signed, since the sign is the direction of the manufactured effect. (No standardized effect-size bias is computed in the binary setting.)

## Methods (tampering strategies)

Every strategy is an iterative greedy procedure: generate a null dataset, test it, and while non-significant, apply one tampering operation targeting the data point most opposed to the desired result, then re-test. @tbl-strategies defines the operations. All strategies push toward a positive effect by convention; tests are two-sided at $\alpha = .05$ (equal-variance *t*-test; Pearson correlation test; chi-square test without continuity correction).

| Strategy | Mean difference | Correlation | 2×2 association |
|:------------------|:----------------------|:----------------------|:----------------------|
| Dropping | Delete the most hostile case (lowest in intervention or highest in control) | Delete the most discordant case (most negative centred cross-product) | Delete a 1 from not-exposed or a 0 from exposed |
| Label switching | Move the most hostile case's *condition label* to the other group, alternating direction to keep groups balanced | *Re-pairing:* swap the *y*-values of two cases so the most discordant case receives the *y* matching its *x*-rank | Move a not-exposed 1 to exposed or an exposed 0 to not-exposed, alternating direction |
| Duplication | Duplicate a case from each group's most favourable 10% | Duplicate one of the most concordant cases | (coincides with fabrication) |
| Fabrication | Add invented cases: control $\sim \mathcal{N}(0, 0.2)$, intervention $\sim \mathcal{N}(5, 0.2)$ | Add a high-leverage concordant pair at $\pm(3, 3)$ with small scatter | Add a 1 to exposed or a 0 to not-exposed |
| Replacement | Replace the most hostile case with an invented one (whichever of the two candidate replacements most lowers *p*) | Replace the most discordant case's *y* with a concordant high-leverage value | (coincides with fabrication) |

: The tampering strategies and their operationalization in each setting. Each operation counts as one tampered data point. For binary data, duplication, fabrication, and replacement collapse onto the same atomic operation (adding a favourable cell) and are treated as a single strategy. {#tbl-strategies}

The label-switching column deserves emphasis because of its forensic properties: condition switching leaves every univariate distribution untouched, and re-pairing preserves both marginal distributions *exactly* (the set of *y*-values is only permuted), so neither is detectable by any univariate check. Swapping *entire* conditions, as opposed to individual cases, is handled analytically rather than simulated: it merely doubles the directional false-positive rate to $\alpha$, adding $\tfrac{1}{2}\alpha$.

**No tampering budget is imposed.** Each strategy runs until significance is reached or a structural limit is hit — the data are exhausted under dropping, no improving re-pairing exists, a donor cell empties in the 2×2 — with a loop guard at 10× the sample size as a termination backstop for pathological greedy stalls (e.g., a one-step greedy cannot cross the two-sided test's $p \approx 1$ valley when the initial dataset happens to show a strong effect in the *opposite* direction; such iterations, well under 1% of runs, are recorded as never attaining significance). Because the full unbudgeted counts are recorded, the proportion of datasets exceeding *any* tampering budget is computable after the fact; we report budgets of 10% and 25% of the sample.

## Performance measures and Monte Carlo uncertainty

An iteration that never attains significance has no finite tamper count — it is right-censored. All percentile-based summaries (median, 95% prediction interval) are therefore computed over all iterations with the never-significant ones ranked above every finite count; a percentile falling among them is reported as *never* rather than being silently replaced by the survivors' percentile, which would answer a different question ("given that tampering worked, how many edits?") and bias the effort downward. The mean cannot be computed this way at all (the censored counts are unbounded above), so means are reported only as conditional-on-success skew diagnostics. The 95% prediction interval is the 2.5th–97.5th percentile of the tamper counts across iterations: the range within which a single new study's required effort would fall 95% of the time.

Every reported estimate carries a Monte Carlo standard error: binomial for proportions, closed-form for means, and nonparametric bootstrap for medians [@morris_using_2019; @siepe_simulation_2024]. In tables, decimal places are set from the MCSE (two significant figures, with the estimate rounded to match) rather than fixed per table. Simulations were parallelized with reproducible L'Ecuyer–CMRG streams, so results are identical regardless of worker count.

# Results

```{r}
#| label: fig-meandiff
#| fig-cap: 'Mean-difference (t-test) setting: the number of tampered data points needed to first reach p < .05 (top) and the same effort as a proportion of the sample (bottom). Points are censoring-aware medians and bars are 95% prediction intervals (2.5th–97.5th percentile across iterations); a bar is omitted where its upper percentile falls among datasets that never attained significance. Strategies are ordered by median proportion tampered. K = 10,000 iterations per condition.'
#| fig-height: 6.5
#| fig-width: 7.5

tampers |>
  filter(setting == "Mean difference") |>
  order_methods("Mean difference") |>
  plot_effort("Tampered data points to p < .05")
```

```{r}
#| label: fig-correlation
#| fig-cap: 'Correlation setting: tampered data points needed to first reach p < .05 (top) and as a proportion of the sample (bottom). Points are censoring-aware medians, bars 95% prediction intervals; a bar is omitted where its upper percentile is never attained. K = 10,000 iterations per condition.'
#| fig-height: 6.5
#| fig-width: 7.5

tampers |>
  filter(setting == "Correlation") |>
  order_methods("Correlation") |>
  plot_effort("Tampered data points to p < .05")
```

```{r}
#| label: fig-2x2
#| fig-cap: '2×2 association (chi-square) setting, by outcome prevalence: tampered cells needed to first reach p < .05 as a proportion of the sample. Points are censoring-aware medians, bars 95% prediction intervals; a bar is omitted where its upper percentile is never attained. At prevalence .05 and small N, a majority of datasets cannot be pushed to significance at all by dropping (the median itself is never attained and no point is drawn). K = 10,000 iterations per condition.'
#| fig-height: 4
#| fig-width: 7.5

d2 <- tampers |>
  filter(setting == "2x2 association") |>
  order_methods("2x2 association") |>
  mutate(prevalence_label = paste0("prevalence = ", prevalence),
         prop_pi_upper_f = ifelse(is.finite(prop_pi_upper), prop_pi_upper, NA_real_),
         prop_median_f = ifelse(is.finite(prop_median), prop_median, NA_real_))

ggplot(d2, aes(n_total, prop_median_f, color = method)) +
  geom_point(position = position_dodge(width = 27), size = 1) +
  geom_linerange(aes(ymin = prop_pi_lower, ymax = prop_pi_upper_f), position = position_dodge(width = 27)) +
  facet_wrap(~ prevalence_label) +
  labs(x = "Total sample size (N)", y = "Proportion of sample tampered", color = "Strategy") +
  scale_y_continuous(labels = percent_format(), limits = c(0, NA), breaks = breaks_pretty(n = 6)) +
  scale_x_continuous(breaks = x_breaks) +
  scale_color_viridis_d(begin = 0, end = 0.85) +
  theme_linedraw() +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")
```

## Effort is small and proportionally flat in sample size

Across both continuous settings, the absolute number of tampered data points needed to manufacture significance grows with sample size, but the *proportion* of the sample does not (@fig-meandiff, @fig-correlation): at the largest sample size examined ($N = 400$), the median proportion tampered ranges from `r fmt_pct(key$prop400_min)` to `r fmt_pct(key$prop400_max)` across strategies. Manufacturing a false positive is no harder, proportionally, in a large study than in a small one. Framed as a budget: at $N = 400$, the proportion of datasets that would require more than 10% of the sample to be tampered is at most `r fmt_pct(key$gt10_400_max)` for the least efficient continuous-setting strategy, and for the cheapest it is `r fmt_pct(key$gt10_400_min)`. Only `r fmt_pct(key$never_400_max)` of datasets resist the least effective strategy altogether at that sample size.

The 2×2 setting is the partial exception that illustrates the mechanism (@fig-2x2). With a rare outcome (prevalence .05) and small samples there is simply not enough signal-bearing data to tamper with: at $N = 20$, `r fmt_pct(key$lowprev_drop_never_20, 0)` of datasets cannot be pushed to significance by dropping at all — the strategy exhausts the table's hostile cells first. Where tampering *is* structurally possible, rare outcomes are cheap to tip, because each fabricated or relabelled case moves the test statistic disproportionately.

## The stealthiest strategies are among the cheapest

Strategies in every figure are ordered by their median proportion tampered, computed from the data at render time. The forensically critical result is where the label-switching family sits in that ordering: condition switching (median `r key$swap_median_100` switched labels at $N = 100$) and re-pairing (median `r key$repair_median_100` re-paired values at $N = 100$) reach significance with effort comparable to or lower than the visibly destructive strategies, while leaving every univariate distribution intact — re-pairing preserves both marginals *exactly*. No check of either variable in isolation can detect them; only the joint structure is tampered. Efficiency and detectability are thus not in tension for the would-be fraudster: the cheap strategies include the invisible ones.

```{r}
#| label: tbl-results
#| tbl-cap: "Tampering effort at a representative sample size (total N = 100), by setting and strategy. Values in parentheses are ±1 Monte Carlo standard error. Never is the proportion of datasets for which significance was never attained, tampering having run to a structural limit. The two budget columns are post hoc tampering budgets: the proportion of datasets needing more than that share of the sample tampered, never-significant datasets exceeding every budget by definition. The median is computed over all iterations with never-significant ones ranked above every finite count, and is also given as a proportion of the sample (Prop.); prediction intervals are shown in the figures. D/f/r denotes the duplication, fabrication and replacement strategies, which coincide for binary data."

tampers |>
  filter(n_total == 100) |>
  arrange(setting, prevalence, method) |>
  transmute(Setting = case_when(
              setting == "2x2 association" ~ paste0("2×2, prev. ", prevalence),
              TRUE ~ as.character(setting)),
            Strategy = str_replace(as.character(method),
                                   "Duplication/fabrication/replacement", "D/f/r"),
            # proportions at three decimals and counts at one keep the table inside the
            # text block; the compendia report every column at full MCSE-driven precision
            Never = mcse_cell(never, never_mcse, min_dp = 3, max_dp = 3),
            `> 10%` = mcse_cell(prop_gt_10, prop_gt_10_mcse, min_dp = 3, max_dp = 3),
            `> 25%` = mcse_cell(prop_gt_25, prop_gt_25_mcse, min_dp = 3, max_dp = 3),
            Median = mcse_cell(median, median_mcse, max_dp = 1),
            Prop. = mcse_cell(prop_median, prop_median_mcse, min_dp = 3, max_dp = 3)) |>
  kable(booktabs = TRUE, linesep = "", align = c("l", "l", rep("r", 5))) |>
  kable_styling(latex_options = c("hold_position", "scale_down")) |>
  collapse_rows(columns = 1, latex_hline = "major", valign = "top")
```

## The manufactured effect is just large enough

```{r}
#| label: fig-bias
#| fig-cap: 'Effect-size bias of the final tampered datasets, relative to the true effect of zero: Hedges'' g in the mean-difference setting (left) and Pearson''s r in the correlation setting (right). Points are medians across iterations, bars 95% prediction intervals; strategies are coloured by family, so that condition switching (mean difference) and re-pairing (correlation) — the two distribution-preserving strategies — share a colour. The dashed curve is the critical effect size — the smallest effect significant at $\alpha$ = .05 (two-sided) for the original sample size; strategies that change N (dropping, duplication, fabrication) make this an approximation. Because the greedy procedure stops at the first crossing, the manufactured effects sit just above this curve.'
#| fig-height: 4
#| fig-width: 7.5

crit <- tibble(n_total = seq(20, 400, by = 2)) |>
  mutate(n_per = n_total / 2,
         t_crit = qt(0.975, df = n_total - 2),
         d_crit = t_crit * sqrt(2 / n_per),
         g_crit = d_crit * (1 - 3 / (4 * (n_total - 2) - 1)),
         r_crit = sqrt(t_crit^2 / (t_crit^2 + n_total - 2)))

# the two panels hold different strategy sets (condition switching vs re-pairing), so
# map colour to the strategy FAMILY: identical levels in both panels means patchwork can
# collect a single legend, and the shared naming makes the cross-setting parallel explicit
family_levels <- c("Replacement", "Label switching / re-pairing", "Fabrication", "Dropping", "Duplication")
bias_fam <- bias |>
  mutate(family = fct_relevel(case_match(method,
                                         c("Condition switching", "Re-pairing") ~ "Label switching / re-pairing",
                                         .default = method),
                              family_levels))

p_g <- ggplot(bias_fam |> filter(setting == "Mean difference"),
              aes(n_total, es_median, color = family)) +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "grey50") +
  geom_line(data = crit, aes(n_total, g_crit), inherit.aes = FALSE, linetype = "dashed", colour = "grey30") +
  geom_point(position = position_dodge(width = 27), size = 1) +
  geom_linerange(aes(ymin = es_pi_lower, ymax = es_pi_upper), position = position_dodge(width = 27)) +
  labs(x = "Total sample size (N)", y = "Hedges' g in final dataset", color = NULL) +
  scale_x_continuous(breaks = c(20, 100, 200, 300, 400)) +
  scale_color_viridis_d(begin = 0, end = 0.85) +
  theme_linedraw() +
  theme(panel.grid.minor = element_blank())

p_r <- ggplot(bias_fam |> filter(setting == "Correlation"),
              aes(n_total, es_median, color = family)) +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "grey50") +
  geom_line(data = crit, aes(n_total, r_crit), inherit.aes = FALSE, linetype = "dashed", colour = "grey30") +
  geom_point(position = position_dodge(width = 27), size = 1) +
  geom_linerange(aes(ymin = es_pi_lower, ymax = es_pi_upper), position = position_dodge(width = 27)) +
  labs(x = "Total sample size (N)", y = "Pearson's r in final dataset", color = NULL) +
  scale_x_continuous(breaks = c(20, 100, 200, 300, 400)) +
  scale_color_viridis_d(begin = 0, end = 0.85) +
  theme_linedraw() +
  theme(panel.grid.minor = element_blank())

p_g + p_r + plot_layout(ncol = 2, guides = "collect") &
  theme(legend.position = "bottom") &
  guides(color = guide_legend(nrow = 2, byrow = TRUE))
```

Because the true effect is zero, the signed effect size of each final dataset is itself the bias the tampering introduced. @fig-bias shows that the median manufactured effect tracks the critical effect size for the sample size at hand: the greedy procedure stops the moment significance is crossed, so it produces effects just large enough to clear the threshold and no larger. This has an uncomfortable implication for detection: manufactured effects of this kind are, by construction, of exactly the magnitude that a reader would consider unremarkable for the sample size at hand — they look like ordinary, modestly powered findings, and they shrink as N grows just as a real literature's small-study effects are expected to.

# Discussion

## Summary of findings against the aims

The distance from a null dataset to a statistically significant one is short and cheap. The median effort is a small percentage of the sample across strategies and settings, and for most strategies a minority of datasets require more than 10% of the sample to be tampered. Where tampering fails outright it is because it is structurally impossible rather than merely expensive — the rare-outcome 2×2 cells, where too few hostile cases exist to move the table at all. The proportion of the sample that must be altered does not grow with sample size, so the intuition that large studies are safe from this kind of manipulation is wrong in proportional terms. The strategies that preserve univariate distributions — condition switching and re-pairing, precisely the ones invisible to univariate forensic checks — are among the cheapest. And the effects manufactured by a stop-at-first-crossing procedure are calibrated by construction to look unremarkable.

## Ethics and disclosure

This article documents weaponisable knowledge, and the framing matters: the simulations are designed to make a continuum visible, not to provide a recipe. The manipulations simulated here have been publicly described for years [@chambers_seven_2017; @brown_condition_2020]; what we add is quantification, and the quantification is the argument. Excluding inconvenient participants is usually filed under *p*-hacking even when the exclusions serve only to reach significance; typing the other condition's label into those same rows is fabrication. The simulations show that these differ by keystrokes, not by effort or by the damage done to the evidence — deleting a hostile case and inventing a favourable one are, in effort terms, the same act. Treating the first as common [@john_measuring_2012] and the second as rare [@fanelli_how_2009] therefore risks badly underestimating the prevalence of the latter, since the behavioural barrier between them is trivially small.

## Limitations

- **Proof of principle, not a fraud model.** The strategies are deliberately simple and greedy. Real, detection-aware fraud would trade efficiency for plausibility; conversely, several naive strategies here leave detectable signatures (truncated distributions after dropping; leverage outliers after fabrication; shifted marginal prevalence in the 2×2). The distribution-preserving strategies are the existence proof that effective *and* invisible tampering is cheap.
- **No detectability axis.** We measure effort and bias, not the probability that each manipulation would be caught by modern forensic tools [@brown_grim_2017; @heathers_recovering_2018; @wilkinson_inspect-sr_2025]. Crossing effort with detectability is the most valuable extension.
- **Greedy myopia is part of the procedure.** One-step greedy occasionally stalls — most notably when the initial dataset shows a strong effect opposite to the tampering direction, where a two-sided test's $p \approx 1$ valley cannot be crossed one step at a time. These rare failures are recorded as never attaining significance and are a true property of naive tampering, not an artefact; a less myopic tamperer would do better, so our effort estimates are, if anything, conservative upper bounds.
- **Elementary designs and Normal/binary DGPs.** Likert-type outcomes, non-Normal marginals, covariate-adjusted models, and an "alter existing values" strategy distinct from replacement are out of scope.
- **Single threshold.** All results are conditioned on $\alpha$ = .05, two-sided. No tampering budget is imposed in the design, so results under any budget are recoverable from the recorded counts.

# Data and code availability {.unnumbered}

All simulation code, the full per-strategy research compendia (with extended ADEMP specifications, per-condition tables, and Monte Carlo standard errors for every estimate), and the code that renders this manuscript — every number, figure, and table in it is computed from the simulation results at render time — are available at <https://github.com/ianhussey/simulation-impact-of-data-tampering>.

# References {.unnumbered}

::: {#refs}
:::
