---
title: "Little big lies: A compendium and simulation of data tampering and fabrication strategies"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

# TODO

- convert condition swaps to N rather than N pairs for consistency?
- correlation DIFFERENT summary and plot PLOT FUNCTIONs NEEDED? DOES IT DOUBLE N?

# Overview 

## For types of tampering and fabrication

- unwarranted dropping of cases, by N or proportion [done]
- swapping conditions, by N or proportion [done]
- swapping entire conditions, so that significant results in the wrong direction are simply relabeled as in the other direction. (this will increase false positives by half alpha - no need to simulate)
- changing/offsetting existing cases - changing sign [done] its a bit arbitrary but could illustrate the point
- duplicating existing cases 
- duplicate cases from the top 10% of values?
- fabricating new cases [done]
- alter? same strategy as fabricate but change existing data points
- duplicate all these sims with likert data (skewed, bounded, granular)

Some mention of Spiess et al. (2024) and dfstat seems warranted, even if only to say that many published articles results are driven by a very small number of data points and the effect disappears if they're removed. 

early stage data processing is important

Uri's project?

## Ethics and optics 

Think about the ethics and options of distributing this weaponisable knowledge. Are there ways to hide detectability in it? 

Add explicit discussion about how arbitrary the line is between p-hacking (scientifically naughty) and fabrication (legally actionable): 

- excluding some participants is usually seen as p-hacking, even if these exclusions are totally unjustifiable and unprincipled, and seem to just be done to produce significant p values.
- but if instead of clicking the delete key on those rows you instead type the other condition name in, suddenly its fabrication and you could lose your job over it.
- the idea that one of these is common and one of these is rare, or one of these is merely iffy and the other is Clearly Unacceptable, risks underestimating the prevalence of the latter.
- as usual, this boils down to a Trolly Problem: is deleting data really ethically different from making up data, if both serve the individuals' self-advancement at the expense of undermining the integrity of the research and the validity of the evidence? 

## Realism of sims

- Perhaps the sim that drops participants feels too obvious for a fraudster, because it can create an obvious truncation of otherwise normally distributed scores. However: 1) many detected cases of fraud have not been very elaborate or smart, and 2) this simulation is a proof of principle, there are of course more elaborate strategies possible which would seek to try to retain the plausibility of data (e.g., removing data beyond a threshold probabilistically).

Alway comes back to illustration of minimal tampering and effort. Anything can be found with arbitrary fabrication.

# Dependencies

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

```{r}

library(tidyr)
library(dplyr)
library(readr)
library(forcats)
library(tibble)
library(purrr) 
library(furrr)
library(MASS)
library(ggplot2)
library(scales)
library(knitr)
library(kableExtra)
library(janitor)
library(ggtext)
library(patchwork)
library(wesanderson)

# set up parallel processing
plan(multisession, workers = availableCores())

options(scipen=999)

dir.create("../data")
dir.create("../communications/plots")

```

# Common functions and global variables

```{r}

# n iterations for all sims
n_iterations <- 2000L

# define data generating function ----
generate_data <- function(n_per_condition,
                          mean_control,
                          mean_intervention,
                          sd_control,
                          sd_intervention) {
  
  data_control <- 
    tibble(condition = "control",
           score = rnorm(n = n_per_condition, mean = mean_control, sd = sd_control))
  
  data_intervention <- 
    tibble(condition = "intervention",
           score = rnorm(n = n_per_condition, mean = mean_intervention, sd = sd_intervention))
  
  data <- bind_rows(data_control,
                    data_intervention) 
  
  return(data)
}

# define data analysis function ----
# slow, because t.test is slow. 
analyse_data <- function(data) {
  
  res_t_test <- t.test(formula = score ~ condition,
                       data = data,
                       var.equal = TRUE,
                       alternative = "two.sided")
  
  res <- tibble(p = res_t_test$p.value)
  
  return(res)
}

# # manual t-test p-value calculation function
# analyse_data <- function(data) {
#   x <- data$score[data$condition == "control"]
#   y <- data$score[data$condition == "intervention"]
#   
#   # Basic checks like t.test() would do
#   if (length(x) < 2 || length(y) < 2 || all(is.na(x)) || all(is.na(y))) {
#     return(tibble(p = NA_real_))
#   }
#   
#   if (sd(x, na.rm = TRUE) == 0 || sd(y, na.rm = TRUE) == 0) {
#     return(tibble(p = NA_real_))
#   }
# 
#   # Remove NAs for safety
#   x <- x[!is.na(x)]
#   y <- y[!is.na(y)]
#   
#   n1 <- length(x)
#   n2 <- length(y)
#   m1 <- mean(x)
#   m2 <- mean(y)
#   s1 <- sd(x)
#   s2 <- sd(y)
# 
#   s_pooled <- sqrt(((n1 - 1) * s1^2 + (n2 - 1) * s2^2) / (n1 + n2 - 2))
#   
#   # Catch divide-by-zero or NaN
#   if (is.nan(s_pooled) || s_pooled == 0) {
#     return(tibble(p = NA_real_))
#   }
# 
#   t_stat <- (m1 - m2) / (s_pooled * sqrt(1/n1 + 1/n2))
#   df <- n1 + n2 - 2
#   p_val <- 2 * pt(-abs(t_stat), df)
# 
#   if (is.na(p_val)) p_val <- NA_real_
# 
#   return(tibble(p = p_val))
# }

summarise_across_iterations <- function(data){
  data |>
    group_by(n_per_condition) |>
    summarize(proportion_never_sig = mean(tampers_needed == floor(0.5 * n_per_condition * 2)),
              min = min(tampers_needed[tampers_needed < floor(0.5 * n_per_condition * 2)]),
              max = max(tampers_needed[tampers_needed < floor(0.5 * n_per_condition * 2)]),
              mean = mean(tampers_needed[tampers_needed < floor(0.5 * n_per_condition * 2)]),
              sd = sd(tampers_needed[tampers_needed < floor(0.5 * n_per_condition * 2)]),
              median = median(tampers_needed[tampers_needed < floor(0.5 * n_per_condition * 2)]),
              mad = mad(tampers_needed[tampers_needed < floor(0.5 * n_per_condition * 2)]),
              lower = quantile(tampers_needed[tampers_needed < floor(0.5 * n_per_condition * 2)], probs = 0.25),
              upper = quantile(tampers_needed[tampers_needed < floor(0.5 * n_per_condition * 2)], probs = 0.75)) |>
    ungroup() |>
    mutate(n_total = n_per_condition*2,
           prop_min = min/n_total,
           prop_max = max/n_total,
           prop_mean = mean/n_total,
           prop_sd = sd/n_total,
           prop_median = median/n_total,
           prop_mad = median/n_total,
           prop_lower = lower / n_total,
           prop_upper = upper / n_total) |>
    relocate(n_total, .before = n_per_condition) |>
    dplyr::select(-n_per_condition)
}

plot_distributions <- function(data, title, x_axis){
  ggplot(data, aes(x = tampers_needed)) +
    geom_histogram(binwidth = 1) +
    labs(title = title, 
         x = x_axis,
         y = "Count") +
    ylim(0, NA) +
    theme_linedraw() +
    scale_x_continuous(breaks = breaks_pretty(n = 5)) +
    facet_wrap(~ n_per_condition) # N PER CONDITION NOT TOTAL N
}

plot_tampers_n <- function(data, title, y_axis){
  ggplot(data, aes(n_total, median)) +
    geom_point() +
    geom_linerange(aes(ymin = lower, ymax = upper)) +
    labs(title = title,
         x = "Sample size",
         y = y_axis) +
    theme_linedraw() +
    scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
    scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2)
}

plot_tampers_prop <- function(data, title, y_axis){
  ggplot(data, aes(n_total, prop_median)) +
    geom_point() +
    geom_linerange(aes(ymin = prop_lower, ymax = prop_upper)) +
    labs(title = title,
         x = "Sample size",
         y = y_axis) +
    theme_linedraw() +
    scale_y_continuous(labels = percent_format(), limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
    scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2)
}

```

# Unwarranted data exclusion (dropping N)

## t-test

### Simulation functions

```{r}

iterative_exclusion_once <- function(data) {
  
  # we only consider rows that have not yet been dropped (i.e., score not NA)
  data_incl <- data |> filter(!is.na(score))
  
  # if either group has fewer than 2 participants, you can decide how to handle
  # (for example, skip dropping or forcibly return data unchanged).
  # For simplicity, let's handle the case but proceed only if both groups still have variance:
  if(any(table(data_incl$condition) < 2)){
    # not enough participants to meaningfully compute z-scores, so just return the same data
    return(data)
  }
  
  # compute z-scores for intervention
  data_incl <- data_incl |>
    group_by(condition) |>
    mutate(
      z_score_group = (score - mean(score)) / sd(score)
    ) |>
    ungroup()
  
  # smallest in intervention (lowest z-score)
  smallest_in_intervention <- data_incl |>
    filter(condition == "intervention") |>
    slice_min(z_score_group, n = 1)
  
  # largest in control (highest z-score)
  largest_in_control <- data_incl |>
    filter(condition == "control") |>
    slice_max(z_score_group, n = 1)
  
  # among those two, pick the single participant with the largest absolute z-score
  to_drop <- bind_rows(smallest_in_intervention,
                       largest_in_control) |>
    slice_max(abs(z_score_group), n = 1)
  
  # apply the drop
  data_dropped <- data |>
    mutate(score = ifelse(row_number() %in% to_drop$row_number, NA, score))
  
  return(data_dropped)
}


run_iterative_exclusion_study <- function(n_per_condition,
                                          mean_control,
                                          mean_intervention,
                                          sd_control,
                                          sd_intervention,
                                          max_drop_proportion = 0.5) {
  
  # 1) Generate data
  data <- generate_data(n_per_condition = n_per_condition,
                        mean_control = mean_control,
                        mean_intervention = mean_intervention,
                        sd_control = sd_control,
                        sd_intervention = sd_intervention)
  
  # Store row_number so we can track drops more easily
  data <- data |> mutate(row_number = row_number())
  
  # 2) Analyze once
  test_res <- analyse_data(data)
  p_val <- test_res$p[1]
  
  # If significant from the start, return 0
  if(p_val < 0.05) {
    return(0)
  }
  
  # otherwise, attempt iterative exclusion
  n_total <- nrow(data)
  max_drops <- floor(max_drop_proportion * n_total)
  drops_so_far <- 0
  
  # keep dropping while p >= 0.05, if we haven't hit max
  while(p_val >= 0.05 && drops_so_far < max_drops) {
    # drop exactly one new participant
    data <- iterative_exclusion_once(data)
    drops_so_far <- drops_so_far + 1
    
    # re-test
    test_res <- analyse_data(data)
    p_val <- test_res$p[1]
  }
  
  # after the loop, either we reached significance or we used up 50% of the data
  if(p_val < 0.05) {
    return(drops_so_far)
  } else {
    # indicate never reached significance. Return the total drops, or you could return NA/Inf.
    return(drops_so_far)
  }
}

```

### Run simulation

```{r}

if(file.exists("../data/simulation_iterative_exclusions_meandiff.rds")){
  simulation_iterative_exclusions_meandiff <- read_rds("../data/simulation_iterative_exclusions_meandiff.rds")
} else {
  
  # Create a parameter grid
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    iteration = 1:n_iterations
  )
  
  set.seed(123)
  
  simulation_iterative_exclusions_meandiff <- experiment_parameters_grid |>
    mutate(tampers_needed = future_pmap_dbl(list(n_per_condition, 
                                                 mean_control, 
                                                 mean_intervention, 
                                                 sd_control, 
                                                 sd_intervention),
                                            run_iterative_exclusion_study,
                                            max_drop_proportion = 0.5, 
                                            .progress = TRUE,
                                            .options = furrr_options(seed = TRUE)))
  
  write_rds(simulation_iterative_exclusions_meandiff, "../data/simulation_iterative_exclusions_meandiff.rds")
}

# Summarize results
simulation_iterative_exclusions_meandiff_summary <- summarise_across_iterations(simulation_iterative_exclusions_meandiff) 

```

### Table

```{r}

simulation_iterative_exclusions_meandiff_summary |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

### Plot

```{r}

plot_distributions(simulation_iterative_exclusions_meandiff, "Dropping", "N drops")

p_iterative_exclusions_meandiff_n <- plot_tampers_n(simulation_iterative_exclusions_meandiff_summary, "Dropping", "N drops")

p_iterative_exclusions_meandiff_prop <- plot_tampers_prop(simulation_iterative_exclusions_meandiff_summary, "Dropping", "N drops")

```

## Correlation

### Simulation functions

```{r eval=FALSE, include=FALSE}

# Generate bivariate normal data with no correlation
generate_uncorrelated_data <- function(n) {
  mu <- c(0, 0)
  sigma <- matrix(c(1, 0, 0, 1), nrow = 2)  # Identity matrix: no correlation
  data <- as_tibble(mvrnorm(n, mu = mu, Sigma = sigma))
  names(data) <- c("x", "y")
  data$row_id <- 1:n
  return(data)
}

# Analyze correlation
analyse_correlation <- function(data) {
  cor.test(data$x, data$y)$p.value
}

# Iteratively delete points that dampen a positive correlation
iterative_drop_once <- function(data) {
  data <- data |> filter(!is.na(x) & !is.na(y))
  
  # Identify outliers in the top-left or bottom-right
  data <- data |>
    mutate(
      score = -1 * x * y,  # large negative x*y = top-left or bottom-right
      abs_score = abs(score)
    )
  
  to_drop <- data |> slice_max(score, n = 1)
  data <- data |>
    mutate(
      x = ifelse(row_id == to_drop$row_id, NA, x),
      y = ifelse(row_id == to_drop$row_id, NA, y)
    )
  
  return(data)
}

# Run one full trial
run_iterative_correlation_exclusion <- function(n, max_drop_proportion = 0.5) {
  data <- generate_uncorrelated_data(n)
  p <- analyse_correlation(data)
  
  if (p < 0.05) return(0)  # already a false positive
  
  max_drops <- floor(max_drop_proportion * n)
  drops <- 0
  
  while (p >= 0.05 && drops < max_drops) {
    data <- iterative_drop_once(data)
    drops <- drops + 1
    p <- analyse_correlation(data |> filter(!is.na(x) & !is.na(y)))
  }
  
  if (p < 0.05) return(drops) else return(drops)
}

```

### Run simulation

```{r eval=FALSE, include=FALSE}

# Run simulations
if(file.exists("../data/simulation_iterative_exclusion_correlation.rds")){
  simulation_iterative_exclusion_correlation <- read_rds("../data/simulation_iterative_exclusion_correlation.rds")
} else {
  
  # Parameter grid
  experiment_grid <- expand_grid(
    n = c(10*2, seq(from = 25*2, to = 200*2, by = 25*2)), # to match total sample size of other sims
    iteration = 1:n_iterations
  )
  
  set.seed(123)
  
  simulation_iterative_exclusion_correlation <- experiment_grid |>
    mutate(tampers_needed = future_map2_dbl(
      n, iteration,
      ~run_iterative_correlation_exclusion(.x),
      .options = furrr_options(seed = TRUE),
      .progress = TRUE
    ))
  
  write_rds(simulation_iterative_exclusion_correlation, "../data/simulation_iterative_exclusion_correlation.rds")
  
}

# summarize
simulation_iterative_exclusion_correlation_summary <- summarise_across_iterations(simulation_iterative_exclusion_correlation)

```

### Table

```{r eval=FALSE, include=FALSE}

simulation_iterative_exclusion_correlation_summary |> 
  mutate(across(where(is.numeric), ~ janitor::round_half_up(., digits = 2))) |>
  knitr::kable() |>
  kableExtra::kable_classic(full_width = FALSE)

```

### Plot

```{r eval=FALSE, include=FALSE}

plot_distributions(simulation_iterative_exclusion_correlation, "Dropping", "N drops")

p_iterative_exclusion_meandiff_n <- plot_tampers_n(simulation_iterative_exclusion_correlation_summary, "Dropping", "N drops")

p_iterative_exclusion_meandiff_prop <- plot_tampers_prop(simulation_iterative_exclusion_correlation_summary, "Dropping", "N drops")

```

# Condition swapping 

## t-test

### Simulation functions

```{r}

# Swap condition of most extreme participant from each group
iterative_condition_swap_once <- function(data) {
  
  data <- data |> 
    group_by(condition) |> 
    mutate(z_score = (score - mean(score)) / sd(score)) |> 
    ungroup()
  
  # Identify one participant in each group to switch
  to_swap_intervention <- data |> filter(condition == "intervention") |> slice_min(z_score, n = 1)
  to_swap_control      <- data |> filter(condition == "control") |> slice_max(z_score, n = 1)
  
  if(nrow(to_swap_intervention) < 1 || nrow(to_swap_control) < 1) {
    return(data)
  }
  
  rows_to_swap <- c(to_swap_intervention$row_number, to_swap_control$row_number)
  
  data <- data |> 
    mutate(condition = case_when(
      row_number %in% rows_to_swap & condition == "intervention" ~ "control",
      row_number %in% rows_to_swap & condition == "control" ~ "intervention",
      TRUE ~ condition
    )) |> 
    dplyr::select(-z_score)
  
  return(data)
}

# Main simulation function
run_iterative_condition_swap_study <- function(n_per_condition,
                                               mean_control,
                                               mean_intervention,
                                               sd_control,
                                               sd_intervention,
                                               max_swap_proportion = 0.5) {
  
  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) |> 
    mutate(row_number = row_number())
  
  res <- analyse_data(data)
  p_val <- res$p[1]
  
  if(p_val < 0.05) {
    return(0)
  }
  
  n_total <- nrow(data)
  max_swaps <- floor(max_swap_proportion * n_total)
  swaps_so_far <- 0
  
  while(p_val >= 0.05 && swaps_so_far < max_swaps) {
    data <- iterative_condition_swap_once(data)
    swaps_so_far <- swaps_so_far + 2
    res <- analyse_data(data)
    p_val <- res$p[1]
  }
  
  return(swaps_so_far)
}

```

### Run simulation

```{r}

# Run simulation or load existing
if(file.exists("../data/simulation_iterative_swaps_meandiff.rds")) {
  simulation_iterative_swaps_meandiff <- read_rds("../data/simulation_iterative_swaps_meandiff.rds")
} else {
  
  # Parameter grid
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    iteration = 1:n_iterations
  )
  
  set.seed(123)
  
  simulation_iterative_swaps_meandiff <- experiment_parameters_grid |> 
    mutate(tampers_needed = future_pmap_dbl(list(n_per_condition, 
                                                 mean_control, 
                                                 mean_intervention, 
                                                 sd_control, 
                                                 sd_intervention),
                                            run_iterative_condition_swap_study,
                                            max_swap_proportion = 0.5, 
                                            .progress = TRUE,
                                            .options = furrr_options(seed = TRUE)))
  
  write_rds(simulation_iterative_swaps_meandiff, "../data/simulation_iterative_swaps_meandiff.rds")
}

# summarize
simulation_iterative_swaps_meandiff_summary <- summarise_across_iterations(simulation_iterative_swaps_meandiff)

```

### Table

```{r}

simulation_iterative_swaps_meandiff_summary |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

### Plot

```{r}

plot_distributions(simulation_iterative_swaps_meandiff, "Swaps", "N condition swaps")

p_iterative_swaps_meandiff_n <- plot_tampers_n(simulation_iterative_swaps_meandiff_summary, "Swaps", "N condition swaps")

p_iterative_swaps_meandiff_prop <- plot_tampers_prop(simulation_iterative_swaps_meandiff_summary, "Swaps", "N condition swaps")

```

# Duplicate data points

From top 10% most favorable data points

## t-test

### Simulation functions

```{r}

# Function to perform one round of top-10%-based duplication
duplicate_from_top_10_percent_once <- function(data, original_top10_control, original_top10_intervention) {
  
  # Sample one from each condition's original top 10%
  dup_control <- sample_n(original_top10_control, 1)
  dup_intervention <- sample_n(original_top10_intervention, 1)
  
  # Append to data
  data <- bind_rows(data, dup_control, dup_intervention)
  
  return(data)
}

# Main simulation function
run_iterative_duplication_study <- function(n_per_condition,
                                            mean_control,
                                            mean_intervention,
                                            sd_control,
                                            sd_intervention,
                                            max_duplication_proportion = 0.5,
                                            proportion) {
  
  # Generate original data and label with row IDs
  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) %>%
    mutate(row_id = row_number())
  
  # Compute original top 10% favorable scores per condition
  top10_control <- data %>% 
    filter(condition == "control") %>% 
    slice_max(score, prop = proportion)
  
  top10_intervention <- data %>% 
    filter(condition == "intervention") %>% 
    slice_min(score, prop = proportion)
  
  # Initial p-value
  p_val <- analyse_data(data)$p[1]
  if (p_val < 0.05) return(0)
  
  n_total <- nrow(data)
  max_dups <- floor(max_duplication_proportion * n_total)
  dups_so_far <- 0
  
  while (p_val >= 0.05 && dups_so_far < max_dups) {
    data <- duplicate_from_top_10_percent_once(data, top10_control, top10_intervention)
    p_val <- analyse_data(data)$p[1]
    dups_so_far <- dups_so_far + 2
  }
  
  return(dups_so_far)
}

```

### Run simulation

```{r}

if (file.exists("../data/simulation_iterative_duplication_meandiff.rds")) {
  simulation_iterative_duplication_meandiff <- read_rds("../data/simulation_iterative_duplication_meandiff.rds")
} else {
  
  # Parameter grid
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    proportion = 0.1,
    iteration = 1:n_iterations
  )
  
  set.seed(123)
  
  simulation_iterative_duplication_meandiff <- experiment_parameters_grid %>%
    mutate(tampers_needed = future_pmap_dbl(list(n_per_condition,
                                                 mean_control,
                                                 mean_intervention,
                                                 sd_control,
                                                 sd_intervention,
                                                 proportion),
                                            run_iterative_duplication_study,
                                            max_duplication_proportion = 0.5,
                                            .progress = TRUE,
                                            .options = furrr_options(seed = TRUE)))
  
  write_rds(simulation_iterative_duplication_meandiff, "../data/simulation_iterative_duplication_meandiff.rds")
}

# summarise
simulation_iterative_duplication_meandiff_summary <- summarise_across_iterations(simulation_iterative_duplication_meandiff) 

```

### Table

```{r}

simulation_iterative_duplication_meandiff_summary |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

### Plot

```{r}

plot_distributions(simulation_iterative_duplication_meandiff, "Duplications", "N duplications")

p_iterative_duplication_meandiff_n <- plot_tampers_n(simulation_iterative_duplication_meandiff_summary, "Duplications", "N duplications")

p_iterative_duplication_meandiff_prop <- plot_tampers_prop(simulation_iterative_duplication_meandiff_summary, "Duplications", "N duplications")

```

## Correlation

TODO

# Change sign (%)

## t-test

Very long run time compared to the others?

### Simulation functions

Optimised function avoids data set duplication and use of t.test which is slow - might want to use the original for a publication, even if slow, for the sake of standardization

```{r}

# slow
# iterative_sign_switch_once <- function(data) {
#   # Compute z-scores across the full sample
#   data <- data %>%
#     mutate(z_score = (score - mean(score)) / sd(score),
#            flipped_score = -score)
# 
#   # Try flipping each participant, and check resulting p-value
#   candidate_results <- map_dfr(1:nrow(data), function(i) {
#     test_data <- data
#     test_data$score[i] <- test_data$flipped_score[i]
# 
#     p <- t.test(score ~ condition, data = test_data, var.equal = TRUE)$p.value
# 
#     tibble(row_id = test_data$row_id[i], new_p = p)
#   })
# 
#   # Identify the row that gives the lowest p-value
#   best_candidate <- candidate_results %>%
#     slice_min(new_p, n = 1)
# 
#   # Apply the flip
#   data$score[data$row_id == best_candidate$row_id] <- -data$score[data$row_id == best_candidate$row_id]
# 
#   return(data %>% dplyr::select(-z_score, -flipped_score))
# }

# optimised
iterative_sign_switch_once <- function(data) {
  # Precompute group info
  condition <- data$condition
  group_A <- condition == unique(condition)[1]
  group_B <- !group_A
  score <- data$score
  
  # Precompute group sizes
  n_A <- sum(group_A)
  n_B <- sum(group_B)
  
  best_p <- 1
  best_i <- NA
  
  # Try flipping each participant and compute t stat manually
  for (i in seq_along(score)) {
    score_flipped <- score
    score_flipped[i] <- -score[i]
    
    m_A <- mean(score_flipped[group_A])
    m_B <- mean(score_flipped[group_B])
    s_A <- sd(score_flipped[group_A])
    s_B <- sd(score_flipped[group_B])
    
    # Pooled SD
    s_pooled <- sqrt(((n_A - 1) * s_A^2 + (n_B - 1) * s_B^2) / (n_A + n_B - 2))
    t_stat <- (m_A - m_B) / (s_pooled * sqrt(1 / n_A + 1 / n_B))
    df <- n_A + n_B - 2
    p_val <- 2 * pt(-abs(t_stat), df)
    
    if (p_val < best_p) {
      best_p <- p_val
      best_i <- i
    }
  }
  
  # Apply the flip only to the best row
  data$score[best_i] <- -data$score[best_i]
  
  return(data)
}


run_iterative_sign_switch_study <- function(n_per_condition,
                                            mean_control,
                                            mean_intervention,
                                            sd_control,
                                            sd_intervention,
                                            max_flips_proportion = 0.5) {
  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) %>%
    mutate(row_id = row_number())
  
  p_val <- t.test(score ~ condition, data = data, var.equal = TRUE)$p.value
  if (p_val < 0.05) return(0)
  
  n_total <- nrow(data)
  max_flips <- floor(max_flips_proportion * n_total)
  flips <- 0
  
  while (p_val >= 0.05 && flips < max_flips) {
    data <- iterative_sign_switch_once(data)
    p_val <- t.test(score ~ condition, data = data, var.equal = TRUE)$p.value
    flips <- flips + 1
  }
  
  return(flips)
}

```

### Run simulation

```{r}

if (file.exists("../data/simulation_iterative_changesigns_meandiff.rds")) {
  simulation_iterative_changesigns_meandiff <- read_rds("../data/simulation_iterative_changesigns_meandiff.rds")
} else {
  
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    iteration = 1:n_iterations
  )
  
  set.seed(123)
  
  simulation_iterative_changesigns_meandiff <- experiment_parameters_grid %>%
    mutate(tampers_needed = future_pmap_dbl(list(n_per_condition,
                                                 mean_control,
                                                 mean_intervention,
                                                 sd_control,
                                                 sd_intervention),
                                            run_iterative_sign_switch_study,
                                            max_flips_proportion = 0.5,
                                            .progress = TRUE,
                                            .options = furrr_options(seed = TRUE)))
  
  write_rds(simulation_iterative_changesigns_meandiff, "../data/simulation_iterative_changesigns_meandiff.rds")
}

# summarise
simulation_iterative_changesigns_meandiff_summary <- summarise_across_iterations(simulation_iterative_changesigns_meandiff)

```

### Table

```{r}

simulation_iterative_changesigns_meandiff_summary |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

### Plot

```{r}

plot_distributions(simulation_iterative_changesigns_meandiff, "Sign changing", "N sign changed")

p_iterative_changesigns_meandiff_n <- plot_tampers_n(simulation_iterative_changesigns_meandiff_summary, "Sign changing", "N sign changed")

p_iterative_changesigns_meandiff_prop <- plot_tampers_prop(simulation_iterative_changesigns_meandiff_summary, "Sign changing", "N sign changed")

```

## Correlation

TODO

# Fabrication

## Illustrate 

for the correlation sim, m = .93 so that cohen's d = 5.0 == pearson's r = .93

```{r}

x_vals <- seq(-7, 7, length.out = 1000)

df <- tibble(
  x = rep(x_vals, 2),
  density = c(dnorm(x_vals, mean = 0, sd = 1),
              dnorm(x_vals, mean = 5, sd = 0.2)),
  distribution = rep(c("N(0,1)", "N(1,0.2)"), each = length(x_vals))
)

ggplot(df, aes(x = x, y = density, color = distribution)) +
  geom_line(linewidth = 1) +
  labs(x = "Value", y = "Density") +
  theme_linedraw()

```

## t-test

### Simulation functions

```{r}

run_iterative_fabrication_study <- function(n_per_condition,
                                            mean_control,
                                            mean_intervention,
                                            sd_control,
                                            sd_intervention,
                                            fabricated_control_mean,
                                            fabricated_control_sd,
                                            fabricated_intervention_mean,
                                            fabricated_intervention_sd,
                                            max_add_proportion = 0.5) {
  
  # Generate original data
  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) %>%
    mutate(row_id = row_number())
  
  # Initial p-value
  p_val <- analyse_data(data)$p[1]
  if (p_val < 0.05) return(0)
  
  n_total <- nrow(data)
  max_adds <- floor(max_add_proportion * n_total)
  adds_so_far <- 0
  
  while (p_val >= 0.05 && adds_so_far < max_adds) {
    # Add one control and one intervention point per iteration
    new_control <- tibble(condition = "control", score = rnorm(1, mean = fabricated_control_mean, sd = fabricated_control_sd))
    new_intervention <- tibble(condition = "intervention", score = rnorm(1, mean = fabricated_intervention_mean, sd = fabricated_intervention_sd))
    
    data <- bind_rows(data, new_control, new_intervention)
    adds_so_far <- adds_so_far + 2
    
    p_val <- analyse_data(data)$p[1]
  }
  
  return(adds_so_far)
}

```

### Run simulation

```{r}

if (file.exists("../data/simulation_iterative_fabrication_meandiff.rds")) {
  simulation_iterative_fabrication_meandiff <- read_rds("../data/simulation_iterative_fabrication_meandiff.rds")
} else {
  
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    fabricated_control_mean = 0,
    fabricated_control_sd = 0.2,
    fabricated_intervention_mean = 5,
    fabricated_intervention_sd = 0.2,
    iteration = 1:n_iterations
  )
  
  set.seed(123)
  
  simulation_iterative_fabrication_meandiff <- experiment_parameters_grid %>%
    mutate(tampers_needed = future_pmap_dbl(list(n_per_condition,
                                                 mean_control,
                                                 mean_intervention,
                                                 sd_control,
                                                 sd_intervention,
                                                 fabricated_control_mean,
                                                 fabricated_control_sd,
                                                 fabricated_intervention_mean,
                                                 fabricated_intervention_sd),
                                            run_iterative_fabrication_study,
                                            max_add_proportion = 0.5,
                                            .progress = TRUE,
                                            .options = furrr_options(seed = TRUE)))
  
  write_rds(simulation_iterative_fabrication_meandiff, "../data/simulation_iterative_fabrication_meandiff.rds")
}

# summarise
simulation_iterative_fabrication_meandiff_summary <- summarise_across_iterations(simulation_iterative_fabrication_meandiff)

```

### Table

```{r}

simulation_iterative_fabrication_meandiff_summary |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

### Plot

```{r}

plot_distributions(simulation_iterative_fabrication_meandiff, "Fabrications", "N fabricated")

p_iterative_fabrication_meandiff_n <- plot_tampers_n(simulation_iterative_fabrication_meandiff_summary, "Fabrications", "N fabricated")

p_iterative_fabrication_meandiff_prop <- plot_tampers_prop(simulation_iterative_fabrication_meandiff_summary, "Fabrications", "N fabricated")

```

## Correlation

TODO

# Replace

## t-test

### Simulation functions

```{r}

run_iterative_replacement_study <- function(n_per_condition,
                                            mean_control,
                                            mean_intervention,
                                            sd_control,
                                            sd_intervention,
                                            fabricated_control_mean,
                                            fabricated_control_sd,
                                            fabricated_intervention_mean,
                                            fabricated_intervention_sd,
                                            max_replace_proportion = 0.5) {
  
  # Generate original data
  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) %>%
    mutate(row_id = row_number())
  
  # Initial p-value
  p_val <- analyse_data(data)$p[1]
  if (p_val < 0.05) return(0)
  
  n_total <- nrow(data)
  max_replacements <- floor(max_replace_proportion * n_total)
  replacements_so_far <- 0
  
  while (!is.na(p_val) && p_val >= 0.05 && replacements_so_far < max_replacements) {
    # Filter out NA scores
    data_incl <- data |> filter(!is.na(score))
    
    if (any(table(data_incl$condition) < 2)) break
    
    data_incl <- data_incl |> 
      group_by(condition) |> 
      mutate(z_score_group = (score - mean(score)) / sd(score)) |> 
      ungroup()
    
    smallest_in_intervention <- data_incl |> 
      filter(condition == "intervention") |> 
      slice_min(z_score_group, n = 1)
    
    largest_in_control <- data_incl |> 
      filter(condition == "control") |> 
      slice_max(z_score_group, n = 1)
    
    to_replace <- bind_rows(smallest_in_intervention, largest_in_control) |> 
      slice_max(abs(z_score_group), n = 1)
    
    replacement_row <- if (to_replace$condition == "control") {
      tibble(score = rnorm(1, fabricated_control_mean, fabricated_control_sd))
    } else {
      tibble(score = rnorm(1, fabricated_intervention_mean, fabricated_intervention_sd))
    }
    
    data <- data |> mutate(score = ifelse(row_id == to_replace$row_id, replacement_row$score, score))
    
    replacements_so_far <- replacements_so_far + 1
    p_val <- analyse_data(data)$p[1]
  }
  
  return(replacements_so_far)
}

```

### Run simulation

```{r}

if (file.exists("../data/simulation_iterative_replacement_meandiff.rds")) {
  simulation_iterative_replacement_meandiff <- read_rds("../data/simulation_iterative_replacement_meandiff.rds")
} else {
  
  # Parameter grid
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    fabricated_control_mean = 0,
    fabricated_control_sd = 0.2,
    fabricated_intervention_mean = 5,
    fabricated_intervention_sd = 0.2,
    iteration = 1:n_iterations
  )
  
  set.seed(123)
  
  simulation_iterative_replacement_meandiff <- experiment_parameters_grid %>%
    mutate(tampers_needed = future_pmap_dbl(list(n_per_condition,
                                                 mean_control,
                                                 mean_intervention,
                                                 sd_control,
                                                 sd_intervention,
                                                 fabricated_control_mean,
                                                 fabricated_control_sd,
                                                 fabricated_intervention_mean,
                                                 fabricated_intervention_sd),
                                            run_iterative_replacement_study,
                                            max_replace_proportion = 0.5,
                                            .progress = TRUE,
                                            .options = furrr_options(seed = TRUE)))
  
  write_rds(simulation_iterative_replacement_meandiff, "../data/simulation_iterative_replacement_meandiff.rds")
}

# Summarize results
simulation_iterative_replacement_meandiff_summary <- summarise_across_iterations(simulation_iterative_replacement_meandiff)

```

### Table

```{r}

simulation_iterative_replacement_meandiff_summary |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

### Plot

```{r}

plot_distributions(simulation_iterative_replacement_meandiff, "Replacing", "N replaced")

p_iterative_replacement_meandiff_n <- plot_tampers_n(simulation_iterative_replacement_meandiff_summary, "Replacment", "N replaced")

p_iterative_replacement_meandiff_prop <- plot_tampers_prop(simulation_iterative_replacement_meandiff_summary, "Replacment", "N replaced")

```

## Correlation

TODO

# Combined tables and plots

## t-tests

### Faceted plots

```{r fig.height=15, fig.width=11}

p_iterative_exclusions_meandiff_n + 
  p_iterative_exclusions_meandiff_prop + 
  p_iterative_swaps_meandiff_n +
  p_iterative_swaps_meandiff_prop +
  p_iterative_duplication_meandiff_n +
  p_iterative_duplication_meandiff_prop +
  p_iterative_changesigns_meandiff_n +
  p_iterative_changesigns_meandiff_prop +
  p_iterative_fabrication_meandiff_n +
  p_iterative_fabrication_meandiff_prop +
  p_iterative_replacement_meandiff_n + 
  p_iterative_replacement_meandiff_prop +
  plot_layout(ncol = 2)

```

### Single plot

```{r fig.height=8, fig.width=10}

simulation_iterative_meandiff_summary <- bind_rows(
  simulation_iterative_exclusions_meandiff_summary |> mutate(Method = "Dropping"),
  simulation_iterative_swaps_meandiff_summary |> mutate(Method = "Condition swapping"),
  simulation_iterative_duplication_meandiff_summary |> mutate(Method = "Duplication"),
  #simulation_iterative_changesigns_meandiff_summary |> mutate(Method = "Sign changing"),
  simulation_iterative_fabrication_meandiff_summary |> mutate(Method = "Fabrication"),
  simulation_iterative_replacement_meandiff_summary |> mutate(Method = "Replacement")
) |>
  mutate(Method = fct_relevel(Method, "Duplication", "Fabrication", "Dropping", "Replacement", "Condition swapping", "Sign changing"))

p_meandiff_n <- 
  ggplot(simulation_iterative_meandiff_summary, 
         aes(n_total, median, color = Method, fill = Method)) +
  geom_point(position = position_dodge(width = 27)) +
  geom_linerange(aes(ymin = lower, ymax = upper), position = position_dodge(width = 27)) +
  labs(title = "Count of tampered cases needed to reach statistical significance",
       subtitle = "White gaps represent medians, colored bars represent interquartile range, whiskers represent range",
       x = "Sample size",
       y = "Count") +
  theme_linedraw() +
  scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
  scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2) +
  #scale_color_viridis_d(begin = 0.2, end = 0.8, option = "viridis") +
  #scale_fill_viridis_d(begin = 0.2, end = 0.8, option = "viridis") 
  scale_color_manual(values = wes_palette("Cavalcanti1")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1")) 

p_meandiff_prop <- 
  ggplot(simulation_iterative_meandiff_summary, 
         aes(n_total, prop_median, color = Method, fill = Method)) +
  geom_point(position = position_dodge(width = 27)) +
  geom_linerange(aes(ymin = prop_lower, ymax = prop_upper), position = position_dodge(width = 27)) +
  labs(title = "Proportion of tampered cases needed to reach statistical significance",
       subtitle = "White gaps represent medians, colored bars represent interquartile range, whiskers represent range",
       x = "Sample size",
       y = "Proportion") +
  theme_linedraw() +
  scale_y_continuous(labels = percent_format(), limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
  scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2) +
  #scale_color_viridis_d(begin = 0.2, end = 0.8, option = "viridis") +
  #scale_fill_viridis_d(begin = 0.2, end = 0.8, option = "viridis") 
  scale_color_manual(values = wes_palette("Cavalcanti1")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1")) 

p_meandiff_n +
  p_meandiff_prop +
  plot_layout(ncol = 1, guides = "collect")

ggsave(filename = "../communications/plots/iterative_meandiff_combined.pdf",
       device = "pdf",
       width = 9,
       height = 8)

ggsave(filename = "../communications/plots/iterative_meandiff_combined.png",
       device = "png",
       width = 9,
       height = 8)

```

### Single boxplot

```{r fig.height=8, fig.width=10}

simulation_iterative_meandiff <- bind_rows(
  simulation_iterative_exclusions_meandiff |> mutate(Method = "Dropping"),
  simulation_iterative_swaps_meandiff |> mutate(Method = "Condition swapping"),
  simulation_iterative_duplication_meandiff |> mutate(Method = "Duplication"),
  #simulation_iterative_changesigns_meandiff |> mutate(Method = "Sign changing"),
  simulation_iterative_fabrication_meandiff |> mutate(Method = "Fabrication"),
  simulation_iterative_replacement_meandiff |> mutate(Method = "Replacement")
) |>
  mutate(Method = fct_relevel(Method, "Duplication", "Fabrication", "Dropping", "Replacement", "Condition swapping", "Sign changing"),
         n_total = n_per_condition*2)

p_meandiff_n_boxplot <- 
  ggplot(simulation_iterative_meandiff, 
         aes(as.factor(n_total), 
             tampers_needed, 
             color = Method,
             fill = Method)) +
  #geom_boxplot(outlier.shape = NA, width = 0.5, position = position_dodge(width = 0.9), coef = 1) +
  ggthemes::geom_tufteboxplot(median.type = "line", 
                              hoffset = 0, 
                              width = 6,
                              position = position_dodge(width = 0.85)) +
  labs(title = "Boxplot of counts of tampered cases needed to reach statistical significance",
       subtitle = "White gaps represent medians, colored bars represent interquartile range, whiskers represent range",
       x = "Sample size",
       y = "Count") +
  theme_linedraw() +
  theme(panel.grid.minor = element_blank()) +
  scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
  scale_x_discrete(breaks = c(10, seq(25, 200, by = 25))*2) +
  #scale_color_viridis_d(begin = 0.2, end = 0.8, option = "viridis") +
  #scale_fill_viridis_d(begin = 0.2, end = 0.8, option = "viridis") 
  scale_color_manual(values = wes_palette("Cavalcanti1")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1")) 

p_meandiff_prop_boxplot <- 
  ggplot(simulation_iterative_meandiff, 
         aes(as.factor(n_total), 
             tampers_needed/n_total, 
             color = Method,
             fill = Method)) +
  #geom_boxplot(outlier.shape = NA, width = 0.5, position = position_dodge(width = 0.9), coef = 1) +
  ggthemes::geom_tufteboxplot(median.type = "line", 
                              hoffset = 0, 
                              width = 6,
                              position = position_dodge(width = 0.85)) +
  labs(title = "Boxplot of proportions of tampered cases needed to reach statistical significance",
       subtitle = "White gaps represent medians, colored bars represent interquartile range, whiskers represent range",
       x = "Sample size",
       y = "Proportion") +
  theme_linedraw() +
  theme(panel.grid.minor = element_blank()) +
  scale_y_continuous(labels = percent_format(), limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
  scale_x_discrete(breaks = c(10, seq(25, 200, by = 25))*2) +
  #scale_color_viridis_d(begin = 0.2, end = 0.8, option = "viridis") +
  #scale_fill_viridis_d(begin = 0.2, end = 0.8, option = "viridis") 
  scale_color_manual(values = wes_palette("Cavalcanti1")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1")) 

p_meandiff_n_boxplot +
  p_meandiff_prop_boxplot +
  plot_layout(ncol = 1, guides = "collect")

ggsave(filename = "../communications/plots/iterative_meandiff_combined_boxplot.pdf",
       device = "pdf",
       width = 10,
       height = 8)

ggsave(filename = "../communications/plots/iterative_meandiff_combined_boxplot.png",
       device = "png",
       width = 10,
       height = 8)

```

### Table

```{r}

simulation_iterative_meandiff_summary |>
  relocate(Method, .before = n_total) |>
  dplyr::select(-mean, -sd, -prop_mean, -prop_sd) |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Correlations

```{r fig.height=3, fig.width=11}

p_iterative_exclusion_correlation_n +
  p_iterative_exclusion_correlation_prop +
  plot_layout(ncol = 2)

```

# Session info

```{r}

sessionInfo()

```


