---
title: "Little big lies: iterative data tampering in two-group mean-difference (t-test) designs"
author: "Ian Hussey"
date: today
format:
  html:
    embed-resources: true # ignored when rendering book
    toc: true
    toc_float: true
    code-fold: true
    code-overflow: wrap
    code-tools: true
    warning: false
    df-print: kable
    theme:
      light: cosmo
      dark: darkly
    highlight-style:
      light: breeze
      dark: dracula
    include-in-header:
      text: |
        <style>
          .kable-table table { width: 70%; margin: 0 auto; }
          .quarto-color-scheme-toggle::after { content: " Dark mode"; font-family: sans-serif; font-size: 0.85em; }
        </style>
---

# Introduction

This introduction is structured following the ADEMP framework (Aims, Data-generating process, Estimands, Methods, Performance) as described by Morris, White, and Crowther (2019) and operationalised for psychology by Siepe et al. (2024). Working through the framework explicitly makes the design decisions behind the simulations visible and auditable.

This file covers the **two-group mean-difference (Student's *t*-test)** setting. The companion files `compendium_iterative_correlation.Rmd` and `compendium_iterative_chisquare.Rmd` apply the same logic to a Pearson correlation and a 2×2 association respectively.

## Project description

This is a simulation study of **data tampering and fabrication**. It asks: starting from data generated under the null hypothesis (no true group difference), how much must a dishonest analyst tamper with the data before a two-sample *t*-test crosses *p* < .05? "Tampering" is operationalised as a family of iterative strategies — dropping cases, swapping condition labels, duplicating cases, flipping signs, fabricating new cases, and replacing cases — each applied a few data points at a time until significance is reached or a cap is hit.

The motivating context is the continuum between questionable research practices and outright fabrication: the same null dataset is pushed to significance by manipulations ranging from those usually filed under "*p*-hacking" (unprincipled exclusion of inconvenient cases) to those that are unambiguously fraud (typing a different condition label, inventing rows). Quantifying the *effort* — the count of altered data points — makes concrete how little manipulation can be required.

## Prior related work

There is no closed-form benchmark for the quantity of interest: the minimal number of tampered points needed to manufacture significance is a property of a greedy tampering algorithm applied to random data, not an estimator of a population parameter, so the study is **descriptive** rather than a calibration exercise. It connects to work on the fragility of findings to a few influential observations (the `dfstat`/`dfbeta` tradition; see also Spiess et al., 2024 *[citation to be completed]*) and to explicit "how to cheat" framings such as Chambers' (2017) "dirty dozen", whose second tip (condition swapping) is simulated directly here.

## Aims

The task is the **characterisation of a tampering procedure**. For each strategy and sample size the aims are:

1. To estimate **how many tampered data points are required to reach *p* < .05** from null data, and how this scales with sample size, both in absolute terms and as a proportion of the sample.
2. To estimate the **proportion of datasets that never reach significance** within a pre-set tampering budget (cap).
3. To quantify the **effect-size bias** of the final tampered dataset relative to the truth — Hedges' *g* relative to a true standardised effect of 0.
4. To **compare strategies** on these measures.

## Data-Generating Process

### DGP specification approach

The DGP is **fully parametric** and generated under the **true null**, so any significant result is a false positive produced entirely by the tampering. Two independent conditions of `n_per_condition` observations each are drawn from a Normal distribution with common SD: $Y_{ij} \overset{\text{iid}}{\sim} \mathcal{N}(\mu_j, \sigma^2)$ with $\mu_{\text{control}} = \mu_{\text{intervention}} = 0$ and $\sigma = 1$. Produced by `generate_data()` via `rnorm()`.

### DGP factors

Two factors are varied: the **tampering strategy** and the **per-group sample size** `n_per_condition`.

### Factor values and settings

- `n_per_condition` ∈ {10, 25, 50, 75, 100, 125, 150, 175, 200}.
- Tampering strategy ∈ {dropping, condition swapping, duplication, fabrication, replacement}.
- Held constant: μ = 0 in both groups and σ = 1 (the null); α = .05, two-sided; a tampering budget of at most 50% of the sample (`max_*_proportion = 0.5`); and the greedy "most-helpful-first" rule (each step alters the single data point that most moves the test toward significance).

### Factor combination and number of conditions

The strategy × sample-size grid is crossed **fully factorially**: 5 strategies × 9 sample sizes = **45 conditions**, each replicated `n_iterations` times.

## Estimands and Targets

- **Primary target — tamper count.** `tampers_needed`, the number of tampered data points required to first reach *p* < .05, summarised by its **median** (with IQR; rationale in *A note on summarising with the median rather than the mean*) and the **proportion never significant** within the cap.
- **Secondary target — effect-size bias.** Because the true effect is 0, the final-dataset **Hedges' *g*** (relative to 0) *is* the bias; reported as the signed value (a signed median, not an absolute magnitude, since the sign is the direction of the manufactured effect).

There is no analytic ground truth; the Monte Carlo estimates are the result.

## Methods and extracted quantities

Each strategy is an iterative greedy procedure: generate a null dataset, test it, and while non-significant (and under the cap) apply one tampering operation targeting the data point(s) most opposed to the desired result, then re-test. The analysis is the equal-variance two-sample *t*-test (`var.equal = TRUE`, two-sided, α = .05). From each replication the worker returns `tampers_needed`, `reached_sig` (whether *p* < .05 was actually reached, so no replication is silently dropped), and the final-dataset effect size (`hedges_g`, `bias_mean_diff`). Swapping *entire* conditions is handled analytically (it merely adds ½α to the false-positive rate) rather than simulated.

## Performance and Uncertainty

### Performance measures

The tamper-count distribution is summarised by its **median** and IQR (with mean/SD also tabulated to expose the skew) and as a **proportion of the sample**; the **proportion never significant** is reported separately. The bias is summarised by the **signed median Hedges' *g*** (which, since the truth is 0, is the bias itself). Full rationale for the median is in *A note on summarising with the median rather than the mean*.

### Monte Carlo uncertainty

Every summary carries a Monte Carlo standard error: the binomial SE for the never-significant proportion, the SE of the mean for the mean count, and a nonparametric bootstrap (Morris et al., 2019) for the median count and median bias. See the *Monte Carlo standard errors* subsection below.

### Number of simulation repetitions

Each condition is replicated `n_iterations` times: **10000** for a full run and **100** for a fast local/development render (auto-selected by the `on_hpc` knob; see *Running on an HPC cluster*). This is adequate for the descriptive comparison intended here.

### Non-convergence and missing values

The *t*-test is closed-form. The realistic failure modes are degenerate intermediate datasets (a group falling below two cases under dropping/replacement — guarded; zero within-group variance), which return a missing *p*-value and are classified as never-significant via `reached_sig`; summaries use `na.rm`.

### Interpretation of performance measures *(optional)*

There is no pass/fail criterion; the comparison is relative. A strategy is "cheap" to the extent that its median tamper count — especially its **proportion of the sample** — is small and its never-significant proportion is low. The expectation is that the proportion of data that must be tampered is small and does not grow with sample size, and that the distribution-preserving strategy (condition swapping) is both effective and the hardest to detect.

## Ethics, optics, and the *p*-hacking / fabrication continuum

This report documents weaponisable knowledge, and the framing matters; the simulations are designed to make a continuum visible, not to provide a recipe:

- Excluding participants is usually filed under *p*-hacking even when the exclusions are unprincipled and serve only to produce a significant *p*-value.
- But if, instead of deleting those rows, the analyst types the *other* condition's label into them, the same act becomes fabrication — and a firing offence.
- Treating one as common-and-merely-iffy and the other as rare-and-clearly-unacceptable risks badly underestimating the prevalence of the latter.
- The point is a trolley problem: is deleting data really ethically different from making it up, when both serve the individual's advancement at the expense of the evidence?

## Realism and scope

These are **proofs of principle**, not forensic models of real fraud. Some strategies (e.g. naive dropping) can leave obvious signatures such as a truncated otherwise-Normal distribution; many detected fraud cases were nonetheless unsophisticated, and more elaborate plausibility-preserving variants are easy to construct. The recurring theme is the illustration of **minimal tampering and effort**. Out of scope here: Likert-type outcomes; an "alter existing values" strategy distinct from fabrication; and a detectability axis.

# Dependencies

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

```{r}

library(tidyr)
library(dplyr)
library(readr)
library(forcats)
library(tibble)
library(purrr) 
library(furrr)
library(MASS)
library(effectsize)
library(ggplot2)
library(scales)
library(knitr)
library(kableExtra)
library(janitor)
library(ggtext)
library(patchwork)

options(scipen=999)

dir.create("../data", showWarnings = FALSE)
dir.create("../communications/plots", showWarnings = FALSE)

# --- Parallel plan: HPC-aware worker auto-detection -------------------------
# parallelly::availableCores() honours a SLURM allocation, cgroup CPU quotas,
# and affinity masks, so it returns the cores granted to THIS session rather than
# all physical cores. Set n_workers_override to pin a count; leave NULL to auto-
# detect. On the "Everything" HPC, 4-16 workers is the useful range.
n_workers_override <- NULL
detected_cores <- unname(parallelly::availableCores())
n_workers <- if (is.null(n_workers_override)) detected_cores else as.integer(n_workers_override)
n_workers <- max(1L, as.integer(n_workers))
message(sprintf("Parallel plan: %d workers (parallelly detected %d).", n_workers, detected_cores))

# --- Pin BLAS/LAPACK to one thread BEFORE workers spawn ---------------------
# With dozens of worker processes, multi-threaded BLAS would oversubscribe the
# node (workers x BLAS threads all fighting for cores). The operations here use
# tiny matrices, so single-threaded BLAS loses nothing. Setting these in the
# master means the multisession (PSOCK) workers inherit them at spawn.
Sys.setenv(FLEXIBLAS_NUM_THREADS = "1", OMP_NUM_THREADS = "1",
           OPENBLAS_NUM_THREADS = "1", MKL_NUM_THREADS = "1", BLIS_NUM_THREADS = "1",
           VECLIB_MAXIMUM_THREADS = "1") # VECLIB covers macOS Accelerate (local testing)
if (requireNamespace("flexiblas", quietly = TRUE)) flexiblas::flexiblas_set_num_threads(1L)
if (requireNamespace("RhpcBLASctl", quietly = TRUE)) {
  RhpcBLASctl::blas_set_num_threads(1)
  RhpcBLASctl::omp_set_num_threads(1)
}

# RStudio Server is not fork-safe, so use multisession (separate R processes),
# not multicore. furrr_options(seed = TRUE) advances an L'Ecuyer-CMRG stream per
# task, giving reproducibility independent of the worker count.
plan(multisession, workers = n_workers)

```

# Common functions and global variables

```{r}

# --- Replication knob (local vs cluster) ------------------------------------
# RStudio Server does not pass the job's SLURM_* allocation variables into the R
# session, but the HPC sets some SLURM_* global that survives and is absent on a
# laptop, so "any SLURM_* variable present" is a reliable cluster detector.
# run_full forces the choice: NULL = auto-detect, TRUE = full run, FALSE = quick
# local smoke test.
run_full <- TRUE
on_hpc <- if (!is.null(run_full)) isTRUE(run_full) else any(grepl("^SLURM_", names(Sys.getenv())))

# n iterations for all sims: full cluster run vs fast local smoke test
n_iterations <- if (on_hpc) 10000L else 100L

# Cache tag: keeps the full and dev runs in separate .rds files so they never
# collide. cache_path() builds a tagged path under ../data for each simulation.
cache_tag <- paste0(if (on_hpc) "full" else "dev", "_k", n_iterations)
cache_path <- function(stem) paste0("../data/", stem, "_", cache_tag, ".rds")

# Packages each worker must attach (belt-and-braces; future also auto-detects
# the packages of exported globals). MASS is only needed by the correlation sims.
worker_packages <- c("dplyr", "tidyr", "tibble", "purrr", "effectsize", "MASS")

# define data generating function ----
generate_data <- function(n_per_condition,
                          mean_control,
                          mean_intervention,
                          sd_control,
                          sd_intervention) {
  
  data_control <- 
    tibble(condition = "control",
           score = rnorm(n = n_per_condition, mean = mean_control, sd = sd_control))
  
  data_intervention <- 
    tibble(condition = "intervention",
           score = rnorm(n = n_per_condition, mean = mean_intervention, sd = sd_intervention))
  
  data <- bind_rows(data_control,
                    data_intervention) 
  
  return(data)
}

# define data analysis function ----
# slow, because t.test is slow. 
analyse_data <- function(data) {
  
  res_t_test <- t.test(formula = score ~ condition,
                       data = data,
                       var.equal = TRUE,
                       alternative = "two.sided")
  
  res <- tibble(p = res_t_test$p.value)

  return(res)
}

# extract Hedges' g and the raw mean-difference bias from a (possibly tampered) dataset ----
# The data are generated under the null (mean_control = mean_intervention = 0), so the
# true standardised effect is 0. Hedges' g of the final, tampered dataset is therefore a
# direct quantification of the bias the tampering has introduced relative to mu = 0.
# Hedges' g is signed intervention-minus-control (positive = intervention scored higher).
# The signed g is reported throughout: because the truth is 0, the signed median IS the
# bias, and it also shows the direction each strategy induces. (An absolute |g| would be
# misleading: it is strictly positive even under no systematic bias.)
compute_bias <- function(data) {

  data <- data |> dplyr::filter(!is.na(score))

  # force factor order so g is computed as intervention - control
  data <- data |> mutate(condition = factor(condition, levels = c("intervention", "control")))

  hedges_g_value <- tryCatch(
    as.numeric(effectsize::hedges_g(score ~ condition, data = data)$Hedges_g),
    error = function(e) NA_real_
  )

  means <- data |>
    dplyr::group_by(condition) |>
    dplyr::summarise(m = mean(score), .groups = "drop")

  m_control      <- means$m[means$condition == "control"]
  m_intervention <- means$m[means$condition == "intervention"]
  if (length(m_control) == 0)      m_control      <- NA_real_
  if (length(m_intervention) == 0) m_intervention <- NA_real_

  tibble(hedges_g       = hedges_g_value,
         bias_mean_diff = m_intervention - m_control)
}

# bundle the number of tampered cases with the final-dataset effect size / bias ----
# reached_sig records whether the study actually crossed p < .05 (TRUE) or stopped
# because it hit the tampering cap (FALSE). Tracking this explicitly — rather than
# inferring it from tampers_needed vs the cap — means a result can never be silently
# dropped from both the central-tendency stats and the never-significant proportion.
summarise_tamper_result <- function(data, tampers_needed, reached_sig) {
  bind_cols(tibble(tampers_needed = tampers_needed, reached_sig = reached_sig),
            compute_bias(data))
}

# # manual t-test p-value calculation function
# analyse_data <- function(data) {
#   x <- data$score[data$condition == "control"]
#   y <- data$score[data$condition == "intervention"]
#   
#   # Basic checks like t.test() would do
#   if (length(x) < 2 || length(y) < 2 || all(is.na(x)) || all(is.na(y))) {
#     return(tibble(p = NA_real_))
#   }
#   
#   if (sd(x, na.rm = TRUE) == 0 || sd(y, na.rm = TRUE) == 0) {
#     return(tibble(p = NA_real_))
#   }
# 
#   # Remove NAs for safety
#   x <- x[!is.na(x)]
#   y <- y[!is.na(y)]
#   
#   n1 <- length(x)
#   n2 <- length(y)
#   m1 <- mean(x)
#   m2 <- mean(y)
#   s1 <- sd(x)
#   s2 <- sd(y)
# 
#   s_pooled <- sqrt(((n1 - 1) * s1^2 + (n2 - 1) * s2^2) / (n1 + n2 - 2))
#   
#   # Catch divide-by-zero or NaN
#   if (is.nan(s_pooled) || s_pooled == 0) {
#     return(tibble(p = NA_real_))
#   }
# 
#   t_stat <- (m1 - m2) / (s_pooled * sqrt(1/n1 + 1/n2))
#   df <- n1 + n2 - 2
#   p_val <- 2 * pt(-abs(t_stat), df)
# 
#   if (is.na(p_val)) p_val <- NA_real_
# 
#   return(tibble(p = p_val))
# }

# Monte Carlo standard errors (MCSEs) ----
# The mean and proportion MCSEs use the closed-form expressions from Siepe et al. (2024)
# / Morris et al. (2019), matching simulation_template_estimation.qmd:
#   MCSE(mean)       = SD / sqrt(K)
#   MCSE(proportion) = sqrt(p (1 - p) / K)
# The median has no neat closed form under skew, so its MCSE is obtained by the
# nonparametric bootstrap recommended by Morris et al. (2019). The RNG state is saved
# and restored so that computing summaries never perturbs the simulation seeds.
mcse_mean <- function(x){
  x <- x[!is.na(x)]
  if (length(x) < 2) return(NA_real_)
  sd(x) / sqrt(length(x))
}

mcse_proportion <- function(p, k){
  if (is.na(p) || k < 1) return(NA_real_)
  sqrt(p * (1 - p) / k)
}

mcse_median <- function(x, n_boot = 1000, seed = 42){
  x <- x[!is.na(x)]
  n <- length(x)
  if (n < 2) return(NA_real_)
  old_seed <- if (exists(".Random.seed", envir = .GlobalEnv)) get(".Random.seed", envir = .GlobalEnv) else NULL
  set.seed(seed)
  boot_medians <- replicate(n_boot, median(x[sample.int(n, n, replace = TRUE)]))
  if (!is.null(old_seed)) assign(".Random.seed", old_seed, envir = .GlobalEnv)
  sd(boot_medians)
}

summarise_err <- function(data){
  data |>
    group_by(n_per_condition, mean_intervention) |>
    # central-tendency stats are computed only over iterations that actually reached
    # significance (reached_sig); the rest are summarised by proportion_never_sig. The
    # two are complementary, so no iteration is dropped from both (the previous
    # tampers_needed-vs-cap comparison silently lost cases whose count overshot an odd cap).
    summarize(k = n(), # total iterations in this cell
              k_reached = sum(reached_sig), # iterations that reached significance
              proportion_never_sig = mean(!reached_sig),
              proportion_never_sig_mcse = mcse_proportion(proportion_never_sig, k),
              min = min(tampers_needed[reached_sig]),
              max = max(tampers_needed[reached_sig]),
              mean = mean(tampers_needed[reached_sig]),
              mean_mcse = mcse_mean(tampers_needed[reached_sig]),
              sd = sd(tampers_needed[reached_sig]),
              median = median(tampers_needed[reached_sig]),
              median_mcse = mcse_median(tampers_needed[reached_sig]),
              mad = mad(tampers_needed[reached_sig]),
              lower = quantile(tampers_needed[reached_sig], probs = 0.25),
              upper = quantile(tampers_needed[reached_sig], probs = 0.75)) |>
    ungroup() |>
    mutate(n_total = n_per_condition*2,
           prop_min = min/n_total,
           prop_max = max/n_total,
           prop_mean = mean/n_total,
           prop_mean_mcse = mean_mcse/n_total,
           prop_sd = sd/n_total,
           prop_median = median/n_total,
           prop_median_mcse = median_mcse/n_total,
           prop_mad = median/n_total,
           prop_lower = lower / n_total,
           prop_upper = upper / n_total) |>
    relocate(n_total, .before = n_per_condition) |>
    relocate(mean_mcse, .after = mean) |>
    relocate(median_mcse, .after = median) |>
    dplyr::select(-n_per_condition)
}

# summarise the bias (Hedges' g relative to mu = 0) in the final tampered datasets ----
summarise_bias <- function(data){
  data |>
    group_by(n_per_condition, mean_intervention) |>
    summarize(g_median        = median(hedges_g, na.rm = TRUE),
              g_median_mcse   = mcse_median(hedges_g),
              g_lower         = quantile(hedges_g, probs = 0.25, na.rm = TRUE),
              g_upper         = quantile(hedges_g, probs = 0.75, na.rm = TRUE),
              bias_median       = median(bias_mean_diff, na.rm = TRUE),
              bias_median_mcse  = mcse_median(bias_mean_diff),
              bias_lower        = quantile(bias_mean_diff, probs = 0.25, na.rm = TRUE),
              bias_upper        = quantile(bias_mean_diff, probs = 0.75, na.rm = TRUE)) |>
    ungroup() |>
    mutate(n_total = n_per_condition * 2) |>
    relocate(n_total, .before = n_per_condition) |>
    dplyr::select(-n_per_condition)
}

# plot the (signed) bias (Hedges' g relative to mu = 0) against sample size ----
plot_bias <- function(data, title){
  ggplot(data, aes(n_total, g_median)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
    geom_point() +
    geom_linerange(aes(ymin = g_lower, ymax = g_upper)) +
    labs(title = title,
         x = "Sample size",
         y = "Hedges' g in final dataset (bias relative to μ = 0)") +
    theme_linedraw() +
    scale_y_continuous(breaks = breaks_pretty(n = 8)) +
    scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2)
}

plot_distributions <- function(data, title, x_axis){
  ggplot(data, aes(x = tampers_needed)) +
    geom_histogram(binwidth = 1) +
    labs(title = title, 
         x = x_axis,
         y = "Count") +
    ylim(0, NA) +
    theme_linedraw() +
    scale_x_continuous(breaks = breaks_pretty(n = 5)) +
    facet_wrap(~ n_per_condition) # N PER CONDITION NOT TOTAL N
}

plot_tampers_n <- function(data, title, y_axis){
  ggplot(data, aes(n_total, median)) +
    geom_point() +
    geom_linerange(aes(ymin = lower, ymax = upper)) +
    labs(title = title,
         x = "Sample size",
         y = y_axis) +
    theme_linedraw() +
    scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
    scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2)
}

plot_tampers_prop <- function(data, title, y_axis){
  ggplot(data, aes(n_total, prop_median)) +
    geom_point() +
    geom_linerange(aes(ymin = prop_lower, ymax = prop_upper)) +
    labs(title = title,
         x = "Sample size",
         y = y_axis) +
    theme_linedraw() +
    scale_y_continuous(labels = percent_format(), limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
    scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2)
}

```

# A note on summarising with the median rather than the mean

Each simulation records `tampers_needed`: the number of tampering operations applied before the result crossed *p* < .05. This quantity is summarised throughout with the **median** (and the interquartile range / `mad`) rather than the mean. There are three reasons.

1. **The distribution is right-skewed.** Most datasets reach significance after only a few tampers, but a long tail require many. Part of the point of these simulations is to *show* that skew, so the central-tendency measure should not hide it. The summary tables report the `mean` and `sd` alongside the `median` and `mad` precisely so the skew is visible: the mean sits above the median, and the size of that gap indexes the length of the tail.

2. **The distribution is censored.** Tampering stops at a cap (`floor(0.5 × n_total)`), and a proportion of datasets never reach significance within that cap (reported separately as `proportion_never_sig`). Those censored cases have no finite tamper count, so the *unconditional* mean is not well defined; the central-tendency statistics are therefore computed only over the datasets that did reach significance before the cap. The median is robust to this upper censoring — it is unaffected as long as the central value lies below the cap — whereas the mean is not.

3. **The median refers to an actual case.** For an integer count, a mean such as "2.7 tampers" corresponds to no real study, whereas a median of 2 describes a realisable scenario ("half the time, two edits were enough"). Because the substantive claim concerns the realistic effort an individual would expend, the statistic that points at an actual case is the more faithful one.

Note that the central-tendency statistics **include** the datasets that were significant from the outset and so required zero tampers (the ~5% expected under the null at α = .05). The median therefore answers "starting from null data, how many edits to significance?" rather than "*given* that tampering was needed, how many?". This framing was chosen deliberately because it matches the question of how readily null data can be pushed to significance.

## Monte Carlo standard errors

Because each summary statistic is itself estimated from a finite number of simulation iterations (`n_iterations`), it carries Monte Carlo error. Following the approach in `simulation_template_estimation.qmd` (and Siepe et al., 2024; Morris et al., 2019), every table reports a Monte Carlo standard error (MCSE) next to each estimated quantity, so that apparent differences between conditions can be judged against simulation noise. The MCSEs are computed as:

- **Proportion never significant** (`proportion_never_sig_mcse`): the binomial standard error $\sqrt{\hat p\,(1-\hat p)/K}$, where *K* = `n_iterations`.
- **Mean** (`mean_mcse`): the standard error of the mean $\text{SD}/\sqrt{K_\text{reached}}$, where $K_\text{reached}$ is the number of iterations that reached significance before the cap (i.e. the iterations the mean is computed over).
- **Median** (`median_mcse`): obtained by nonparametric bootstrap (1,000 resamples), because the count distribution is skewed and has no convenient closed-form quantile MCSE. The same bootstrap is used for the median Hedges' g and median raw-bias columns.

The `median_mcse` is the relevant uncertainty for the headline statistic; with `n_iterations` in the low thousands it is typically a small fraction of a tamper, confirming the medians are estimated precisely. Because the tamper counts are discrete, the bootstrap median MCSE is sometimes exactly **0**: this is not an error but indicates that the median is invariant to resampling — its sampling variability is below the one-unit resolution of the data, so the median is estimated to sub-count precision. (The bootstrap saves and restores the RNG state, so it never disturbs the simulation seeds.) The `mean_mcse`, which is always non-zero, provides a complementary, continuous-scale check on Monte Carlo precision in these cases.

# Dropping

Unwarranted exclusion of cases: each round, drop the single most hostile remaining case (the lowest scorer in the intervention group or the highest scorer in the control group, whichever is the more extreme within its own group), then re-test. Repeat until significance is reached or the cap is hit.

## Simulation functions

```{r}

# Drop the single most hostile remaining case (lowest in intervention or highest in control)
iterative_drop_once <- function(data) {

  data_incl <- data |> filter(!is.na(score))

  data_incl <- data_incl |>
    group_by(condition) |>
    mutate(z_score_group = (score - mean(score)) / sd(score)) |>
    ungroup()

  smallest_in_intervention <- data_incl |>
    filter(condition == "intervention") |>
    slice_min(z_score_group, n = 1)

  largest_in_control <- data_incl |>
    filter(condition == "control") |>
    slice_max(z_score_group, n = 1)

  to_drop <- bind_rows(smallest_in_intervention, largest_in_control) |>
    slice_max(abs(z_score_group), n = 1)

  data <- data |> mutate(score = ifelse(row_id == to_drop$row_id, NA, score))

  return(data)
}

# Main simulation function
run_iterative_dropping_study <- function(n_per_condition,
                                         mean_control,
                                         mean_intervention,
                                         sd_control,
                                         sd_intervention,
                                         max_drop_proportion) {

  # Generate original data
  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) |>
    mutate(row_id = row_number())

  # Initial p-value
  p_val <- analyse_data(data)$p[1]
  if (p_val < 0.05) return(summarise_tamper_result(data, 0, TRUE))

  n_total <- nrow(data)
  max_drops <- floor(max_drop_proportion * n_total)
  drops_so_far <- 0

  while (!is.na(p_val) && p_val >= 0.05 && drops_so_far < max_drops) {
    # Stop if dropping would leave a condition with fewer than 2 cases
    data_incl <- data |> filter(!is.na(score))
    if (any(table(data_incl$condition) < 2)) break

    data <- iterative_drop_once(data)
    drops_so_far <- drops_so_far + 1
    p_val <- analyse_data(data)$p[1]
  }

  return(summarise_tamper_result(data, drops_so_far, !is.na(p_val) && p_val < 0.05))
}

```

## Run simulation

```{r}

simulation_iterative_dropping_meandiff_path <- cache_path("simulation_iterative_dropping_meandiff")
if (file.exists(simulation_iterative_dropping_meandiff_path) &&
    "reached_sig" %in% names(read_rds(simulation_iterative_dropping_meandiff_path))) {
  simulation_iterative_dropping_meandiff <- read_rds(simulation_iterative_dropping_meandiff_path)
} else {

  # Parameter grid
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    max_drop_proportion = 0.5,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_iterative_dropping_meandiff <- experiment_parameters_grid |>
    mutate(res = future_pmap(.l = list(n_per_condition = n_per_condition,
                                       mean_control = mean_control,
                                       mean_intervention = mean_intervention,
                                       sd_control = sd_control,
                                       sd_intervention = sd_intervention,
                                       max_drop_proportion = max_drop_proportion),
                             .f = run_iterative_dropping_study,
                             .progress = TRUE,
                             .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages))) |>
    unnest(res)

  write_rds(simulation_iterative_dropping_meandiff, simulation_iterative_dropping_meandiff_path)
}

# summarize tamper counts and bias (Hedges' g relative to mu = 0)
simulation_iterative_dropping_meandiff_err <- summarise_err(simulation_iterative_dropping_meandiff)
simulation_iterative_dropping_meandiff_bias    <- summarise_bias(simulation_iterative_dropping_meandiff)

```

## Table

```{r}

simulation_iterative_dropping_meandiff_err |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Plot

```{r}

plot_distributions(simulation_iterative_dropping_meandiff, "Dropping", "N dropped")

p_iterative_dropping_meandiff_n <- plot_tampers_n(simulation_iterative_dropping_meandiff_err, "Dropping", "N dropped")

p_iterative_dropping_meandiff_prop <- plot_tampers_prop(simulation_iterative_dropping_meandiff_err, "Dropping", "N dropped")

```

## Bias (Hedges' g relative to μ = 0)

```{r}

simulation_iterative_dropping_meandiff_bias |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

p_iterative_dropping_meandiff_bias <- plot_bias(simulation_iterative_dropping_meandiff_bias, "Dropping")

p_iterative_dropping_meandiff_bias

```

# Condition swapping

## Simulation functions

```{r}

# Swap one participant from one condition to the other.
# A swap moves a single participant's label, so it is counted as ONE tamper (matching
# the other methods). To keep the two group sizes balanced over pairs of swaps, the
# direction alternates: when move_intervention_to_control is TRUE we move the most
# hostile intervention case (lowest score, which drags the intervention mean down) to
# control; otherwise we move the most hostile control case (highest score) to intervention.
iterative_condition_swap_once <- function(data, move_intervention_to_control) {

  data <- data |>
    group_by(condition) |>
    mutate(z_score = (score - mean(score)) / sd(score)) |>
    ungroup()

  if (move_intervention_to_control) {
    to_swap   <- data |> filter(condition == "intervention") |> slice_min(z_score, n = 1)
    new_label <- "control"
  } else {
    to_swap   <- data |> filter(condition == "control") |> slice_max(z_score, n = 1)
    new_label <- "intervention"
  }

  if (nrow(to_swap) < 1) {
    return(data |> dplyr::select(-z_score))
  }

  data |>
    mutate(condition = ifelse(row_number == to_swap$row_number, new_label, condition)) |>
    dplyr::select(-z_score)
}

# Main simulation function
run_iterative_condition_swap_study <- function(n_per_condition,
                                               mean_control,
                                               mean_intervention,
                                               sd_control,
                                               sd_intervention,
                                               max_swap_proportion) {

  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) |>
    mutate(row_number = row_number())

  res <- analyse_data(data)
  p_val <- res$p[1]

  if(p_val < 0.05) {
    return(summarise_tamper_result(data, 0, TRUE))
  }

  n_total <- nrow(data)
  max_swaps <- floor(max_swap_proportion * n_total)
  swaps_so_far <- 0

  while(p_val >= 0.05 && swaps_so_far < max_swaps) {
    # alternate swap direction so group sizes stay balanced over pairs of swaps
    move_intervention_to_control <- (swaps_so_far %% 2 == 0)
    data <- iterative_condition_swap_once(data, move_intervention_to_control)
    swaps_so_far <- swaps_so_far + 1
    res <- analyse_data(data)
    p_val <- res$p[1]
  }

  return(summarise_tamper_result(data, swaps_so_far, !is.na(p_val) && p_val < 0.05))
}

```

## Run simulation

```{r}

# Run simulation or load existing (re-run if a cached file predates the hedges_g column)
simulation_iterative_swaps_meandiff_path <- cache_path("simulation_iterative_swaps_meandiff")
if(file.exists(simulation_iterative_swaps_meandiff_path) &&
   "reached_sig" %in% names(read_rds(simulation_iterative_swaps_meandiff_path))) {
  simulation_iterative_swaps_meandiff <- read_rds(simulation_iterative_swaps_meandiff_path)
} else {

  # Parameter grid
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    max_swap_proportion = 0.5,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_iterative_swaps_meandiff <- experiment_parameters_grid |>
    mutate(res = future_pmap(.l = list(n_per_condition = n_per_condition,
                                       mean_control = mean_control,
                                       mean_intervention = mean_intervention,
                                       sd_control = sd_control,
                                       sd_intervention = sd_intervention,
                                       max_swap_proportion = max_swap_proportion),
                             .f = run_iterative_condition_swap_study,
                             .progress = TRUE,
                             .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages))) |>
    unnest(res)

  write_rds(simulation_iterative_swaps_meandiff, simulation_iterative_swaps_meandiff_path)
}

# summarize tamper counts and bias (Hedges' g relative to mu = 0)
simulation_iterative_swaps_meandiff_err <- summarise_err(simulation_iterative_swaps_meandiff)
simulation_iterative_swaps_meandiff_bias    <- summarise_bias(simulation_iterative_swaps_meandiff)

```

## Table

```{r}

simulation_iterative_swaps_meandiff_err |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Plot

```{r}

plot_distributions(simulation_iterative_swaps_meandiff, "Swaps", "N condition swaps")

p_iterative_swaps_meandiff_n <- plot_tampers_n(simulation_iterative_swaps_meandiff_err, "Swaps", "N condition swaps")

p_iterative_swaps_meandiff_prop <- plot_tampers_prop(simulation_iterative_swaps_meandiff_err, "Swaps", "N condition swaps")

```

## Bias (Hedges' g relative to μ = 0)

```{r}

simulation_iterative_swaps_meandiff_bias |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

p_iterative_swaps_meandiff_bias <- plot_bias(simulation_iterative_swaps_meandiff_bias, "Swaps")

p_iterative_swaps_meandiff_bias

```

# Duplication

From top 10% most favorable data points

## Simulation functions

```{r}

# Function to perform one round of top-10%-based duplication
duplicate_from_top_10_percent_once <- function(data, original_top10_control, original_top10_intervention) {
  
  # Sample one from each condition's original top 10%
  dup_control <- sample_n(original_top10_control, 1)
  dup_intervention <- sample_n(original_top10_intervention, 1)
  
  # Append to data
  data <- bind_rows(data, dup_control, dup_intervention)
  
  return(data)
}

# Main simulation function
run_iterative_duplication_study <- function(n_per_condition,
                                            mean_control,
                                            mean_intervention,
                                            sd_control,
                                            sd_intervention,
                                            max_duplication_proportion,
                                            proportion) {
  
  # Generate original data and label with row IDs
  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) %>%
    mutate(row_id = row_number())
  
  # Compute original top 10% favorable scores per condition
  top10_control <- data %>% 
    filter(condition == "control") %>% 
    slice_max(score, prop = proportion)
  
  top10_intervention <- data %>% 
    filter(condition == "intervention") %>% 
    slice_min(score, prop = proportion)
  
  # Initial p-value
  p_val <- analyse_data(data)$p[1]
  if (p_val < 0.05) return(summarise_tamper_result(data, 0, TRUE))

  n_total <- nrow(data)
  max_dups <- floor(max_duplication_proportion * n_total)
  dups_so_far <- 0

  while (p_val >= 0.05 && dups_so_far < max_dups) {
    data <- duplicate_from_top_10_percent_once(data, top10_control, top10_intervention)
    p_val <- analyse_data(data)$p[1]
    dups_so_far <- dups_so_far + 2
  }

  return(summarise_tamper_result(data, dups_so_far, !is.na(p_val) && p_val < 0.05))
}

```

## Run simulation

```{r}

simulation_iterative_duplication_meandiff_path <- cache_path("simulation_iterative_duplication_meandiff")
if (file.exists(simulation_iterative_duplication_meandiff_path) &&
    "reached_sig" %in% names(read_rds(simulation_iterative_duplication_meandiff_path))) {
  simulation_iterative_duplication_meandiff <- read_rds(simulation_iterative_duplication_meandiff_path)
} else {

  # Parameter grid
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    proportion = 0.1,
    max_duplication_proportion = 0.5,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_iterative_duplication_meandiff <- experiment_parameters_grid %>%
    mutate(res = future_pmap(.l = list(n_per_condition = n_per_condition,
                                       mean_control = mean_control,
                                       mean_intervention = mean_intervention,
                                       sd_control = sd_control,
                                       sd_intervention = sd_intervention,
                                       proportion = proportion),
                             .f = run_iterative_duplication_study,
                             .progress = TRUE,
                             .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages))) |>
    unnest(res)

  write_rds(simulation_iterative_duplication_meandiff, simulation_iterative_duplication_meandiff_path)
}

# summarise tamper counts and bias (Hedges' g relative to mu = 0)
simulation_iterative_duplication_meandiff_err <- summarise_err(simulation_iterative_duplication_meandiff)
simulation_iterative_duplication_meandiff_bias    <- summarise_bias(simulation_iterative_duplication_meandiff)

```

## Table

```{r}

simulation_iterative_duplication_meandiff_err |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Plot

```{r}

plot_distributions(simulation_iterative_duplication_meandiff, "Duplications", "N duplications")

p_iterative_duplication_meandiff_n <- plot_tampers_n(simulation_iterative_duplication_meandiff_err, "Duplications", "N duplications")

p_iterative_duplication_meandiff_prop <- plot_tampers_prop(simulation_iterative_duplication_meandiff_err, "Duplications", "N duplications")

```

## Bias (Hedges' g relative to μ = 0)

```{r}

simulation_iterative_duplication_meandiff_bias |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

p_iterative_duplication_meandiff_bias <- plot_bias(simulation_iterative_duplication_meandiff_bias, "Duplications")

p_iterative_duplication_meandiff_bias

```

# Fabrication

## Illustrate 

for the correlation sim, m = .93 so that cohen's d = 5.0 == pearson's r = .93

```{r}

x_vals <- seq(-7, 7, length.out = 1000)

df <- tibble(
  x = rep(x_vals, 2),
  density = c(dnorm(x_vals, mean = 0, sd = 1),
              dnorm(x_vals, mean = 5, sd = 0.2)),
  distribution = rep(c("N(0,1)", "N(1,0.2)"), each = length(x_vals))
)

ggplot(df, aes(x = x, y = density, color = distribution)) +
  geom_line(linewidth = 1) +
  labs(x = "Value", y = "Density") +
  theme_linedraw()

```

## Simulation functions

```{r}

run_iterative_fabrication_study <- function(n_per_condition,
                                            mean_control,
                                            mean_intervention,
                                            sd_control,
                                            sd_intervention,
                                            fabricated_control_mean,
                                            fabricated_control_sd,
                                            fabricated_intervention_mean,
                                            fabricated_intervention_sd,
                                            max_add_proportion) {
  
  # Generate original data
  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) %>%
    mutate(row_id = row_number())
  
  # Initial p-value
  p_val <- analyse_data(data)$p[1]
  if (p_val < 0.05) return(summarise_tamper_result(data, 0, TRUE))

  n_total <- nrow(data)
  max_adds <- floor(max_add_proportion * n_total)
  adds_so_far <- 0

  while (p_val >= 0.05 && adds_so_far < max_adds) {
    # Add one control and one intervention point per iteration
    new_control <- tibble(condition = "control", score = rnorm(1, mean = fabricated_control_mean, sd = fabricated_control_sd))
    new_intervention <- tibble(condition = "intervention", score = rnorm(1, mean = fabricated_intervention_mean, sd = fabricated_intervention_sd))

    data <- bind_rows(data, new_control, new_intervention)
    adds_so_far <- adds_so_far + 2

    p_val <- analyse_data(data)$p[1]
  }

  return(summarise_tamper_result(data, adds_so_far, !is.na(p_val) && p_val < 0.05))
}

```

## Run simulation

```{r}

simulation_iterative_fabrication_meandiff_path <- cache_path("simulation_iterative_fabrication_meandiff")
if (file.exists(simulation_iterative_fabrication_meandiff_path) &&
    "reached_sig" %in% names(read_rds(simulation_iterative_fabrication_meandiff_path))) {
  simulation_iterative_fabrication_meandiff <- read_rds(simulation_iterative_fabrication_meandiff_path)
} else {

  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    fabricated_control_mean = 0,
    fabricated_control_sd = 0.2,
    fabricated_intervention_mean = 5,
    fabricated_intervention_sd = 0.2,
    max_add_proportion = 0.5,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_iterative_fabrication_meandiff <- experiment_parameters_grid %>%
    mutate(res = future_pmap(.l = list(n_per_condition = n_per_condition,
                                       mean_control = mean_control,
                                       mean_intervention = mean_intervention,
                                       sd_control = sd_control,
                                       sd_intervention = sd_intervention,
                                       fabricated_control_mean = fabricated_control_mean,
                                       fabricated_control_sd = fabricated_control_sd,
                                       fabricated_intervention_mean = fabricated_intervention_mean,
                                       fabricated_intervention_sd = fabricated_intervention_sd,
                                       max_add_proportion = max_add_proportion),
                             .f = run_iterative_fabrication_study,
                             .progress = TRUE,
                             .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages))) |>
    unnest(res)

  write_rds(simulation_iterative_fabrication_meandiff, simulation_iterative_fabrication_meandiff_path)
}

# summarise tamper counts and bias (Hedges' g relative to mu = 0)
simulation_iterative_fabrication_meandiff_err <- summarise_err(simulation_iterative_fabrication_meandiff)
simulation_iterative_fabrication_meandiff_bias    <- summarise_bias(simulation_iterative_fabrication_meandiff)

```

## Table

```{r}

simulation_iterative_fabrication_meandiff_err |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Plot

```{r}

plot_distributions(simulation_iterative_fabrication_meandiff, "Fabrications", "N fabricated")

p_iterative_fabrication_meandiff_n <- plot_tampers_n(simulation_iterative_fabrication_meandiff_err, "Fabrications", "N fabricated")

p_iterative_fabrication_meandiff_prop <- plot_tampers_prop(simulation_iterative_fabrication_meandiff_err, "Fabrications", "N fabricated")

```

## Bias (Hedges' g relative to μ = 0)

```{r}

simulation_iterative_fabrication_meandiff_bias |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

p_iterative_fabrication_meandiff_bias <- plot_bias(simulation_iterative_fabrication_meandiff_bias, "Fabrications")

p_iterative_fabrication_meandiff_bias

```

# Replacement

## Simulation functions

```{r}

run_iterative_replacement_study <- function(n_per_condition,
                                            mean_control,
                                            mean_intervention,
                                            sd_control,
                                            sd_intervention,
                                            fabricated_control_mean,
                                            fabricated_control_sd,
                                            fabricated_intervention_mean,
                                            fabricated_intervention_sd,
                                            max_replace_proportion) {
  
  # Generate original data
  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) %>%
    mutate(row_id = row_number())
  
  # Initial p-value
  p_val <- analyse_data(data)$p[1]
  if (p_val < 0.05) return(summarise_tamper_result(data, 0, TRUE))

  n_total <- nrow(data)
  max_replacements <- floor(max_replace_proportion * n_total)
  replacements_so_far <- 0
  
  while (!is.na(p_val) && p_val >= 0.05 && replacements_so_far < max_replacements) {
    # Filter out NA scores
    data_incl <- data |> filter(!is.na(score))
    
    if (any(table(data_incl$condition) < 2)) break
    
    data_incl <- data_incl |> 
      group_by(condition) |> 
      mutate(z_score_group = (score - mean(score)) / sd(score)) |> 
      ungroup()
    
    smallest_in_intervention <- data_incl |> 
      filter(condition == "intervention") |> 
      slice_min(z_score_group, n = 1)
    
    largest_in_control <- data_incl |> 
      filter(condition == "control") |> 
      slice_max(z_score_group, n = 1)
    
    to_replace <- bind_rows(smallest_in_intervention, largest_in_control) |> 
      slice_max(abs(z_score_group), n = 1)
    
    replacement_row <- if (to_replace$condition == "control") {
      tibble(score = rnorm(1, fabricated_control_mean, fabricated_control_sd))
    } else {
      tibble(score = rnorm(1, fabricated_intervention_mean, fabricated_intervention_sd))
    }
    
    data <- data |> mutate(score = ifelse(row_id == to_replace$row_id, replacement_row$score, score))
    
    replacements_so_far <- replacements_so_far + 1
    p_val <- analyse_data(data)$p[1]
  }

  return(summarise_tamper_result(data, replacements_so_far, !is.na(p_val) && p_val < 0.05))
}

```

## Run simulation

```{r}

simulation_iterative_replacement_meandiff_path <- cache_path("simulation_iterative_replacement_meandiff")
if (file.exists(simulation_iterative_replacement_meandiff_path) &&
    "reached_sig" %in% names(read_rds(simulation_iterative_replacement_meandiff_path))) {
  simulation_iterative_replacement_meandiff <- read_rds(simulation_iterative_replacement_meandiff_path)
} else {

  # Parameter grid
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    fabricated_control_mean = 0,
    fabricated_control_sd = 0.2,
    fabricated_intervention_mean = 5,
    fabricated_intervention_sd = 0.2,
    max_replace_proportion = 0.5,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_iterative_replacement_meandiff <- experiment_parameters_grid %>%
    mutate(res = future_pmap(.l = list(n_per_condition = n_per_condition,
                                       mean_control = mean_control,
                                       mean_intervention = mean_intervention,
                                       sd_control = sd_control,
                                       sd_intervention = sd_intervention,
                                       fabricated_control_mean = fabricated_control_mean,
                                       fabricated_control_sd = fabricated_control_sd,
                                       fabricated_intervention_mean = fabricated_intervention_mean,
                                       fabricated_intervention_sd = fabricated_intervention_sd,
                                       max_replace_proportion = max_replace_proportion),
                             .f = run_iterative_replacement_study,
                             .progress = TRUE,
                             .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages))) |>
    unnest(res)

  write_rds(simulation_iterative_replacement_meandiff, simulation_iterative_replacement_meandiff_path)
}

# Summarize tamper counts and bias (Hedges' g relative to mu = 0)
simulation_iterative_replacement_meandiff_err <- summarise_err(simulation_iterative_replacement_meandiff)
simulation_iterative_replacement_meandiff_bias    <- summarise_bias(simulation_iterative_replacement_meandiff)

```

## Table

```{r}

simulation_iterative_replacement_meandiff_err |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Plot

```{r}

plot_distributions(simulation_iterative_replacement_meandiff, "Replacing", "N replaced")

p_iterative_replacement_meandiff_n <- plot_tampers_n(simulation_iterative_replacement_meandiff_err, "Replacment", "N replaced")

p_iterative_replacement_meandiff_prop <- plot_tampers_prop(simulation_iterative_replacement_meandiff_err, "Replacment", "N replaced")

```

## Bias (Hedges' g relative to μ = 0)

```{r}

simulation_iterative_replacement_meandiff_bias |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

p_iterative_replacement_meandiff_bias <- plot_bias(simulation_iterative_replacement_meandiff_bias, "Replacement")

p_iterative_replacement_meandiff_bias

```

# Combined tables and plots

## Empirical Rejection Rate plots

```{r fig.height=8, fig.width=10}

simulation_iterative_meandiff_err <- bind_rows(
  simulation_iterative_dropping_meandiff_err |> mutate(Method = "Dropping"),
  simulation_iterative_swaps_meandiff_err |> mutate(Method = "Condition swapping"),
  simulation_iterative_duplication_meandiff_err |> mutate(Method = "Duplication"),
  simulation_iterative_fabrication_meandiff_err |> mutate(Method = "Fabrication"),
  simulation_iterative_replacement_meandiff_err |> mutate(Method = "Replacement")
) |>
  mutate(Method = fct_relevel(Method, "Duplication", "Fabrication", "Dropping", "Replacement", "Condition swapping"))

p_meandiff_n <- 
  ggplot(simulation_iterative_meandiff_err, 
         aes(n_total, median, color = Method, fill = Method)) +
  geom_point(position = position_dodge(width = 27)) +
  geom_linerange(aes(ymin = lower, ymax = upper), position = position_dodge(width = 27)) +
  labs(title = "Count of tampered cases needed to reach statistical significance",
       subtitle = "Median and 95% Prediction Intervals in the final tampered dataset",
       x = "Sample size",
       y = "Count") +
  theme_linedraw() +
  scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
  scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2) +
  scale_color_viridis_d(begin = 0, end = 0.85) +
  scale_fill_viridis_d(begin = 0, end = 0.85)

p_meandiff_prop <- 
  ggplot(simulation_iterative_meandiff_err, 
         aes(n_total, prop_median, color = Method, fill = Method)) +
  geom_point(position = position_dodge(width = 27)) +
  geom_linerange(aes(ymin = prop_lower, ymax = prop_upper), position = position_dodge(width = 27)) +
  labs(title = "Proportion of tampered cases needed to reach statistical significance",
       subtitle = "Median and 95% Prediction Intervals in the final tampered dataset",
       x = "Sample size",
       y = "Proportion") +
  theme_linedraw() +
  scale_y_continuous(labels = percent_format(), limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
  scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2) +
  scale_color_viridis_d(begin = 0, end = 0.85) +
  scale_fill_viridis_d(begin = 0, end = 0.85) 

p_meandiff_n +
  p_meandiff_prop +
  plot_layout(ncol = 1, guides = "collect")

ggsave(filename = "../communications/plots/iterative_meandiff_combined.pdf",
       device = "pdf",
       width = 9,
       height = 8)

ggsave(filename = "../communications/plots/iterative_meandiff_combined.png",
       device = "png",
       width = 9,
       height = 8)

```

## Bias plots

Hedges' g of each final, tampered dataset, quantified relative to the true standardised effect of μ = 0. Because the truth is 0, the **signed** median Hedges' g *is* the bias the tampering introduces, and its sign shows the direction each strategy induces (e.g. duplication pushes the opposite way to the others). An absolute |Hedges' g| would be misleading: it is strictly positive even under no systematic bias, so it overstates bias and hides direction.

```{r fig.height=6, fig.width=8}

# bind the raw, unnested per-iteration results across methods (with a Method label)
simulation_iterative_meandiff <- bind_rows(
  simulation_iterative_dropping_meandiff    |> mutate(Method = "Dropping"),
  simulation_iterative_swaps_meandiff       |> mutate(Method = "Condition swapping"),
  simulation_iterative_duplication_meandiff |> mutate(Method = "Duplication"),
  simulation_iterative_fabrication_meandiff |> mutate(Method = "Fabrication"),
  simulation_iterative_replacement_meandiff |> mutate(Method = "Replacement")
) |>
  mutate(Method = fct_relevel(Method, "Duplication", "Fabrication", "Dropping", "Replacement", "Condition swapping"),
         n_total = n_per_condition * 2)

# combined bias summary across methods (uses the raw, unnested simulation results)
simulation_iterative_meandiff_bias <- simulation_iterative_meandiff |>
  group_by(Method, n_total) |>
  summarize(g_median          = median(hedges_g, na.rm = TRUE),
            g_median_mcse     = mcse_median(hedges_g),
            g_lower           = quantile(hedges_g, probs = 0.25, na.rm = TRUE),
            g_upper           = quantile(hedges_g, probs = 0.75, na.rm = TRUE),
            bias_median       = median(bias_mean_diff, na.rm = TRUE),
            bias_median_mcse  = mcse_median(bias_mean_diff),
            .groups = "drop")

p_meandiff_bias <-
  ggplot(simulation_iterative_meandiff_bias,
         aes(n_total, g_median, color = Method, fill = Method)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_point(position = position_dodge(width = 27)) +
  geom_linerange(aes(ymin = g_lower, ymax = g_upper), position = position_dodge(width = 27)) +
  labs(title = "Bias of the tampered effect size relative to the truth (μ = 0)",
       subtitle = "Median and 95% Prediction Intervals in the final tampered dataset",
       x = "Sample size",
       y = "Hedges' g in final dataset") +
  theme_linedraw() +
  scale_y_continuous(breaks = breaks_pretty(n = 8)) +
  scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2) +
  scale_color_viridis_d(begin = 0, end = 0.85) +
  scale_fill_viridis_d(begin = 0, end = 0.85)

p_meandiff_bias

ggsave(filename = "../communications/plots/iterative_meandiff_bias.pdf",
       device = "pdf",
       width = 9,
       height = 5)

ggsave(filename = "../communications/plots/iterative_meandiff_bias.png",
       device = "png",
       width = 9,
       height = 5)

```

```{r}

simulation_iterative_meandiff_bias |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Table

```{r}

simulation_iterative_meandiff_err |>
  relocate(Method, .before = n_total) |>
  dplyr::select(-k, -prop_mean, -prop_mean_mcse, -prop_sd) |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

The `mean` and `sd` columns are reported here alongside the `median` and `mad` (and the 25th/75th percentiles, `lower`/`upper`) so that the skew is visible numerically: because the number of tampers needed is a right-skewed count, the mean sits above the median, and the gap between them is itself a summary of how long the upper tail is. The plots and the headline summaries throughout use the median rather than the mean — see the note under *t-test simulations* for the rationale.

# Running on an HPC cluster

This report is written to run unchanged on a laptop and on a SLURM-managed HPC (e.g. the "Everything" cluster: SLURM + interactive RStudio Server), following the practices in `simulation_template_hpc.qmd`. The cluster scaffolding lives in the *Dependencies* and *Common functions and global variables* chunks:

- **Worker auto-detection.** `parallelly::availableCores()` honours the SLURM allocation, cgroup CPU quotas, and affinity masks, so the parallel plan adapts to the cores actually granted to the session. Set `n_workers_override` to pin a count. On Everything, 4–16 workers is the useful range (little gain beyond that for this workload).
- **BLAS thread pinning.** BLAS/LAPACK is pinned to one thread *before* the workers spawn, so dozens of worker processes do not oversubscribe the node. The matrices here are tiny, so single-threaded BLAS costs nothing.
- **Replication knob.** `run_full` / `on_hpc` auto-detect the environment (any `SLURM_*` variable present ⇒ cluster) and set `n_iterations` accordingly — a fast local smoke test (100) versus the full cluster run (10000). Force it with `run_full <- TRUE`/`FALSE`.
- **Tagged caches.** `cache_path()` tags every `.rds` with `dev`/`full` and the iteration count (e.g. `..._full_k10000.rds`), so local and cluster runs never overwrite each other. Re-rendering reads the cached summary instead of recomputing.
- **Reproducible parallel seeding.** `furrr_options(seed = TRUE)` advances an L'Ecuyer-CMRG stream per task, so results are identical regardless of the worker count; `scheduling = 2` enables dynamic load balancing, and `packages = worker_packages` attaches the needed libraries on each worker.

**Granularity (a deliberate deviation from the template).** The template parallelises over *conditions* and summarises inside each worker, so the master never holds raw replicate rows. These simulations instead parallelise over *replicates* (one task per condition × iteration) and keep the raw per-iteration results, because the distribution histograms and bootstrap-based median MCSEs need the full per-iteration vectors — summarising in the worker would discard exactly the data the report relies on. With many small tasks and `scheduling = 2`, the cost (large-*n* conditions) is spread across workers automatically, which also makes the explicit grid shuffle the template uses (a load-balancing aid for coarse condition-level tasks) unnecessary here.

To run on Everything: request a node and launch RStudio Server inside the allocation, e.g. `salloc --nodes=1 --cpus-per-task=16 --mem=64G --time=08:00:00`; optionally export `FLEXIBLAS_NUM_THREADS=1 OMP_NUM_THREADS=1` in the shell before launching R (belt-and-braces; confirm with `RhpcBLASctl::blas_get_num_procs()`); then render. The full run regenerates the `*_full_k10000.rds` caches.

# Discussion

## Performance of the strategies across conditions

The substantive results are the per-strategy tables and the *Combined tables and plots* section; this discussion summarises how they should be read against the aims. Two patterns are the point and can be checked against the rendered figures:

- **Effort is small and roughly scale-invariant in proportional terms.** Although the absolute tamper count grows with sample size, the *proportion* of the sample that must be altered to reach significance is small and does not increase with N — manufacturing a false positive is no harder, proportionally, in a large study than a small one.
- **The distribution-preserving strategy is both effective and stealthy.** Condition swapping leaves every univariate distribution untouched, so it is the hardest to detect by routine checks, yet reaches significance with few operations. Strategies that add high-leverage points (fabrication, replacement) are the most efficient per operation but the most visible. Because the truth is zero, the median Hedges' *g* at the point significance is reached sits near the critical effect size for the given N — just large enough to clear the threshold and no larger, as the greedy procedure stops on first crossing.

## Conclusions with regard to the aims

The aims are met descriptively by the per-strategy tables (counts, proportions, MCSEs) and bias summaries, and comparatively by the combined plots. The headline conclusion is that the **distance from a null dataset to a "significant" one is short and cheap**, and that the line between practices labelled *p*-hacking and those labelled fabrication corresponds to a difference in keystrokes, not in effort or in the damage done to the evidence.

## Limitations and intended use

- **Proof of principle, not a fraud model.** The strategies are deliberately simple and greedy; real, detection-aware fraud would trade efficiency for plausibility, and the naive versions here can leave detectable signatures such as truncated distributions.
- **No detectability axis.** The study measures effort and bias, not how easily each manipulation would be caught; adding a detectability measure is the most valuable extension.
- **Normal, continuous DGP only.** Likert-type outcomes are not examined; an "alter existing values" strategy distinct from fabrication is left for future work.
- **Single threshold and budget.** Results are conditioned on α = .05 and a 50% tampering cap, both reported transparently.

# References

Chambers, C. (2017). *The seven deadly sins of psychology: A manifesto for reforming the culture of scientific practice.* Princeton University Press.

Morris, T. P., White, I. R., & Crowther, M. J. (2019). Using simulation studies to evaluate statistical methods. *Statistics in Medicine, 38*(11), 2074–2102. https://doi.org/10.1002/sim.8086

Siepe, B. S., Bartoš, F., Morris, T. P., Boulesteix, A.-L., Heck, D. W., & Pawel, S. (2024). Simulation studies for methodological research in psychology: A standardized template for planning, preregistration, and reporting. *Psychological Methods.* https://doi.org/10.1037/met0000695

*[citation to be completed]* Spiess et al. (2024) / the `dfstat`–`dfbeta` influential-observation literature.

# Session info

```{r}

sessionInfo()

```
