---
title: "A compendium of data tampering and fabrication strategies"
subtitle: "And the relative degree to which their minimal application inflates the false positive rate of an independent Student's t-test"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

# TODO

- the dropping and sign changing both alter N participants, whereas the condition swapping swaps N pairs of participants. Alter this to instead do N participants?

# Overview 

## For types of tampering and fabrication

- unwarranted dropping of cases, by N or proportion [done]
- swapping conditions, by N or proportion [done]
- changing/offsetting existing cases - changing sign [done] its a bit arbitrary but could illustrate the point
- duplicating existing cases 
- fabricating new cases - make this less arbitrary by asking how extreme would these cases need to be to produce significant results eg via dfstat

- The simulations thus far use a fixed number of tampers and observe the false positive rate. What about an iterative strategy, where you keep tampering until you get significance, and quantify the necessary number of tampers?

## For additional analyses

- done for t tests but what about regression/correlation/mediation?
- Above consider p-value false positive rate. Add effect size bias. 

## Ethics and optics 

Think about the ethics and options of distributing this weaponisable knowledge. Are there ways to hide detectability in it? 

Add explicit discussion about how arbitrary the line is between p-hacking (scientifically naughty) and fabrication (legally actionable): 

- excluding some participants is usually seen as p-hacking, even if these exclusions are totally unjustifiable and unprincipled, and seem to just be done to produce significant p values.
- but if instead of clicking the delete key on those rows you instead type the other condition name in, suddenly its fabrication and you could lose your job over it.
- the idea that one of these is common and one of these is rare, or one of these is merely iffy and the other is Clearly Unacceptable, risks underestimating the prevalence of the latter.
- as usual, this boils down to a Trolly Problem: is deleting data really ethically different from making up data, if both serve the individuals' self-advancement at the expense of undermining the integrity of the research and the validity of the evidence? 

## Realism of sims

- Perhaps the sim that drops participants feels too obvious for a fraudster, because it can create an obvious truncation of otherwise normally distributed scores. However: 1) many detected cases of fraud have not been very elaborate or smart, and 2) this simulation is a proof of principle, there are of course more elaborate strategies possible which would seek to try to retain the plausibility of data (e.g., removing data beyond a threshold probabilistically).

# Dependencies

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

```{r}

library(tidyr)
library(dplyr)
library(readr)
library(tibble)
library(purrr) 
library(furrr)
library(ggplot2)
library(patchwork)
library(scales)
library(knitr)
library(kableExtra)
library(janitor)
library(ggtext)
# devtools::install_github("ianhussey/simulateR")
library(simulateR)

# set up parallel processing
plan(multisession, workers = availableCores())

options(scipen=999)

dir.create("../data")
dir.create("../communications/plots")

```

# 1a Unwarranted data exclusion (dropping N)

Flexibility in data exclusions has been studied as a p-hacking strategy. However, the strategies considered are usually individually plausible (e.g., ±2 SD from the mean, etc.). Exclusions are sometimes less justifiable or entirely unjustifiable, such as 'qualitative' strategies that seem to remove data points that increase the apparent size of the difference in means between groups (REF: e.g., the response to the Beckers et al. verification report). 

This simulation examines these unjustifiable strategic exclusions.

## Simulation functions

Note: the `alter_data()` function assumes simulated data uses mean_intervention >= mean_control

```{r}

# define data generating function ----
generate_data <- function(n_per_condition,
                          mean_control,
                          mean_intervention,
                          sd_control,
                          sd_intervention) {
  
  data_control <- 
    tibble(condition = "control",
           score = rnorm(n = n_per_condition, mean = mean_control, sd = sd_control))
  
  data_intervention <- 
    tibble(condition = "intervention",
           score = rnorm(n = n_per_condition, mean = mean_intervention, sd = sd_intervention))
  
  data <- bind_rows(data_control,
                    data_intervention) 
  
  return(data)
}


alter_data <- function(data, n_participants_dropped) {
  
  data <- data |>
    rownames_to_column(var = "rowname") |>
    mutate(altered = FALSE)
  
  # identify the N smallest scores in the intervention group
  smallest_intervention <- data |>
    filter(condition == "intervention") |>
    mutate(z_score = (score - mean(score)) / sd(score),
           abs_z_score = abs(z_score)) |>
    arrange(z_score) |> # assumes simulated data uses mean_intervention >= mean_control
    slice_head(n = n_participants_dropped) # oversample n dropped
  
  # identify the N largest scores in the control group
  largest_control <- data |>
    filter(condition == "control") |>
    mutate(z_score = score / sd(score),
           abs_z_score = abs(z_score)) |>
    arrange(desc(z_score)) |> # assumes simulated data uses mean_intervention >= mean_control
    slice_head(n = n_participants_dropped) # oversample n dropped
  
  extreme_scores <- 
    bind_rows(smallest_intervention, 
              largest_control) |>
    arrange(desc(abs_z_score)) |>
    slice_head(n = n_participants_dropped) |> # correctly sample n dropped
    pull(rowname)
  
  # set 'altered' to TRUE for these rows, and score_altered to NA
  data <- data |>
    mutate(altered = ifelse(rowname %in% extreme_scores, TRUE, FALSE),
           score = ifelse(altered, NA, score)) |>
    select(-rowname)
  
  return(data)
}


# define data analysis function ----
analyse_data <- function(data) {
  
  res_t_test <- t.test(formula = score ~ condition, 
                       data = data,
                       var.equal = TRUE,
                       alternative = "two.sided")
  
  res <- tibble(p = res_t_test$p.value)
  
  return(res)
}

```

## Run simulation

```{r}

# define experiment parameters ----
experiment_parameters_grid <- expand_grid(
  n_per_condition = c(10, seq(from = 25, to = 200, by = 25)),
  mean_control = 0,
  mean_intervention = 0, 
  sd_control = 1,
  sd_intervention = 1,
  #n_participants_dropped_per_condition = c(1, 2, 3, 4, 5),
  n_participants_dropped = 1:10,
  iteration = 1:10000
)

if(file.exists("../data/simulation_fixed_exclusions_n.rds")){
  
  simulation_fixed_exclusions_n <- read_rds("../data/simulation_fixed_exclusions_n.rds")
  
} else {
  
  # set the seed ----
  # for the pseudo random number generator to make results reproducible
  set.seed(123)
  
  # run simulation ----
  simulation_fixed_exclusions_n <- 
    # using the experiment parameters
    experiment_parameters_grid |>
    
    # generate data using the data generating function and the parameters relevant to data generation
    mutate(generated_data = future_pmap(list(n_per_condition,
                                             mean_control,
                                             mean_intervention,
                                             sd_control,
                                             sd_intervention),
                                        generate_data,
                                        .progress = TRUE,
                                        .options = furrr_options(seed = TRUE))) |>
    
    # generate data using the data generating function and the parameters relevant to data generation
    mutate(altered_data = future_pmap(list(generated_data,
                                           n_participants_dropped),
                                      alter_data,
                                      .progress = TRUE,
                                      .options = furrr_options(seed = TRUE))) |>
    
    # apply the analysis function to the generated data using the parameters relevant to analysis
    mutate(analysis_results_original_data = future_pmap(list(generated_data),
                                                        analyse_data,
                                                        .progress = TRUE,
                                                        .options = furrr_options(seed = TRUE)),
           analysis_results_altered_data = future_pmap(list(altered_data),
                                                       analyse_data,
                                                       .progress = TRUE,
                                                       .options = furrr_options(seed = TRUE)))
  
  write_rds(simulation_fixed_exclusions_n,
            "../data/simulation_fixed_exclusions_n.rds", 
            compress = "gz")
  
  simulation_fixed_exclusions_n |>
    select(-generated_data, -altered_data) |>
    write_rds("../data/simulation_fixed_exclusions_n_nodata.rds",
              compress = "gz")
}

```

## Illustrate drops

```{r}

set.seed(42)

dat_illustrate <- simulation_fixed_exclusions_n |>
  rownames_to_column() |>
  unnest(analysis_results_original_data) |>
  rename(p_original = p) |>
  unnest(analysis_results_altered_data) |>
  rename(p_altered = p) |>
  filter(n_participants_dropped == 1 &
           p_original > .05 & p_altered < .05) |>
  mutate(diff = p_original - p_altered) |>
  arrange(desc(diff))

# dat_illustrate_most_extreme_rowname <- dat_illustrate |>
#   slice_head(n = 1) |>
#   pull(rowname)

random_row <- dat_illustrate |>
  slice_sample(n = 1) |>
  pull(rowname)

full_join(
  dat_illustrate |> 
    filter(rowname == random_row) |> 
    select(generated_data) |>
    unnest(generated_data) |>
    rownames_to_column() |>
    rename(score_original = score),
  dat_illustrate |> 
    filter(rowname == random_row) |> 
    select(altered_data) |>
    unnest(altered_data) |>
    rownames_to_column() |>
    rename(score_altered = score),
  by = c("rowname", "condition")
) |>
  ggplot(aes(condition, score_original, color = altered)) +
  geom_point(position = position_jitter(width = 0.1)) +
  scale_color_viridis_d(begin = 0.3, end = 0.7) +
  theme_linedraw() +
  ylab("Score")

dat_illustrate |>
  filter(rowname == random_row) |>
  select(p_original, p_altered) |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 3) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Summary table 

```{r}

# summarise simulation results over the iterations ----
simulation_fixed_exclusions_n_summary <- simulation_fixed_exclusions_n |>
  # convert `analysis_results` nested-data-frame column to regular columns in the df. in this case, the p value.
  unnest(analysis_results_original_data) |>
  rename(p_original_data = p) |>
  unnest(analysis_results_altered_data) |>
  rename(p_altered_data = p) |>
  # summarize across iterations
  group_by(n_per_condition, n_participants_dropped) |>
  summarize(original_data_false_positive_rate = mean(p_original_data < .05),
            altered_data_false_positive_rate = mean(p_altered_data < .05)) |>
  ungroup() |>
  mutate(proportion_dropped = n_participants_dropped/n_per_condition)

simulation_fixed_exclusions_n_summary |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

# simulation_fixed_exclusions_n_summary |>
#   filter(n_participants_dropped == 1) |>
#   mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
#   kable() |>
#   kable_classic(full_width = FALSE)
# 
# simulation_fixed_exclusions_n_summary |>
#   filter(n_participants_dropped == 5) |>
#   mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
#   kable() |>
#   kable_classic(full_width = FALSE)

```

## Plot

```{r}

p_fixed_exclusions_n <- ggplot(simulation_fixed_exclusions_n_summary, 
                               aes(n_participants_dropped, 
                                   altered_data_false_positive_rate, 
                                   color = as.factor(n_per_condition*2))) + 
  geom_hline(yintercept = 0.05, linetype = "dotted") +
  geom_line() +
  #geom_smooth(method = "lm", alpha = 0.1) +
  geom_point() +
  scale_color_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  theme_linedraw() +
  labs(color = "N",
       x = "N participants excluded",
       y = "False positive rate") +
  scale_y_continuous(limits = c(0,1), breaks = c(0, 0.05, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1)) +
  scale_x_continuous(breaks = breaks_pretty(n = 10)) +
  guides(color = guide_legend(reverse = TRUE))

p_fixed_exclusions_n

# ggsave(filename = "../communications/plots/fixed_exclusions_n.pdf",
#        device = "pdf",
#        width = 6,
#        height = 5)

```

# 1b Unwarranted data exclusion (dropping %)

This is similar to the last simulation but instead of excluding a fixed N of participants regardless of total sample size I exclude a fixed percentage of participants.

## Simulation functions

Note: the `alter_data()` function assumes simulated data uses mean_intervention >= mean_control

```{r}

# define data generating function ----
generate_data <- function(n_per_condition,
                          mean_control,
                          mean_intervention,
                          sd_control,
                          sd_intervention) {
  
  data_control <- 
    tibble(condition = "control",
           score = rnorm(n = n_per_condition, mean = mean_control, sd = sd_control))
  
  data_intervention <- 
    tibble(condition = "intervention",
           score = rnorm(n = n_per_condition, mean = mean_intervention, sd = sd_intervention))
  
  data <- bind_rows(data_control,
                    data_intervention) 
  
  return(data)
}


alter_data <- function(data, n_participants_dropped) {
  
  data <- data |>
    rownames_to_column(var = "rowname") |>
    mutate(altered = FALSE)
  
  # identify the N smallest scores in the intervention group
  smallest_intervention <- data |>
    filter(condition == "intervention") |>
    mutate(z_score = score / sd(score),
           abs_z_score = abs(z_score)) |>
    arrange(z_score) |> # assumes simulated data uses mean_intervention >= mean_control
    slice_head(n = n_participants_dropped) # oversample n dropped
  
  # identify the N largest scores in the control group
  largest_control <- data |>
    filter(condition == "control") |>
    mutate(z_score = (score - mean(score)) / sd(score),
           abs_z_score = abs(z_score)) |>
    arrange(desc(z_score)) |> # assumes simulated data uses mean_intervention >= mean_control
    slice_head(n = n_participants_dropped) # oversample n dropped
  
  extreme_scores <- 
    bind_rows(smallest_intervention, 
              largest_control) |>
    arrange(desc(abs_z_score)) |>
    slice_head(n = n_participants_dropped) |> # correctly sample n dropped
    pull(rowname)
  
  # set 'altered' to TRUE for these rows, and score_altered to NA
  data <- data |>
    mutate(altered = ifelse(rowname %in% extreme_scores, TRUE, FALSE),
           score = ifelse(altered, NA, score)) |>
    select(-rowname)
  
  return(data)
}


# define data analysis function ----
analyse_data <- function(data) {
  
  res_t_test <- t.test(formula = score ~ condition, 
                       data = data,
                       var.equal = TRUE,
                       alternative = "two.sided")
  
  res <- tibble(p = res_t_test$p.value)
  
  return(res)
}

```

## Run simulation

```{r}

# define experiment parameters ----
experiment_parameters_grid <- expand_grid(
  n_per_condition = c(10, seq(from = 25, to = 200, by = 25)),
  mean_control = 0,
  mean_intervention = 0, 
  sd_control = 1,
  sd_intervention = 1,
  proportion_participants_dropped = c(.01, .025, .05, .10, .15, .10, .15, .20, .25, .30, .35),
  iteration = 1:10000
)

if(file.exists("../data/simulation_fixed_exclusions_percent.rds")){
  
  simulation_fixed_exclusions_percent <- read_rds("../data/simulation_fixed_exclusions_percent.rds")
  
} else {
  
  # set the seed ----
  # for the pseudo random number generator to make results reproducible
  set.seed(123)
  
  # run simulation ----
  simulation_fixed_exclusions_percent <- 
    # using the experiment parameters
    experiment_parameters_grid |>
    
    # generate data using the data generating function and the parameters relevant to data generation
    mutate(generated_data = future_pmap(list(n_per_condition,
                                             mean_control,
                                             mean_intervention,
                                             sd_control,
                                             sd_intervention),
                                        generate_data,
                                        .progress = TRUE,
                                        .options = furrr_options(seed = TRUE))) |>
    
    mutate(n_participants_dropped = janitor::round_half_up(proportion_participants_dropped * n_per_condition*2)) |>
    
    # generate data using the data generating function and the parameters relevant to data generation
    mutate(altered_data = future_pmap(list(generated_data,
                                           n_participants_dropped),
                                      alter_data,
                                      .progress = TRUE,
                                      .options = furrr_options(seed = TRUE))) |>
    
    # apply the analysis function to the generated data using the parameters relevant to analysis
    mutate(analysis_results_original_data = future_pmap(list(generated_data),
                                                        analyse_data,
                                                        .progress = TRUE,
                                                        .options = furrr_options(seed = TRUE)),
           analysis_results_altered_data = future_pmap(list(altered_data),
                                                       analyse_data,
                                                       .progress = TRUE,
                                                       .options = furrr_options(seed = TRUE)))
  
  write_rds(simulation_fixed_exclusions_percent,
            "../data/simulation_fixed_exclusions_percent.rds", 
            compress = "gz")
  
  simulation_fixed_exclusions_percent |>
    select(-generated_data, -altered_data) |>
    write_rds("../data/simulation_fixed_exclusions_percent_nodata.rds",
              compress = "gz")
}

```

## Illustrate drops

```{r}

set.seed(42)

dat_illustrate <- simulation_fixed_exclusions_percent |>
  rownames_to_column() |>
  unnest(analysis_results_original_data) |>
  rename(p_original = p) |>
  unnest(analysis_results_altered_data) |>
  rename(p_altered = p) |>
  filter(p_original > .05 & p_altered < .05) |>
  mutate(diff = p_original - p_altered) |>
  arrange(desc(diff))

dat_illustrate |>
  slice_head(n = 10) |>
  select(n_per_condition, proportion_participants_dropped, p_original, p_altered, diff) |>
  kable() |>
  kable_classic(full_width = FALSE)

random_row <- dat_illustrate |>
  slice_sample(n = 1) |>
  pull(rowname)

full_join(
  dat_illustrate |> 
    filter(rowname == random_row) |> 
    select(generated_data) |>
    unnest(generated_data) |>
    rownames_to_column() |>
    rename(score_original = score),
  dat_illustrate |> 
    filter(rowname == random_row) |> 
    select(altered_data) |>
    unnest(altered_data) |>
    rownames_to_column() |>
    rename(score_altered = score),
  by = c("rowname", "condition")
) |>
  ggplot(aes(condition, score_original, color = altered)) +
  geom_point(position = position_jitter(width = 0.1)) +
  scale_color_viridis_d(begin = 0.3, end = 0.7) +
  theme_linedraw() +
  ylab("Score")

dat_illustrate |>
  filter(rowname == random_row) |>
  select(p_original, p_altered) |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 3) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Summary table 

```{r}

# summarise simulation results over the iterations ----
simulation_fixed_exclusions_percent_summary <-
  simulation_fixed_exclusions_percent |>
  # convert `analysis_results` nested-data-frame column to regular columns in the df. in this case, the p value.
  unnest(analysis_results_original_data) |>
  rename(p_original_data = p) |>
  unnest(analysis_results_altered_data) |>
  rename(p_altered_data = p) |>
  # summarize across iterations
  group_by(n_per_condition, proportion_participants_dropped) |>
  summarize(original_data_false_positive_rate = mean(p_original_data < .05),
            altered_data_false_positive_rate = mean(p_altered_data < .05)) |>
  ungroup() 

simulation_fixed_exclusions_percent_summary |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

# simulation_fixed_exclusions_percent_summary |>
#   filter(proportion_participants_dropped == .1) |>
#   mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
#   kable() |>
#   kable_classic(full_width = FALSE)
# 
# simulation_fixed_exclusions_percent_summary |>
#   filter(proportion_participants_dropped == .35) |>
#   mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
#   kable() |>
#   kable_classic(full_width = FALSE)

```

## Plot

```{r}

p_fixed_exclusions_percent <- ggplot(simulation_fixed_exclusions_percent_summary,
                                     aes(proportion_participants_dropped, 
                                         altered_data_false_positive_rate, 
                                         color = as.factor(n_per_condition*2))) + 
  geom_hline(yintercept = 0.05, linetype = "dotted") +
  geom_line() +
  geom_point() +
  scale_color_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  theme_linedraw() +
  labs(color = "N",
       x = "% participants excluded",
       y = "False positive rate") +
  scale_y_continuous(limits = c(0,1), breaks = c(0, 0.05, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1)) +
  scale_x_continuous(labels = percent_format(), breaks = c(.01, .025, .05, .10, .15, .20, .25, .30, .35)) +
  guides(color = guide_legend(reverse = TRUE))

p_fixed_exclusions_percent

# ggsave(filename = "../communications/plots/fixed_exclude_percent.pdf",
#        device = "pdf",
#        width = 6,
#        height = 5)

```

# 2a Condition swapping participants (N participants)

## Simulation functions

Involves swapping the conditions of individual participants, but doing half of swaps from each group so as to retain group balance

Note: the `alter_data()` function assumes simulated data uses mean_intervention >= mean_control

```{r}

# define data generating function ----
generate_data <- function(n_per_condition,
                          mean_control,
                          mean_intervention,
                          sd_control,
                          sd_intervention) {
  
  data_control <- 
    tibble(condition = "control",
           score = rnorm(n = n_per_condition, mean = mean_control, sd = sd_control))
  
  data_intervention <- 
    tibble(condition = "intervention",
           score = rnorm(n = n_per_condition, mean = mean_intervention, sd = sd_intervention))
  
  data <- bind_rows(data_control,
                    data_intervention) 
  
  return(data)
}

alter_data <- function(data, n_participants_swapped) {
  data <- data |>
    rownames_to_column(var = "rowname") |>
    mutate(altered = FALSE)
  
  # Compute z-score within full sample
  data <- data |>
    mutate(z_score = (score - mean(score)) / sd(score))

  # Split groupwise
  control <- data |> filter(condition == "control")
  intervention <- data |> filter(condition == "intervention")
  
  # Identify most extreme low scores in intervention group
  to_swap_from_intervention <- intervention |>
    arrange(z_score) |>  # lowest scores
    slice_head(n = floor(n_participants_swapped / 2)) |>
    pull(rowname)
  
  # Identify most extreme high scores in control group
  to_swap_from_control <- control |>
    arrange(desc(z_score)) |>  # highest scores
    slice_head(n = ceiling(n_participants_swapped / 2)) |>
    pull(rowname)
  
  swap_rows <- c(to_swap_from_control, to_swap_from_intervention)

  # Apply swap
  data <- data |>
    mutate(
      condition = case_when(
        rowname %in% to_swap_from_control ~ "intervention",
        rowname %in% to_swap_from_intervention ~ "control",
        TRUE ~ condition
      ),
      altered = if_else(rowname %in% swap_rows, TRUE, FALSE)
    ) |>
    select(-rowname, -z_score)
  
  return(data)
}


# define data analysis function ----
analyse_data <- function(data) {
  
  res_t_test <- t.test(formula = score ~ condition, 
                       data = data,
                       var.equal = TRUE,
                       alternative = "two.sided")
  
  res <- tibble(p = res_t_test$p.value)
  
  return(res)
}

```

## Run simulation

```{r}

# define experiment parameters ----
experiment_parameters_grid <- expand_grid(
  n_per_condition = c(10, seq(from = 25, to = 200, by = 25)),
  mean_control = 0,
  mean_intervention = 0, 
  sd_control = 1,
  sd_intervention = 1,
  n_participants_swapped = 1:10,
  iteration = 1:10000
)

if(file.exists("../data/simulation_fixed_swaps_n.rds")){
  
  simulation_fixed_swaps_n <- read_rds("../data/simulation_fixed_swaps_n.rds")
  
} else {
  
  # set the seed ----
  # for the pseudo random number generator to make results reproducible
  set.seed(123)
  
  # run simulation ----
  simulation_fixed_swaps_n <- 
    # using the experiment parameters
    experiment_parameters_grid |>
    
    # generate data using the data generating function and the parameters relevant to data generation
    mutate(generated_data = future_pmap(list(n_per_condition,
                                             mean_control,
                                             mean_intervention,
                                             sd_control,
                                             sd_intervention),
                                        generate_data,
                                        .progress = TRUE,
                                        .options = furrr_options(seed = TRUE))) |>
    
    # generate data using the data generating function and the parameters relevant to data generation
    mutate(altered_data = future_pmap(list(generated_data,
                                           n_participants_swapped),
                                      alter_data,
                                      .progress = TRUE,
                                      .options = furrr_options(seed = TRUE))) |>
    
    # apply the analysis function to the generated data using the parameters relevant to analysis
    mutate(analysis_results_original_data = future_pmap(list(generated_data),
                                                        analyse_data,
                                                        .progress = TRUE,
                                                        .options = furrr_options(seed = TRUE)),
           analysis_results_altered_data = future_pmap(list(altered_data),
                                                       analyse_data,
                                                       .progress = TRUE,
                                                       .options = furrr_options(seed = TRUE)))
  
  write_rds(simulation_fixed_swaps_n, 
            "../data/simulation_fixed_swaps_n.rds", 
            compress = "gz")
  
  simulation_fixed_swaps_n |>
    select(-generated_data, -altered_data) |>
    write_rds("../data/simulation_fixed_swaps_n_nodata.rds",
              compress = "gz")
}

```

## Illustrate swaps between conditions 

\TODO

```{r}



```

## Summary table 

```{r}

# summarise simulation results over the iterations ----
simulation_fixed_swaps_n_summary <- simulation_fixed_swaps_n |>
  # convert `analysis_results` nested-data-frame column to regular columns in the df. in this case, the p value.
  unnest(analysis_results_original_data) |>
  rename(p_original_data = p) |>
  unnest(analysis_results_altered_data) |>
  rename(p_altered_data = p) |>
  # summarize across iterations
  group_by(n_per_condition, n_participants_swapped) |>
  summarize(original_data_false_positive_rate = mean(p_original_data < .05),
            altered_data_false_positive_rate = mean(p_altered_data < .05)) |>
  ungroup()

simulation_fixed_swaps_n_summary |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

# simulation_fixed_swaps_n_summary |>
#   filter(n_participants_swapped == 1) |>
#   mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
#   kable() |>
#   kable_classic(full_width = FALSE)
# 
# simulation_fixed_swaps_n_summary |>
#   filter(n_participants_swapped == 5) |>
#   mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
#   kable() |>
#   kable_classic(full_width = FALSE)

```

## Plot

```{r}

p_fixed_swaps_n <- ggplot(simulation_fixed_swaps_n_summary,
                          aes(n_participants_swapped, 
                              altered_data_false_positive_rate, 
                              color = as.factor(n_per_condition*2))) + 
  geom_hline(yintercept = 0.05, linetype = "dotted") +
  geom_line() +
  geom_point() +
  scale_color_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  theme_linedraw() +
  labs(color = "N",
       x = "N participants condition-swapped",
       y = "False positive rate") +
  scale_y_continuous(limits = c(0,1), breaks = c(0, 0.05, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1)) +
  scale_x_continuous(breaks = breaks_pretty(n = 10)) +
  guides(color = guide_legend(reverse = TRUE))

p_fixed_swaps_n

# ggsave(filename = "../communications/plots/fixed_swaps_n.pdf",
#        device = "pdf",
#        width = 6,
#        height = 5)

```

# 2b Condition swapping participants (% participants)

This is similar to the last simulation but instead of excluding a fixed N of participants regardless of total sample size I exclude a fixed percentage of participants.

## Simulation functions

Involves swapping the conditions of individual participants, but doing half of swaps from each group so as to retain group balance

Note: the `alter_data()` function assumes simulated data uses mean_intervention >= mean_control

```{r}

# define data generating function ----
generate_data <- function(n_per_condition,
                          mean_control,
                          mean_intervention,
                          sd_control,
                          sd_intervention) {
  
  data_control <- 
    tibble(condition = "control",
           score = rnorm(n = n_per_condition, mean = mean_control, sd = sd_control))
  
  data_intervention <- 
    tibble(condition = "intervention",
           score = rnorm(n = n_per_condition, mean = mean_intervention, sd = sd_intervention))
  
  data <- bind_rows(data_control,
                    data_intervention) 
  
  return(data)
}

alter_data <- function(data, n_participants_swapped) {
  data <- data |>
    rownames_to_column(var = "rowname") |>
    mutate(altered = FALSE)
  
  # Compute z-score within full sample
  data <- data |>
    mutate(z_score = (score - mean(score)) / sd(score))

  # Split groupwise
  control <- data |> filter(condition == "control")
  intervention <- data |> filter(condition == "intervention")
  
  # Identify most extreme low scores in intervention group
  to_swap_from_intervention <- intervention |>
    arrange(z_score) |>  # lowest scores
    slice_head(n = floor(n_participants_swapped / 2)) |>
    pull(rowname)
  
  # Identify most extreme high scores in control group
  to_swap_from_control <- control |>
    arrange(desc(z_score)) |>  # highest scores
    slice_head(n = ceiling(n_participants_swapped / 2)) |>
    pull(rowname)
  
  swap_rows <- c(to_swap_from_control, to_swap_from_intervention)

  # Apply swap
  data <- data |>
    mutate(
      condition = case_when(
        rowname %in% to_swap_from_control ~ "intervention",
        rowname %in% to_swap_from_intervention ~ "control",
        TRUE ~ condition
      ),
      altered = if_else(rowname %in% swap_rows, TRUE, FALSE)
    ) |>
    select(-rowname, -z_score)
  
  return(data)
}


# define data analysis function ----
analyse_data <- function(data) {
  
  res_t_test <- t.test(formula = score ~ condition, 
                       data = data,
                       var.equal = TRUE,
                       alternative = "two.sided")
  
  res <- tibble(p = res_t_test$p.value)
  
  return(res)
}

```

## Run simulation

```{r}

# define experiment parameters ----
experiment_parameters_grid <- expand_grid(
  n_per_condition = c(10, seq(from = 25, to = 200, by = 25)),
  mean_control = 0,
  mean_intervention = 0, 
  sd_control = 1,
  sd_intervention = 1,
  proportion_participants_swapped = c(.01, .025, .05, .10, .15, .10, .15, .20, .25, .30, .35),
  iteration = 1:10000
)

if(file.exists("../data/simulation_fixed_swaps_percent.rds")){
  
  simulation_fixed_swaps_percent <- read_rds("../data/simulation_fixed_swaps_percent.rds")
  
} else {
  
  # set the seed ----
  # for the pseudo random number generator to make results reproducible
  set.seed(123)
  
  # run simulation ----
  simulation_fixed_swaps_percent <- 
    # using the experiment parameters
    experiment_parameters_grid |>
    
    # generate data using the data generating function and the parameters relevant to data generation
    mutate(generated_data = future_pmap(list(n_per_condition,
                                             mean_control,
                                             mean_intervention,
                                             sd_control,
                                             sd_intervention),
                                        generate_data,
                                        .progress = TRUE,
                                        .options = furrr_options(seed = TRUE))) |>
    
    mutate(n_participants_swapped = janitor::round_half_up(proportion_participants_swapped * n_per_condition * 2)) |>
    
    # generate data using the data generating function and the parameters relevant to data generation
    mutate(altered_data = future_pmap(list(generated_data,
                                           n_participants_swapped),
                                      alter_data,
                                      .progress = TRUE,
                                      .options = furrr_options(seed = TRUE))) |>
    
    # apply the analysis function to the generated data using the parameters relevant to analysis
    mutate(analysis_results_original_data = future_pmap(list(generated_data),
                                                        analyse_data,
                                                        .progress = TRUE,
                                                        .options = furrr_options(seed = TRUE)),
           analysis_results_altered_data = future_pmap(list(altered_data),
                                                       analyse_data,
                                                       .progress = TRUE,
                                                       .options = furrr_options(seed = TRUE)))
  
  write_rds(simulation_fixed_swaps_percent, 
            "../data/simulation_fixed_swaps_percent.rds", 
            compress = "gz")
  
  simulation_fixed_swaps_percent |>
    select(-generated_data, -altered_data) |>
    write_rds("../data/simulation_fixed_swaps_percent_nodata.rds",
              compress = "gz")
}

```

## Illustrate swaps between conditions

\TODO

```{r}



```

## Summary table 

```{r}

# summarise simulation results over the iterations ----
simulation_fixed_swaps_percent_summary <- simulation_fixed_swaps_percent |>
  # convert `analysis_results` nested-data-frame column to regular columns in the df. in this case, the p value.
  unnest(analysis_results_original_data) |>
  rename(p_original_data = p) |>
  unnest(analysis_results_altered_data) |>
  rename(p_altered_data = p) |>
  # summarize across iterations
  group_by(n_per_condition, proportion_participants_swapped) |>
  summarize(original_data_false_positive_rate = mean(p_original_data < .05),
            altered_data_false_positive_rate = mean(p_altered_data < .05)) |>
  ungroup()

simulation_fixed_swaps_percent_summary |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

# simulation_fixed_swaps_percent_summary |>
#   filter(proportion_participants_swapped == .025) |>
#   mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
#   kable() |>
#   kable_classic(full_width = FALSE)
# 
# simulation_fixed_swaps_percent_summary |>
#   filter(proportion_participants_swapped == .15) |>
#   mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
#   kable() |>
#   kable_classic(full_width = FALSE)

```

## Plot

```{r}

p_fixed_swaps_percent <- ggplot(simulation_fixed_swaps_percent_summary,
                                aes(proportion_participants_swapped, 
                                    altered_data_false_positive_rate, 
                                    color = as.factor(n_per_condition*2))) + 
  geom_hline(yintercept = 0.05, linetype = "dotted") +
  geom_line() +
  geom_point() +
  scale_color_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  theme_linedraw() +
  labs(color = "N",
       x = "% participants condition-swapped",
       y = "False positive rate") +
  scale_y_continuous(limits = c(0,1), breaks = c(0, 0.05, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1)) +
  scale_x_continuous(labels = percent_format(), breaks = c(.01, .025, .05, .10, .15, .20, .25, .30, .35)) +
  guides(color = guide_legend(reverse = TRUE))

p_fixed_swaps_percent

# ggsave(filename = "../communications/plots/fixed_swaps_percent.pdf",
#        device = "pdf",
#        width = 6,
#        height = 5)

```

# 3a Change sign (N)

Flexibility in data exclusions has been studied as a p-hacking strategy. However, the strategies considered are usually individually plausible (e.g., ±2 SD from the mean, etc.). Exclusions are sometimes less justifiable or entirely unjustifiable, such as 'qualitative' strategies that seem to remove data points that increase the apparent size of the difference in means between groups (REF: e.g., the response to the Beckers et al. verification report). 

This simulation examines these unjustifiable strategic exclusions.

## Simulation functions

Note: the `alter_data()` function assumes simulated data uses mean_intervention >= mean_control

```{r}

# define data generating function ----
generate_data <- function(n_per_condition,
                          mean_control,
                          mean_intervention,
                          sd_control,
                          sd_intervention) {
  
  data_control <- 
    tibble(condition = "control",
           score = rnorm(n = n_per_condition, mean = mean_control, sd = sd_control))
  
  data_intervention <- 
    tibble(condition = "intervention",
           score = rnorm(n = n_per_condition, mean = mean_intervention, sd = sd_intervention))
  
  data <- bind_rows(data_control,
                    data_intervention) 
  
  return(data)
}


alter_data <- function(data, n_participants_signchanged) {
  
  data <- data |>
    rownames_to_column(var = "rowname") |>
    mutate(altered = FALSE)
  
  # identify the N smallest scores in the intervention group
  smallest_intervention <- data |>
    filter(condition == "intervention") |>
    mutate(z_score = (score - mean(score)) / sd(score),
           abs_z_score = abs(z_score)) |>
    arrange(z_score) |> # assumes simulated data uses mean_intervention >= mean_control
    slice_head(n = n_participants_signchanged) # oversample n signchanged
  
  # identify the N largest scores in the control group
  largest_control <- data |>
    filter(condition == "control") |>
    mutate(z_score = score / sd(score),
           abs_z_score = abs(z_score)) |>
    arrange(desc(z_score)) |> # assumes simulated data uses mean_intervention >= mean_control
    slice_head(n = n_participants_signchanged) # oversample n signchanged
  
  extreme_scores <- 
    bind_rows(smallest_intervention, 
              largest_control) |>
    arrange(desc(abs_z_score)) |>
    slice_head(n = n_participants_signchanged) |> # correctly sample n signchanged
    pull(rowname)
  
  # set 'altered' to TRUE for these rows, and score_altered to NA
  data <- data |>
    mutate(altered = ifelse(rowname %in% extreme_scores, TRUE, FALSE),
           score = ifelse(altered, score*-1, score)) |>
    select(-rowname)
  
  return(data)
}


# define data analysis function ----
analyse_data <- function(data) {
  
  res_t_test <- t.test(formula = score ~ condition, 
                       data = data,
                       var.equal = TRUE,
                       alternative = "two.sided")
  
  res <- tibble(p = res_t_test$p.value)
  
  return(res)
}

```

## Run simulation

```{r}

# define experiment parameters ----
experiment_parameters_grid <- expand_grid(
  n_per_condition = c(10, seq(from = 25, to = 200, by = 25)),
  mean_control = 0,
  mean_intervention = 0, 
  sd_control = 1,
  sd_intervention = 1,
  n_participants_signchanged = 1:10,
  iteration = 1:10000
)

if(file.exists("../data/simulation_fixed_changesigns_n.rds")){
  
  simulation_fixed_changesigns_n <- read_rds("../data/simulation_fixed_changesigns_n.rds")
  
} else {
  
  # set the seed ----
  # for the pseudo random number generator to make results reproducible
  set.seed(123)
  
  # run simulation ----
  simulation_fixed_changesigns_n <- 
    # using the experiment parameters
    experiment_parameters_grid |>
    
    # generate data using the data generating function and the parameters relevant to data generation
    mutate(generated_data = future_pmap(list(n_per_condition,
                                             mean_control,
                                             mean_intervention,
                                             sd_control,
                                             sd_intervention),
                                        generate_data,
                                        .progress = TRUE,
                                        .options = furrr_options(seed = TRUE))) |>
    
    # generate data using the data generating function and the parameters relevant to data generation
    mutate(altered_data = future_pmap(list(generated_data,
                                           n_participants_signchanged),
                                      alter_data,
                                      .progress = TRUE,
                                      .options = furrr_options(seed = TRUE))) |>
    
    # apply the analysis function to the generated data using the parameters relevant to analysis
    mutate(analysis_results_original_data = future_pmap(list(generated_data),
                                                        analyse_data,
                                                        .progress = TRUE,
                                                        .options = furrr_options(seed = TRUE)),
           analysis_results_altered_data = future_pmap(list(altered_data),
                                                       analyse_data,
                                                       .progress = TRUE,
                                                       .options = furrr_options(seed = TRUE)))
  
  write_rds(simulation_fixed_changesigns_n,
            "../data/simulation_fixed_changesigns_n.rds", 
            compress = "gz")
  
  simulation_fixed_changesigns_n |>
    select(-generated_data, -altered_data) |>
    write_rds("../data/simulation_fixed_changesigns_n_nodata.rds",
              compress = "gz")
}

```

## Illustrate drops

```{r}

set.seed(42)

dat_illustrate <- simulation_fixed_changesigns_n |>
  rownames_to_column() |>
  unnest(analysis_results_original_data) |>
  rename(p_original = p) |>
  unnest(analysis_results_altered_data) |>
  rename(p_altered = p) |>
  filter(n_participants_signchanged == 1 &
           p_original > .05 & p_altered < .05) |>
  mutate(diff = p_original - p_altered) |>
  arrange(desc(diff))

# dat_illustrate_most_extreme_rowname <- dat_illustrate |>
#   slice_head(n = 1) |>
#   pull(rowname)

random_row <- dat_illustrate |>
  slice_sample(n = 1) |>
  pull(rowname)

full_join(
  dat_illustrate |> 
    filter(rowname == random_row) |> 
    select(generated_data) |>
    unnest(generated_data) |>
    rownames_to_column() |>
    rename(score_original = score),
  dat_illustrate |> 
    filter(rowname == random_row) |> 
    select(altered_data) |>
    unnest(altered_data) |>
    rownames_to_column() |>
    rename(score_altered = score),
  by = c("rowname", "condition")
) |>
  ggplot(aes(condition, score_original, color = altered)) +
  geom_point(position = position_jitter(width = 0.1)) +
  scale_color_viridis_d(begin = 0.3, end = 0.7) +
  theme_linedraw() +
  ylab("Score")

dat_illustrate |>
  filter(rowname == random_row) |>
  select(p_original, p_altered) |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 3) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Summary table 

```{r}

# summarise simulation results over the iterations ----
simulation_fixed_changesigns_n_summary <- simulation_fixed_changesigns_n |>
  # convert `analysis_results` nested-data-frame column to regular columns in the df. in this case, the p value.
  unnest(analysis_results_original_data) |>
  rename(p_original_data = p) |>
  unnest(analysis_results_altered_data) |>
  rename(p_altered_data = p) |>
  # summarize across iterations
  group_by(n_per_condition, n_participants_signchanged) |>
  summarize(original_data_false_positive_rate = mean(p_original_data < .05),
            altered_data_false_positive_rate = mean(p_altered_data < .05)) |>
  ungroup() |>
  mutate(proportion_tampered = n_participants_signchanged/n_per_condition)

simulation_fixed_changesigns_n_summary |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

# simulation_fixed_changesigns_n_summary |>
#   filter(n_participants_signchanged == 1) |>
#   mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
#   kable() |>
#   kable_classic(full_width = FALSE)
# 
# simulation_fixed_changesigns_n_summary |>
#   filter(n_participants_signchanged == 5) |>
#   mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
#   kable() |>
#   kable_classic(full_width = FALSE)

```

## Plot

```{r}

p_fixed_changesigns_n <- ggplot(simulation_fixed_changesigns_n_summary, 
                                aes(n_participants_signchanged, 
                                    altered_data_false_positive_rate, 
                                    color = as.factor(n_per_condition*2))) + 
  geom_hline(yintercept = 0.05, linetype = "dotted") +
  geom_line() +
  #geom_smooth(method = "lm", alpha = 0.1) +
  geom_point() +
  scale_color_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  theme_linedraw() +
  labs(color = "N",
       x = "N participants sign-changed",
       y = "False positive rate") +
  scale_y_continuous(limits = c(0,1), breaks = c(0, 0.05, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1)) +
  scale_x_continuous(breaks = 1:10) +
  guides(color = guide_legend(reverse = TRUE))

p_fixed_changesigns_n

# ggsave(filename = "../communications/plots/fixed_changesigns_n.pdf",
#        device = "pdf",
#        width = 6,
#        height = 5)

```

# 3b Change sign (%)

This is similar to the last simulation but instead of excluding a fixed N of participants regardless of total sample size I exclude a fixed percentage of participants.

## Simulation functions

Note: the `alter_data()` function assumes simulated data uses mean_intervention >= mean_control

```{r}

# define data generating function ----
generate_data <- function(n_per_condition,
                          mean_control,
                          mean_intervention,
                          sd_control,
                          sd_intervention) {
  
  data_control <- 
    tibble(condition = "control",
           score = rnorm(n = n_per_condition, mean = mean_control, sd = sd_control))
  
  data_intervention <- 
    tibble(condition = "intervention",
           score = rnorm(n = n_per_condition, mean = mean_intervention, sd = sd_intervention))
  
  data <- bind_rows(data_control,
                    data_intervention) 
  
  return(data)
}


alter_data <- function(data, n_participants_signchanged) {
  
  data <- data |>
    rownames_to_column(var = "rowname") |>
    mutate(altered = FALSE)
  
  # identify the N smallest scores in the intervention group
  smallest_intervention <- data |>
    filter(condition == "intervention") |>
    mutate(z_score = score / sd(score),
           abs_z_score = abs(z_score)) |>
    arrange(z_score) |> # assumes simulated data uses mean_intervention >= mean_control
    slice_head(n = n_participants_signchanged) # oversample n signchanged
  
  # identify the N largest scores in the control group
  largest_control <- data |>
    filter(condition == "control") |>
    mutate(z_score = (score - mean(score)) / sd(score),
           abs_z_score = abs(z_score)) |>
    arrange(desc(z_score)) |> # assumes simulated data uses mean_intervention >= mean_control
    slice_head(n = n_participants_signchanged) # oversample n signchanged
  
  extreme_scores <- 
    bind_rows(smallest_intervention, 
              largest_control) |>
    arrange(desc(abs_z_score)) |>
    slice_head(n = n_participants_signchanged) |> # correctly sample n signchanged
    pull(rowname)
  
  # set 'altered' to TRUE for these rows, and score_altered to NA
  data <- data |>
    mutate(altered = ifelse(rowname %in% extreme_scores, TRUE, FALSE),
           score = ifelse(altered, score*-1, score)) |>
    select(-rowname)
  
  return(data)
}


# define data analysis function ----
analyse_data <- function(data) {
  
  res_t_test <- t.test(formula = score ~ condition, 
                       data = data,
                       var.equal = TRUE,
                       alternative = "two.sided")
  
  res <- tibble(p = res_t_test$p.value)
  
  return(res)
}

```

## Run simulation

```{r}

# define experiment parameters ----
experiment_parameters_grid <- expand_grid(
  n_per_condition = c(10, seq(from = 25, to = 200, by = 25)),
  mean_control = 0,
  mean_intervention = 0, 
  sd_control = 1,
  sd_intervention = 1,
  proportion_participants_signchanged = c(.01, .025, .05, .10, .15, .10, .15, .20, .25, .30, .35),
  iteration = 1:10000
)

if(file.exists("../data/simulation_fixed_changesigns_percent.rds")){
  
  simulation_fixed_changesigns_percent <- read_rds("../data/simulation_fixed_changesigns_percent.rds")
  
} else {
  
  # set the seed ----
  # for the pseudo random number generator to make results reproducible
  set.seed(123)
  
  # run simulation ----
  simulation_fixed_changesigns_percent <- 
    # using the experiment parameters
    experiment_parameters_grid |>
    
    # generate data using the data generating function and the parameters relevant to data generation
    mutate(generated_data = future_pmap(list(n_per_condition,
                                             mean_control,
                                             mean_intervention,
                                             sd_control,
                                             sd_intervention),
                                        generate_data,
                                        .progress = TRUE,
                                        .options = furrr_options(seed = TRUE))) |>
    
    mutate(n_participants_signchanged = janitor::round_half_up(proportion_participants_signchanged * n_per_condition*2)) |>
    
    # generate data using the data generating function and the parameters relevant to data generation
    mutate(altered_data = future_pmap(list(generated_data,
                                           n_participants_signchanged),
                                      alter_data,
                                      .progress = TRUE,
                                      .options = furrr_options(seed = TRUE))) |>
    
    # apply the analysis function to the generated data using the parameters relevant to analysis
    mutate(analysis_results_original_data = future_pmap(list(generated_data),
                                                        analyse_data,
                                                        .progress = TRUE,
                                                        .options = furrr_options(seed = TRUE)),
           analysis_results_altered_data = future_pmap(list(altered_data),
                                                       analyse_data,
                                                       .progress = TRUE,
                                                       .options = furrr_options(seed = TRUE)))
  
  write_rds(simulation_fixed_changesigns_percent,
            "../data/simulation_fixed_changesigns_percent.rds", 
            compress = "gz")
  
  simulation_fixed_changesigns_percent |>
    select(-generated_data, -altered_data) |>
    write_rds("../data/simulation_fixed_changesigns_percent_nodata.rds",
              compress = "gz")
}

```

## Illustrate drops

```{r}

set.seed(42)

dat_illustrate <- simulation_fixed_changesigns_percent |>
  rownames_to_column() |>
  unnest(analysis_results_original_data) |>
  rename(p_original = p) |>
  unnest(analysis_results_altered_data) |>
  rename(p_altered = p) |>
  filter(p_original > .05 & p_altered < .05) |>
  mutate(diff = p_original - p_altered) |>
  arrange(desc(diff))

dat_illustrate |>
  slice_head(n = 10) |>
  select(n_per_condition, proportion_participants_signchanged, p_original, p_altered, diff) |>
  kable() |>
  kable_classic(full_width = FALSE)

random_row <- dat_illustrate |>
  slice_sample(n = 1) |>
  pull(rowname)

full_join(
  dat_illustrate |> 
    filter(rowname == random_row) |> 
    select(generated_data) |>
    unnest(generated_data) |>
    rownames_to_column() |>
    rename(score_original = score),
  dat_illustrate |> 
    filter(rowname == random_row) |> 
    select(altered_data) |>
    unnest(altered_data) |>
    rownames_to_column() |>
    rename(score_altered = score),
  by = c("rowname", "condition")
) |>
  ggplot(aes(condition, score_original, color = altered)) +
  geom_point(position = position_jitter(width = 0.1)) +
  scale_color_viridis_d(begin = 0.3, end = 0.7) +
  theme_linedraw() +
  ylab("Score")

dat_illustrate |>
  filter(rowname == random_row) |>
  select(p_original, p_altered) |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 3) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Summary table 

```{r}

# summarise simulation results over the iterations ----
simulation_fixed_changesigns_percent_summary <-
  simulation_fixed_changesigns_percent |>
  # convert `analysis_results` nested-data-frame column to regular columns in the df. in this case, the p value.
  unnest(analysis_results_original_data) |>
  rename(p_original_data = p) |>
  unnest(analysis_results_altered_data) |>
  rename(p_altered_data = p) |>
  # summarize across iterations
  group_by(n_per_condition, proportion_participants_signchanged) |>
  summarize(original_data_false_positive_rate = mean(p_original_data < .05),
            altered_data_false_positive_rate = mean(p_altered_data < .05)) |>
  ungroup() 

simulation_fixed_changesigns_percent_summary |>
  mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

# simulation_fixed_changesigns_percent_summary |>
#   filter(proportion_participants_signchanged == .1) |>
#   mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
#   kable() |>
#   kable_classic(full_width = FALSE)
# 
# simulation_fixed_changesigns_percent_summary |>
#   filter(proportion_participants_signchanged == .35) |>
#   mutate_if(is.numeric, janitor::round_half_up, digits = 2) |>
#   kable() |>
#   kable_classic(full_width = FALSE)

```

## Plot

```{r}

p_fixed_changesigns_percent <- ggplot(simulation_fixed_changesigns_percent_summary,
                                      aes(proportion_participants_signchanged, 
                                          altered_data_false_positive_rate, 
                                          color = as.factor(n_per_condition*2))) + 
  geom_hline(yintercept = 0.05, linetype = "dotted") +
  geom_line() +
  geom_point() +
  scale_color_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  theme_linedraw() +
  labs(color = "N",
       x = "% participants sign-changed",
       y = "False positive rate") +
  scale_y_continuous(limits = c(0,1), breaks = c(0, 0.05, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1)) +
  scale_x_continuous(labels = percent_format(), breaks = c(.01, .025, .05, .10, .15, .20, .25, .30, .35)) +
  guides(color = guide_legend(reverse = TRUE))

p_fixed_changesigns_percent

# ggsave(filename = "../communications/plots/fixed_changesigns_percent.pdf",
#        device = "pdf",
#        width = 6,
#        height = 5)

```

# Combined plots

```{r fig.height=10, fig.width=13}

p_fixed_exclusions_n +
  p_fixed_exclusions_percent +
  p_fixed_swaps_n +
  p_fixed_swaps_percent +
  p_fixed_changesigns_n + 
  p_fixed_changesigns_percent + 
  plot_layout(ncol = 2, guides = "collect")

ggsave(filename = "../communications/plots/fixed_combined.pdf",
       device = "pdf",
       width = 13,
       height = 10)

```

# Session info

```{r}

sessionInfo()

```


