---
title: "Little big lies: iterative data tampering in two-group mean-difference (t-test) designs"
author: "Ian Hussey"
date: today
format:
  html:
    embed-resources: true # ignored when rendering book
    toc: true
    toc_float: true
    code-fold: true
    code-overflow: wrap
    code-tools: true
    warning: false
    df-print: kable
    theme:
      light: cosmo
      dark: darkly
    highlight-style:
      light: breeze
      dark: dracula
    include-in-header:
      text: |
        <style>
          .kable-table table { width: 70%; margin: 0 auto; }
          .quarto-color-scheme-toggle::after { content: " Dark mode"; font-family: sans-serif; font-size: 0.85em; }
        </style>
---

# Introduction

This introduction is structured following the ADEMP framework (Aims, Data-generating process, Estimands, Methods, Performance) as described by Morris, White, and Crowther (2019) and operationalised for psychology by Siepe et al. (2024). Working through the framework explicitly makes the design decisions behind the simulations visible and auditable.

This file covers the **two-group mean-difference (Student's *t*-test)** setting. The companion files `compendium_iterative_correlation.Rmd` and `compendium_iterative_chisquare.Rmd` apply the same logic to a Pearson correlation and a 2×2 association respectively.

## Project description

This is a simulation study of **data tampering and fabrication**. It asks: starting from data generated under the null hypothesis (no true group difference), how much must a dishonest analyst tamper with the data before a two-sample *t*-test crosses *p* < .05? "Tampering" is operationalised as a family of iterative strategies — dropping cases, swapping condition labels, duplicating cases, fabricating new cases, and replacing cases — each applied a few data points at a time until significance is reached or tampering can go no further.

The motivating context is the continuum between questionable research practices and outright fabrication: the same null dataset is pushed to significance by manipulations ranging from those usually filed under "*p*-hacking" (unprincipled exclusion of inconvenient cases) to those that are unambiguously fraud (typing a different condition label, inventing rows). Quantifying the *effort* — the count of altered data points — makes concrete how little manipulation can be required.

## Prior related work

There is no closed-form benchmark for the quantity of interest: the minimal number of tampered points needed to manufacture significance is a property of a greedy tampering algorithm applied to random data, not an estimator of a population parameter, so the study is **descriptive** rather than a calibration exercise. It connects to work on the fragility of findings to a few influential observations (the `dfstat`/`dfbeta` tradition; see also Spiess et al., 2024 *[citation to be completed]*) and to explicit "how to cheat" framings such as Chambers' (2017) "dirty dozen", whose second tip (condition swapping) is simulated directly here.

## Aims

The task is the **characterisation of a tampering procedure**. For each strategy and sample size the aims are:

1. To estimate **how many tampered data points are required to reach *p* < .05** from null data, and how this scales with sample size, both in absolute terms and as a proportion of the sample.
2. To estimate the **proportion of datasets requiring more than a given share of the sample** to be tampered with — post hoc budgets of 10%, 25%, and 50% — and the proportion for which significance is **never attained** (tampering runs to a structural limit without crossing α).
3. To quantify the **effect-size bias** of the final tampered dataset relative to the truth — Hedges' *g* relative to a true standardised effect of 0.
4. To **compare strategies** on these measures.

## Data-Generating Process

### DGP specification approach

The DGP is **fully parametric** and generated under the **true null**, so any significant result is a false positive produced entirely by the tampering. Two independent conditions of `n_per_condition` observations each are drawn from a Normal distribution with common SD: $Y_{ij} \overset{\text{iid}}{\sim} \mathcal{N}(\mu_j, \sigma^2)$ with $\mu_{\text{control}} = \mu_{\text{intervention}} = 0$ and $\sigma = 1$. Produced by `generate_data()` via `rnorm()`.

### DGP factors

Two factors are varied: the **tampering strategy** and the **per-group sample size** `n_per_condition`.

### Factor values and settings

- `n_per_condition` ∈ {10, 25, 50, 75, 100, 125, 150, 175, 200}.
- Tampering strategy ∈ {dropping, condition swapping, duplication, fabrication, replacement}.
- Held constant: μ = 0 in both groups and σ = 1 (the null); α = .05, two-sided; and the greedy "most-helpful-first" rule (each step alters the single data point that most moves the test toward significance). **No tampering budget is imposed**: each strategy runs until significance is reached or a structural limit is hit (a group exhausted; no improving move left), with a loop guard at 10× the sample size as a termination backstop only. Tampering budgets (10%, 25%, 50% of the sample) are applied post hoc at summary time.

### Factor combination and number of conditions

The strategy × sample-size grid is crossed **fully factorially**: 5 strategies × 9 sample sizes = **45 conditions**, each replicated `n_iterations` times.

## Estimands and Targets

- **Primary target — tamper count.** `tampers_needed`, the number of tampered data points required to first reach *p* < .05, summarised by its **median** (with IQR and 95% prediction interval; rationale in *A note on summarising with the median rather than the mean*), by the post hoc **budget-exceedance proportions** (more than 10%, 25%, 50% of the sample), and by the **proportion never attaining significance**.
- **Secondary target — effect-size bias.** Because the true effect is 0, the final-dataset **Hedges' *g*** (relative to 0) *is* the bias; reported as the signed value (a signed median, not an absolute magnitude, since the sign is the direction of the manufactured effect).

There is no analytic ground truth; the Monte Carlo estimates are the result.

## Methods and extracted quantities

Each strategy is an iterative greedy procedure: generate a null dataset, test it, and while non-significant apply one tampering operation targeting the data point(s) most opposed to the desired result, then re-test. The analysis is the equal-variance two-sample *t*-test (`var.equal = TRUE`, two-sided, α = .05). From each replication the worker returns `tampers_needed`, `reached_sig` (whether *p* < .05 was actually reached, so no replication is silently dropped), and the final-dataset effect size (`hedges_g`, `bias_mean_diff`). Swapping *entire* conditions is handled analytically (it merely adds ½α to the false-positive rate) rather than simulated.

## Performance and Uncertainty

### Performance measures

The tamper-count distribution is summarised by its **median**, its interquartile range, and its **95% prediction interval** (the 2.5th and 97.5th percentiles, i.e. the range within which a single new study's tamper count would fall 95% of the time), with the mean and SD also tabulated to expose the skew, and as a **proportion of the sample**. The **post hoc budget-exceedance proportions** (needing more than 10%, 25%, 50% of the sample) and the **empirical detection rate** — equivalently its complement, the **proportion never attaining significance** — are reported separately. The bias is summarised by the **signed median Hedges' *g*** (which, since the truth is 0, is the bias itself). Full rationale for the median is in *A note on summarising with the median rather than the mean*.

### Monte Carlo uncertainty

Every summary carries a Monte Carlo standard error: the binomial SE for the never-significant proportion, the SE of the mean for the mean count, and a nonparametric bootstrap (Morris et al., 2019) for the median count and median bias. See the *Monte Carlo standard errors* subsection below.

### Number of simulation repetitions

Each condition is replicated `n_iterations` times: **10000** for a full run and **100** for a fast local/development render (auto-selected by the `on_hpc` knob; see *Running on an HPC cluster*). This is adequate for the descriptive comparison intended here.

### Non-convergence and missing values

The *t*-test is closed-form. The realistic failure modes are degenerate intermediate datasets (a group falling below two cases under dropping/replacement — guarded; zero within-group variance), which return a missing *p*-value and are classified as never-significant via `reached_sig`; summaries use `na.rm`. Degenerate cells (no iteration reaching significance, or an undefined effect size for every iteration) are guarded in the summarising helpers, which return `NA` for that cell rather than aborting the render.

### Interpretation of performance measures *(optional)*

There is no pass/fail criterion; the comparison is relative. A strategy is "cheap" to the extent that its median tamper count — especially its **proportion of the sample** — is small and its never-significant proportion is low. The expectation is that the proportion of data that must be tampered is small and does not grow with sample size, and that the distribution-preserving strategy (condition swapping) is both effective and the hardest to detect.

## Ethics, optics, and the *p*-hacking / fabrication continuum

This report documents weaponisable knowledge, and the framing matters; the simulations are designed to make a continuum visible, not to provide a recipe:

- Excluding participants is usually filed under *p*-hacking even when the exclusions are unprincipled and serve only to produce a significant *p*-value.
- But if, instead of deleting those rows, the analyst types the *other* condition's label into them, the same act becomes fabrication — and a firing offence.
- Treating one as common-and-merely-iffy and the other as rare-and-clearly-unacceptable risks badly underestimating the prevalence of the latter.
- The point is a trolley problem: is deleting data really ethically different from making it up, when both serve the individual's advancement at the expense of the evidence?

## Realism and scope

These are **proofs of principle**, not forensic models of real fraud. Some strategies (e.g. naive dropping) can leave obvious signatures such as a truncated otherwise-Normal distribution; many detected fraud cases were nonetheless unsophisticated, and more elaborate plausibility-preserving variants are easy to construct. The recurring theme is the illustration of **minimal tampering and effort**. Out of scope here: Likert-type outcomes; an "alter existing values" strategy distinct from fabrication; and a detectability axis.

# Dependencies

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

```{r}

library(tidyr)
library(dplyr)
library(readr)
library(forcats)
library(tibble)
library(purrr) 
library(furrr)
library(MASS)
library(effectsize)
library(ggplot2)
library(scales)
library(knitr)
library(kableExtra)
library(ggtext)
library(patchwork)

options(scipen=999)

dir.create("../data", showWarnings = FALSE)
dir.create("../communications/plots", showWarnings = FALSE)

# --- Pin BLAS/LAPACK to one thread BEFORE workers spawn ---------------------
# With dozens of worker processes, multi-threaded BLAS would oversubscribe the
# node (workers x BLAS threads all fighting for cores). The operations here use
# tiny matrices, so single-threaded BLAS loses nothing. Setting these in the
# master means the multisession (PSOCK) workers inherit them at spawn.
Sys.setenv(FLEXIBLAS_NUM_THREADS = "1", OMP_NUM_THREADS = "1",
           OPENBLAS_NUM_THREADS = "1", MKL_NUM_THREADS = "1", BLIS_NUM_THREADS = "1",
           VECLIB_MAXIMUM_THREADS = "1") # VECLIB covers macOS Accelerate (local testing)
if (requireNamespace("flexiblas", quietly = TRUE)) flexiblas::flexiblas_set_num_threads(1L)
if (requireNamespace("RhpcBLASctl", quietly = TRUE)) {
  RhpcBLASctl::blas_set_num_threads(1)
  RhpcBLASctl::omp_set_num_threads(1)
}

# --- Parallel plan ----------------------------------------------------------
# parallelly::availableCores() honours a SLURM allocation, cgroup CPU quotas,
# and affinity masks, so it returns the cores granted to THIS session rather than
# all physical cores. RStudio Server is not fork-safe, so use multisession
# (separate R processes), not multicore. furrr_options(seed = TRUE) advances an
# L'Ecuyer-CMRG stream per task, giving reproducibility independent of the
# worker count.
plan(multisession, workers = parallelly::availableCores())

```

# Common functions and global variables

```{r}

# --- Replication knob (local vs cluster) ------------------------------------
# RStudio Server does not pass the job's SLURM_* allocation variables into the R
# session, but the HPC sets some SLURM_* global that survives and is absent on a
# laptop, so "any SLURM_* variable present" is a reliable cluster detector.
# run_full forces the choice: NULL = auto-detect, TRUE = full run, FALSE = quick
# local smoke test.
run_full <- TRUE
on_hpc <- if (!is.null(run_full)) isTRUE(run_full) else any(grepl("^SLURM_", names(Sys.getenv())))

# n iterations for all sims: full cluster run vs fast local smoke test
n_iterations <- if (on_hpc) 10000L else 100L

# Cache tag: keeps the full and dev runs in separate .rds files so they never
# collide. cache_path() builds a tagged path under ../data for each simulation.
cache_tag <- paste0(if (on_hpc) "full" else "dev", "_k", n_iterations)
cache_path <- function(stem) paste0("../data/", stem, "_", cache_tag, ".rds")

# Packages each worker must attach (belt-and-braces; future also auto-detects
# the packages of exported globals). MASS is only needed by the correlation sims.
worker_packages <- c("dplyr", "tidyr", "tibble", "purrr", "effectsize", "MASS")

# Loop guard for the tampering while-loops ----
# There is NO tampering budget in the design: each strategy runs until significance
# is reached or a structural limit is hit (data exhausted; no improving move left).
# tamper_loop_guard() is a termination backstop only, set at 10x the sample size --
# far beyond anything the greedy procedures need -- so that a pathological iteration
# (e.g., a duplication pool whose limiting effect is too weak to ever cross alpha)
# cannot loop unboundedly. Iterations that hit the guard or a structural limit are
# recorded as never attaining significance. Tampering budgets (10%, 25%, 50% of the
# sample) are applied POST HOC in summarise_tampers(), which is possible precisely
# because the full, unbudgeted tamper counts are recorded.
tamper_loop_guard <- function(n_total) 10L * n_total

# Display order of the tampering strategies ----
# Strategies are ordered by their impact: from the strategy that needs the SMALLEST
# proportion of the sample to be tampered with to reach p < .05 (largest impact per
# tampered data point) to the one that needs the largest. The order is fixed across
# the mean-difference and correlation reports so that the plots and tables are
# directly comparable, using each strategy's median proportion tampered averaged
# over sample sizes and pooled across those two settings. The effect-size bias at
# the point significance is reached is near-identical across strategies (the greedy
# procedure stops on first crossing), so effort is what discriminates them.
method_levels <- c("Replacement", "Condition swapping", "Fabrication", "Dropping", "Duplication")

# define data generating function ----
generate_data <- function(n_per_condition,
                          mean_control,
                          mean_intervention,
                          sd_control,
                          sd_intervention) {
  
  data_control <- 
    tibble(condition = "control",
           score = rnorm(n = n_per_condition, mean = mean_control, sd = sd_control))
  
  data_intervention <- 
    tibble(condition = "intervention",
           score = rnorm(n = n_per_condition, mean = mean_intervention, sd = sd_intervention))
  
  data <- bind_rows(data_control,
                    data_intervention) 
  
  return(data)
}

# define data analysis function ----
# slow, because t.test is slow. 
analyse_data <- function(data) {

  # Guard against a degenerate group. Uncapped dropping can reduce a condition to a single
  # observation, at which point t.test() aborts with "not enough observations" and takes
  # the whole future_pmap() run with it. The tampering loop's own guard checks the group
  # sizes BEFORE a drop, so a group of exactly two passes it and is then reduced to one.
  # Returning NA here instead ends that iteration cleanly: the while loop's !is.na(p)
  # condition fails, and the iteration is recorded as never attaining significance, which
  # is correct -- with one observation in a group no test can be computed, so significance
  # is not attainable. This matches p_ttest() in compendium_iterative_ttest_optimised.qmd.
  d <- data[!is.na(data$score), , drop = FALSE]
  if (sum(d$condition == "control") < 2 || sum(d$condition == "intervention") < 2) {
    return(tibble(p = NA_real_))
  }

  res_t_test <- t.test(formula = score ~ condition,
                       data = data,
                       var.equal = TRUE,
                       alternative = "two.sided")
  
  res <- tibble(p = res_t_test$p.value)

  return(res)
}

# extract Hedges' g and the raw mean-difference bias from a (possibly tampered) dataset ----
# The data are generated under the null (mean_control = mean_intervention = 0), so the
# true standardised effect is 0. Hedges' g of the final, tampered dataset is therefore a
# direct quantification of the bias the tampering has introduced relative to mu = 0.
# Hedges' g is signed intervention-minus-control (positive = intervention scored higher).
# The signed g is reported throughout: because the truth is 0, the signed median IS the
# bias, and it also shows the direction each strategy induces. (An absolute |g| would be
# misleading: it is strictly positive even under no systematic bias.)
compute_bias <- function(data) {

  data <- data |> dplyr::filter(!is.na(score))

  # A condition reduced to a single observation contributes no within-group variance, so
  # the pooled SD is estimated from the other group alone and Hedges' g, while still
  # numerically computable, is not interpretable. Return NA rather than a number. (The raw
  # mean difference below remains well defined and is still reported.) This matches
  # hedges_g_fast() in compendium_iterative_ttest_optimised.qmd.
  if (sum(data$condition == "control") < 2 || sum(data$condition == "intervention") < 2) {
    return(tibble(hedges_g = NA_real_,
                  bias_mean_diff = mean(data$score[data$condition == "intervention"]) -
                                   mean(data$score[data$condition == "control"])))
  }

  # force factor order so g is computed as intervention - control
  data <- data |> mutate(condition = factor(condition, levels = c("intervention", "control")))

  # ci = NULL skips the confidence interval, which is never used here and which is the only
  # part that can fail: effectsize computes the interval by numerical optimisation, and for
  # a heavily tampered dataset with almost no within-group variance (g in the tens) that
  # optimiser aborts, discarding a perfectly well-defined point estimate.
  hedges_g_value <- tryCatch(
    as.numeric(effectsize::hedges_g(score ~ condition, data = data, ci = NULL)$Hedges_g),
    error = function(e) NA_real_
  )

  means <- data |>
    dplyr::group_by(condition) |>
    dplyr::summarise(m = mean(score), .groups = "drop")

  m_control      <- means$m[means$condition == "control"]
  m_intervention <- means$m[means$condition == "intervention"]
  if (length(m_control) == 0)      m_control      <- NA_real_
  if (length(m_intervention) == 0) m_intervention <- NA_real_

  tibble(hedges_g       = hedges_g_value,
         bias_mean_diff = m_intervention - m_control)
}

# bundle the number of tampered cases with the final-dataset effect size / bias ----
# reached_sig records whether the study actually crossed p < .05 (TRUE) or stopped
# at a structural limit / the loop guard without significance (FALSE). Tracking this
# explicitly — rather than inferring it from the count — means a result can never be
# silently dropped from both the central-tendency stats and the never-attained proportion.
summarise_tamper_result <- function(data, tampers_needed, reached_sig) {
  bind_cols(tibble(tampers_needed = tampers_needed, reached_sig = reached_sig),
            compute_bias(data))
}

# # manual t-test p-value calculation function
# analyse_data <- function(data) {
#   x <- data$score[data$condition == "control"]
#   y <- data$score[data$condition == "intervention"]
#   
#   # Basic checks like t.test() would do
#   if (length(x) < 2 || length(y) < 2 || all(is.na(x)) || all(is.na(y))) {
#     return(tibble(p = NA_real_))
#   }
#   
#   if (sd(x, na.rm = TRUE) == 0 || sd(y, na.rm = TRUE) == 0) {
#     return(tibble(p = NA_real_))
#   }
# 
#   # Remove NAs for safety
#   x <- x[!is.na(x)]
#   y <- y[!is.na(y)]
#   
#   n1 <- length(x)
#   n2 <- length(y)
#   m1 <- mean(x)
#   m2 <- mean(y)
#   s1 <- sd(x)
#   s2 <- sd(y)
# 
#   s_pooled <- sqrt(((n1 - 1) * s1^2 + (n2 - 1) * s2^2) / (n1 + n2 - 2))
#   
#   # Catch divide-by-zero or NaN
#   if (is.nan(s_pooled) || s_pooled == 0) {
#     return(tibble(p = NA_real_))
#   }
# 
#   t_stat <- (m1 - m2) / (s_pooled * sqrt(1/n1 + 1/n2))
#   df <- n1 + n2 - 2
#   p_val <- 2 * pt(-abs(t_stat), df)
# 
#   if (is.na(p_val)) p_val <- NA_real_
# 
#   return(tibble(p = p_val))
# }

# Monte Carlo standard errors (MCSEs) ----
# The mean and proportion MCSEs use the closed-form expressions from Siepe et al. (2024)
# / Morris et al. (2019), matching simulation_template_estimation.qmd:
#   MCSE(mean)       = SD / sqrt(K)
#   MCSE(proportion) = sqrt(p (1 - p) / K)
# The median has no neat closed form under skew, so its MCSE is obtained by the
# nonparametric bootstrap recommended by Morris et al. (2019). The RNG state is saved
# and restored so that computing summaries never perturbs the simulation seeds.
mcse_mean <- function(x){
  x <- x[!is.na(x)]
  if (length(x) < 2) return(NA_real_)
  sd(x) / sqrt(length(x))
}

mcse_proportion <- function(p, k){
  if (is.na(p) || k < 1) return(NA_real_)
  sqrt(p * (1 - p) / k)
}

mcse_median <- function(x, n_boot = 1000, seed = 42){
  x <- x[!is.na(x)]
  n <- length(x)
  if (n < 2) return(NA_real_)
  old_seed <- if (exists(".Random.seed", envir = .GlobalEnv)) get(".Random.seed", envir = .GlobalEnv) else NULL
  set.seed(seed)
  boot_medians <- replicate(n_boot, median(x[sample.int(n, n, replace = TRUE)]))
  if (!is.null(old_seed)) assign(".Random.seed", old_seed, envir = .GlobalEnv)
  # a censored median (Inf) has no estimable MCSE
  if (!all(is.finite(boot_medians))) return(NA_real_)
  sd(boot_medians)
}

# Guards and censoring-aware summaries ----
# Two things can go wrong in a single condition cell and take a whole render with them.
# (a) A degenerate cell: no iteration reached significance, or the effect size is
#     undefined for every iteration (an empty 2x2 margin at very low prevalence, a
#     constant variable, a group reduced below two cases). base R then warns and returns
#     +/-Inf from min()/max() and NaN from mean(). The safe_* wrappers return NA instead,
#     so the cell is reported as missing and every other cell still renders.
# (b) Right censoring: an iteration can end without significance when tampering hits
#     a structural limit (data exhausted; no improving move) or the loop guard. Such an
#     iteration has no finite tamper count -- significance was never attained, so its
#     count is unbounded above. Dropping those iterations and summarising the
#     survivors answers a different question ("given that tampering worked, how many
#     edits?") from the one this report asks ("starting from null data, how many edits?"),
#     and biases the answer downwards. Representing them as +Inf puts them in their
#     correct place at the top of the ordering: a quantile falling among the finite
#     counts is then estimated exactly, and one falling among the censored cases returns
#     Inf -- reported as "never" rather than silently reported as the survivors' quantile.
safe_min  <- function(x){ x <- x[is.finite(x)]; if (length(x) == 0) NA_real_ else min(x) }
safe_max  <- function(x){ x <- x[is.finite(x)]; if (length(x) == 0) NA_real_ else max(x) }
safe_mean <- function(x){ x <- x[is.finite(x)]; if (length(x) == 0) NA_real_ else mean(x) }
safe_sd   <- function(x){ x <- x[is.finite(x)]; if (length(x) < 2)  NA_real_ else sd(x) }

# quantile of a possibly right-censored vector (censored observations coded as Inf)
q_censored <- function(x, probs){
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_real_)
  unname(quantile(x, probs = probs, names = FALSE, type = 7))
}

median_censored <- function(x) q_censored(x, 0.5)

# median absolute deviation of a censored vector; NA when the median itself is censored
mad_censored <- function(x){
  x <- x[!is.na(x)]
  m <- q_censored(x, 0.5)
  if (length(x) == 0 || !is.finite(m)) return(NA_real_)
  1.4826 * q_censored(abs(x - m), 0.5)
}

# Table formatting ----
# Following simulation_template_lego.qmd, a point estimate is printed with its Monte
# Carlo standard error in parentheses and an interval in square brackets. The number of
# decimal places is driven by the MCSE (Morris et al., 2019): the MCSE is shown to two
# significant figures and the estimate to the same number of places. Rounding the whole
# table to a fixed 2 dp -- as an earlier version did -- erased exactly the resolution the
# substantive claims rest on: median proportions of .035, .028 and .023 all printed as
# "0.04"/"0.03"/"0.02", and every proportion MCSE printed as "0.00".
dp_from_mcse <- function(mcse, min_dp = 0, max_dp = 4){
  mcse <- mcse[is.finite(mcse) & mcse > 0]
  if (length(mcse) == 0) return(min_dp)
  min(max_dp, max(min_dp, max(ceiling(-log10(mcse)) + 1)))
}

# A censored quantile is Inf; it is printed as "never" rather than as a number:
# that percentile of datasets ran to a structural limit without attaining
# significance, so no finite tamper count exists for it.
fmt_num <- function(x, dp){
  ifelse(is.na(x), "NA",
         ifelse(is.finite(x), formatC(x, format = "f", digits = dp), "never"))
}

mcse_cell <- function(estimate, mcse, min_dp = 0, max_dp = 4){
  dp <- dp_from_mcse(mcse, min_dp = min_dp, max_dp = max_dp)
  ifelse(is.finite(estimate) & is.finite(mcse),
         paste0(fmt_num(estimate, dp), " (", fmt_num(mcse, dp), ")"),
         fmt_num(estimate, dp))
}

interval_cell <- function(lower, upper, dp = 0){
  paste0("[", fmt_num(lower, dp), ", ", fmt_num(upper, dp), "]")
}

tampers_note <- paste(
  "Values in parentheses are +/-1 Monte Carlo standard error.",
  "IQR is the 25th-75th percentile and PI the 95% prediction interval (2.5th-97.5th",
  "percentile) of the results across iterations; neither is an MCSE, and both describe",
  "the spread of the simulated studies rather than the precision of the summary.",
  "All percentile-based summaries are computed over all iterations, with datasets that",
  "never attained significance (tampering ran to a structural limit without crossing",
  "alpha) ranked above every finite count; 'never' marks a percentile falling among",
  "those datasets. The '> b% of N' columns are post hoc tampering budgets: the",
  "proportion of datasets needing more than b% of the original sample to be tampered",
  "with, never-significant datasets exceeding every budget by definition.",
  "The mean and SD are conditional on attaining significance and are downward-biased",
  "for the unconditional mean whenever 'Never attained' exceeds 0; read them as a",
  "within-cell skew diagnostic, not across cells.",
  "The 0th/100th percentiles (min/max) are held in the summary object but not printed:",
  "they drift with the number of iterations and so are not stable quantities."
)

bias_note <- paste(
  "Values in parentheses are +/-1 Monte Carlo standard error (nonparametric bootstrap",
  "for medians). PI is the 95% prediction interval (2.5th-97.5th percentile) of the",
  "effect sizes across iterations. Because the true effect is zero, the signed median",
  "effect size in the final tampered dataset is itself the bias."
)

# Tamper-count table. extra_cols names any grouping columns (e.g. "Method") that should
# be printed before the sample size.
table_tampers <- function(data, caption, extra_cols = NULL){
  data |>
    transmute(across(all_of(extra_cols)),
              `N total`               = n_total,
              `Never attained`        = mcse_cell(proportion_never_sig, proportion_never_sig_mcse, min_dp = 3),
              `> 10% of N`            = mcse_cell(prop_gt_10, prop_gt_10_mcse, min_dp = 3),
              `> 25% of N`            = mcse_cell(prop_gt_25, prop_gt_25_mcse, min_dp = 3),
              `> 50% of N`            = mcse_cell(prop_gt_50, prop_gt_50_mcse, min_dp = 3),
              `Median tampers`        = mcse_cell(median, median_mcse),
              `IQR`                   = interval_cell(iqr_lower, iqr_upper),
              `95% PI`                = interval_cell(pi_lower, pi_upper),
              `Mean (reached only)`   = mcse_cell(mean_reached, mean_reached_mcse),
              `SD (reached only)`     = fmt_num(sd_reached, 2),
              `Median prop. tampered` = mcse_cell(prop_median, prop_median_mcse, min_dp = 3)) |>
    kable(caption = caption) |>
    kable_classic(full_width = FALSE) |>
    footnote(general = tampers_note, general_title = "Note.", footnote_as_chunk = TRUE)
}

summarise_tampers <- function(data){
  data |>
    # code an iteration that never attained significance as Inf: tampering ran to a
    # structural limit (or the loop guard) without crossing alpha, so its count is
    # "more than anything observed", which ranks it above every finite count
    mutate(tampers_censored = ifelse(reached_sig, tampers_needed, Inf)) |>
    group_by(n_per_condition) |>
    summarize(k = n(), # total iterations in this cell
              k_reached = sum(reached_sig), # iterations that attained significance
              # Empirical detection rate: the proportion of null datasets that can be
              # pushed over p < .05 at all. proportion_never_sig is its complement (the
              # two sum to 1 and share the same binomial MCSE): datasets for which the
              # strategy ran to its structural limit without ever crossing alpha.
              detection_rate = mean(reached_sig),
              detection_rate_mcse = mcse_proportion(detection_rate, k),
              proportion_never_sig = 1 - detection_rate,
              proportion_never_sig_mcse = detection_rate_mcse,
              # --- post hoc tampering budgets: the proportion of datasets needing more
              # than b% of the original sample to be tampered with. Because the full,
              # unbudgeted counts are recorded, any budget can be evaluated at summary
              # time; never-significant datasets exceed every budget by definition (Inf).
              prop_gt_10 = mean(tampers_censored > 0.10 * n_per_condition * 2),
              prop_gt_10_mcse = mcse_proportion(prop_gt_10, k),
              prop_gt_25 = mean(tampers_censored > 0.25 * n_per_condition * 2),
              prop_gt_25_mcse = mcse_proportion(prop_gt_25, k),
              prop_gt_50 = mean(tampers_censored > 0.50 * n_per_condition * 2),
              prop_gt_50_mcse = mcse_proportion(prop_gt_50, k),
              # --- headline summaries: computed over ALL k iterations, with the
              # never-significant ones ranked above every finite count
              median = median_censored(tampers_censored),
              median_mcse = mcse_median(tampers_censored),
              mad = mad_censored(tampers_censored),
              # the interquartile range (a 50% interval) and the 95% prediction interval
              # (2.5th-97.5th percentile), i.e. the range within which a single new study's
              # tamper count would fall 95% of the time. Named so that neither can be
              # mistaken for the other or for the +/-1 MCSE of the median. A bound is Inf
              # (printed as "never") if that percentile falls among the never-significant
              # datasets.
              iqr_lower = q_censored(tampers_censored, 0.25),
              iqr_upper = q_censored(tampers_censored, 0.75),
              pi_lower  = q_censored(tampers_censored, 0.025),
              pi_upper  = q_censored(tampers_censored, 0.975),
              # --- conditional summaries: over the iterations that DID attain
              # significance. The mean is not estimable over all k when any iteration
              # never attains significance (those counts are unbounded above), so the
              # mean and SD here are conditional and downward-biased for the unconditional
              # mean whenever proportion_never_sig > 0. They are a within-cell skew
              # diagnostic against median_reached, not a headline summary. min/max are
              # the 0th and 100th percentiles: unstable (the max drifts upward as k
              # grows), so retained for diagnostics but neither tabled nor plotted.
              median_reached = median_censored(tampers_needed[reached_sig]),
              mean_reached = safe_mean(tampers_needed[reached_sig]),
              mean_reached_mcse = mcse_mean(tampers_needed[reached_sig]),
              sd_reached = safe_sd(tampers_needed[reached_sig]),
              min_reached = safe_min(tampers_needed[reached_sig]),
              max_reached = safe_max(tampers_needed[reached_sig])) |>
    ungroup() |>
    mutate(n_total = n_per_condition*2,
           prop_median = median/n_total,
           prop_median_mcse = median_mcse/n_total,
           prop_mad = mad/n_total,
           prop_iqr_lower = iqr_lower/n_total,
           prop_iqr_upper = iqr_upper/n_total,
           prop_pi_lower  = pi_lower/n_total,
           prop_pi_upper  = pi_upper/n_total,
           prop_mean_reached = mean_reached/n_total,
           prop_mean_reached_mcse = mean_reached_mcse/n_total,
           prop_sd_reached = sd_reached/n_total,
           prop_min_reached = min_reached/n_total,
           prop_max_reached = max_reached/n_total) |>
    relocate(n_total, .before = n_per_condition) |>
    dplyr::select(-n_per_condition)
}

# summarise the bias (Hedges' g relative to mu = 0) in the final tampered datasets ----
summarise_bias <- function(data){
  data |>
    group_by(n_per_condition) |>
    # Named by estimand and metric, following the lego template's convention
    # (mean_diff_bias / smd_bias): both quantities below are biases, one on the
    # standardised scale and one on the raw mean-difference scale.
    summarize(smd_bias_median            = median(hedges_g, na.rm = TRUE),
              smd_bias_median_mcse       = mcse_median(hedges_g),
              smd_bias_iqr_lower         = quantile(hedges_g, probs = 0.25, na.rm = TRUE),
              smd_bias_iqr_upper         = quantile(hedges_g, probs = 0.75, na.rm = TRUE),
              smd_bias_pi_lower          = quantile(hedges_g, probs = 0.025, na.rm = TRUE),
              smd_bias_pi_upper          = quantile(hedges_g, probs = 0.975, na.rm = TRUE),
              mean_diff_bias_median      = median(bias_mean_diff, na.rm = TRUE),
              mean_diff_bias_median_mcse = mcse_median(bias_mean_diff),
              mean_diff_bias_pi_lower    = quantile(bias_mean_diff, probs = 0.025, na.rm = TRUE),
              mean_diff_bias_pi_upper    = quantile(bias_mean_diff, probs = 0.975, na.rm = TRUE)) |>
    ungroup() |>
    mutate(n_total = n_per_condition * 2) |>
    relocate(n_total, .before = n_per_condition) |>
    dplyr::select(-n_per_condition)
}

table_bias <- function(data, caption, extra_cols = NULL){
  data |>
    transmute(across(all_of(extra_cols)),
              `N total`                    = n_total,
              `Median Hedges' g (bias)`    = mcse_cell(smd_bias_median, smd_bias_median_mcse, min_dp = 2),
              `95% PI`                     = interval_cell(smd_bias_pi_lower, smd_bias_pi_upper, dp = 2),
              `Median raw difference (bias)` = mcse_cell(mean_diff_bias_median, mean_diff_bias_median_mcse, min_dp = 2),
              `95% PI (raw)`               = interval_cell(mean_diff_bias_pi_lower, mean_diff_bias_pi_upper, dp = 2)) |>
    kable(caption = caption) |>
    kable_classic(full_width = FALSE) |>
    footnote(general = bias_note, general_title = "Note.", footnote_as_chunk = TRUE)
}

# plot the (signed) bias (Hedges' g relative to mu = 0) against sample size ----
plot_bias <- function(data, title){
  ggplot(data, aes(n_total, smd_bias_median)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
    geom_point() +
    geom_linerange(aes(ymin = smd_bias_pi_lower, ymax = smd_bias_pi_upper)) +
    labs(title = title,
         subtitle = "Median and 95% prediction interval (2.5th-97.5th percentile) across iterations",
         x = "Sample size",
         y = "Hedges' g in final dataset (bias relative to mu = 0)") +
    theme_linedraw() +
    scale_y_continuous(breaks = breaks_pretty(n = 8)) +
    scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2)
}

plot_distributions <- function(data, title, x_axis){
  ggplot(data, aes(x = tampers_needed)) +
    geom_histogram(binwidth = 1) +
    labs(title = title, 
         x = x_axis,
         y = "Count") +
    ylim(0, NA) +
    theme_linedraw() +
    scale_x_continuous(breaks = breaks_pretty(n = 5)) +
    facet_wrap(~ n_per_condition) # N PER CONDITION NOT TOTAL N
}

plot_tampers_n <- function(data, title, y_axis){
  ggplot(data, aes(n_total, median)) +
    geom_point() +
    # an interval bar is omitted (NA) where its upper percentile falls among the
    # never-significant datasets, since no finite value exists to draw it to
    geom_linerange(aes(ymin = pi_lower, ymax = ifelse(is.finite(pi_upper), pi_upper, NA_real_))) +
    labs(title = title,
         subtitle = "Median and 95% prediction interval across iterations; bars omitted where the upper percentile is never attained",
         x = "Sample size",
         y = y_axis) +
    theme_linedraw() +
    scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
    scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2)
}

plot_tampers_prop <- function(data, title, y_axis){
  ggplot(data, aes(n_total, prop_median)) +
    geom_point() +
    geom_linerange(aes(ymin = prop_pi_lower, ymax = ifelse(is.finite(prop_pi_upper), prop_pi_upper, NA_real_))) +
    labs(title = title,
         subtitle = "Median and 95% prediction interval across iterations; bars omitted where the upper percentile is never attained",
         x = "Sample size",
         y = y_axis) +
    theme_linedraw() +
    scale_y_continuous(labels = percent_format(), limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
    scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2)
}

```

# A note on summarising with the median rather than the mean

Each simulation records `tampers_needed`: the number of tampering operations applied before the result crossed *p* < .05. This quantity is summarised throughout with the **median** (with the interquartile range and the 95% prediction interval) rather than the mean. There are three reasons.

1. **The distribution is right-skewed.** Most datasets reach significance after only a few tampers, but a long tail require many. Part of the point of these simulations is to *show* that skew, so the central-tendency measure should not hide it. The summary tables report `mean_reached` and `sd_reached` alongside the median, its interquartile range and its 95% prediction interval, precisely so the skew is visible: the mean sits above the median, and the size of that gap indexes the length of the tail. Read that gap only within a condition, and only where no censoring has occurred — where `proportion_never_sig` is above zero the two statistics are computed over different sets of iterations (all of them for the median, the uncensored ones for the mean), so the comparison no longer isolates skew. `median_reached` is carried in the summary object for anyone who wants the like-for-like comparison.

2. **The distribution is censored, and this is the decisive reason.** There is no design budget, but tampering can still end without significance when a structural limit is hit — the data are exhausted under dropping, no improving re-pairing exists, a donor cell empties — or, pathologically, when the loop guard (10× the sample) fires. Those iterations (reported as `proportion_never_sig`, the censoring fraction) have no finite tamper count at all: significance was never attained, so their count is unbounded above. Under such right censoring:

   - The **median, IQR and prediction interval are computed over all iterations**, with the never-significant ones ranked above every finite count (coded as `Inf`). Any percentile that falls among the finite counts is then estimated exactly and is completely unaffected by the censored cases; any percentile falling among them is reported as `never` rather than being silently replaced by the corresponding percentile of the survivors. This matters: summarising only the iterations that attained significance answers "*given* that tampering worked, how many edits?", whereas the question this report asks is "starting from null data, how many edits?".
   - The **mean cannot be computed that way at all**, because the censored counts are unbounded above. The mean and SD reported here are therefore *conditional on reaching significance* (`mean_reached`, `sd_reached`), and are downward-biased estimates of the unconditional mean by construction, since they discard precisely the largest values. The size of that bias grows with the censoring fraction, so they must not be compared across conditions with different `proportion_never_sig`.

3. **The median refers to an actual case.** For an integer count, a mean such as "2.7 tampers" corresponds to no real study, whereas a median of 2 describes a realisable scenario ("half the time, two edits were enough"). Because the substantive claim concerns the realistic effort an individual would expend, the statistic that points at an actual case is the more faithful one.

Note that the central-tendency statistics **include** the datasets that were significant from the outset and so required zero tampers (the ~5% expected under the null at α = .05). The median therefore answers "starting from null data, how many edits to significance?" rather than "*given* that tampering was needed, how many?". This framing was chosen deliberately because it matches the question of how readily null data can be pushed to significance.

## Monte Carlo standard errors

Because each summary statistic is itself estimated from a finite number of simulation iterations (`n_iterations`), it carries Monte Carlo error. Following the approach in `simulation_template_estimation.qmd` (and Siepe et al., 2024; Morris et al., 2019), every table reports a Monte Carlo standard error (MCSE) next to each estimated quantity, so that apparent differences between conditions can be judged against simulation noise. The MCSEs are computed as:

- **Proportion never significant** (`proportion_never_sig_mcse`): the binomial standard error $\sqrt{\hat p\,(1-\hat p)/K}$, where *K* = `n_iterations`.
- **Mean** (`mean_reached_mcse`): the standard error of the mean $\text{SD}/\sqrt{K_\text{reached}}$, where $K_\text{reached}$ is the number of iterations that attained significance (i.e. the iterations the conditional mean is computed over).
- **Median** (`median_mcse`): obtained by nonparametric bootstrap (1,000 resamples), because the count distribution is skewed and has no convenient closed-form quantile MCSE. The same bootstrap is used for the median effect-size (bias) columns.

Two things are deliberately *not* MCSEs and are never plotted as such. The **interquartile range** (`iqr_lower`/`iqr_upper`) and the **95% prediction interval** (`pi_lower`/`pi_upper`, the 2.5th and 97.5th percentiles) describe the spread of the simulated studies themselves, not the precision with which the summary is estimated; the prediction interval is the wider and is what the plots show. The 0th and 100th percentiles (`min`/`max`) are computed and retained in the summary objects but are neither tabled nor plotted, because the maximum of *K* draws drifts upward as *K* grows and so is not a stable quantity.

A caveat specific to the median of an integer count: at *K* = 10,000 the bootstrap MCSE is frequently exactly 0, because every resample returns the same integer. That is honest rather than an error — the median is pinned — but it means the MCSE column is uninformative for those cells, and it is one respect in which the mean (whose MCSE is smooth and always positive) is better behaved. It is not a reason to prefer the mean as the headline summary, for the censoring reason given above. A censored median has no estimable MCSE at all and is reported as `NA`.

Tables print each estimate with its MCSE in parentheses, following `simulation_template_lego.qmd`. The number of decimal places is set from the MCSE rather than fixed for the whole table: the MCSE is shown to two significant figures and the estimate to the same number of places (Morris et al., 2019). This matters here because the headline quantity — the proportion of the sample that must be tampered with — falls between .02 and .04 across the larger sample sizes, and a fixed 2 dp collapsed those distinct values onto the same printed number.

The `median_mcse` is the relevant uncertainty for the headline statistic; with `n_iterations` in the low thousands it is typically a small fraction of a tamper, confirming the medians are estimated precisely. Because the tamper counts are discrete, the bootstrap median MCSE is sometimes exactly **0**: this is not an error but indicates that the median is invariant to resampling — its sampling variability is below the one-unit resolution of the data, so the median is estimated to sub-count precision. (The bootstrap saves and restores the RNG state, so it never disturbs the simulation seeds.) The `mean_mcse`, which is always non-zero, provides a complementary, continuous-scale check on Monte Carlo precision in these cases.

# Dropping

Unwarranted exclusion of cases: each round, drop the single most hostile remaining case (the lowest scorer in the intervention group or the highest scorer in the control group, whichever is the more extreme within its own group), then re-test. Repeat until significance is reached or a group would fall below two cases.

## Simulation functions

```{r}

# Drop the single most hostile remaining case (lowest in intervention or highest in control)
iterative_drop_once <- function(data) {

  data_incl <- data |> filter(!is.na(score))

  data_incl <- data_incl |>
    group_by(condition) |>
    mutate(z_score_group = (score - mean(score)) / sd(score)) |>
    ungroup()

  smallest_in_intervention <- data_incl |>
    filter(condition == "intervention") |>
    slice_min(z_score_group, n = 1)

  largest_in_control <- data_incl |>
    filter(condition == "control") |>
    slice_max(z_score_group, n = 1)

  # with_ties = FALSE matters here and is not cosmetic. A group of exactly two observations
  # always has z-scores of exactly +/-0.7071, so once both groups are down to two, the two
  # candidates tie on abs(z). slice_max() would then return BOTH rows, and the ifelse()
  # below would compare a length-n row_id vector against a length-2 one, recycling it and
  # setting essentially arbitrary observations to NA. Taking the first row breaks the tie
  # deterministically in favour of the intervention case, matching the optimised
  # implementation in compendium_iterative_ttest_optimised.qmd.
  to_drop <- bind_rows(smallest_in_intervention, largest_in_control) |>
    slice_max(abs(z_score_group), n = 1, with_ties = FALSE)

  data <- data |> mutate(score = ifelse(row_id == to_drop$row_id, NA, score))

  return(data)
}

# Main simulation function
run_iterative_dropping_study <- function(n_per_condition,
                                         mean_control,
                                         mean_intervention,
                                         sd_control,
                                         sd_intervention) {

  # Generate original data
  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) |>
    mutate(row_id = row_number())

  # Initial p-value
  p_val <- analyse_data(data)$p[1]
  if (p_val < 0.05) return(summarise_tamper_result(data, 0, TRUE))

  n_total <- nrow(data)
  loop_guard <- tamper_loop_guard(n_total)
  drops_so_far <- 0

  while (!is.na(p_val) && p_val >= 0.05 && drops_so_far < loop_guard) {
    # Stop if dropping would leave a condition with fewer than 2 cases
    data_incl <- data |> filter(!is.na(score))
    if (any(table(data_incl$condition) < 2)) break

    data <- iterative_drop_once(data)
    drops_so_far <- drops_so_far + 1
    p_val <- analyse_data(data)$p[1]
  }

  return(summarise_tamper_result(data, drops_so_far, !is.na(p_val) && p_val < 0.05))
}

```

## Run simulation

```{r}

simulation_iterative_dropping_meandiff_path <- cache_path("simulation_iterative_dropping_meandiff")
if (file.exists(simulation_iterative_dropping_meandiff_path)) {
  simulation_iterative_dropping_meandiff <- read_rds(simulation_iterative_dropping_meandiff_path)
} else {

  # Parameter grid
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_iterative_dropping_meandiff <- experiment_parameters_grid |>
    slice_sample(prop = 1) |> # shuffle rows so that the expensive large-n tasks are spread evenly across workers
    mutate(res = future_pmap(.l = list(n_per_condition = n_per_condition,
                                       mean_control = mean_control,
                                       mean_intervention = mean_intervention,
                                       sd_control = sd_control,
                                       sd_intervention = sd_intervention),
                             .f = run_iterative_dropping_study,
                             .progress = TRUE,
                             .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages))) |>
    unnest(res)

  write_rds(simulation_iterative_dropping_meandiff, simulation_iterative_dropping_meandiff_path)
}

# summarize tamper counts and bias (Hedges' g relative to mu = 0)
simulation_iterative_dropping_meandiff_tampers <- summarise_tampers(simulation_iterative_dropping_meandiff)
simulation_iterative_dropping_meandiff_bias    <- summarise_bias(simulation_iterative_dropping_meandiff)

```

## Table

```{r}

table_tampers(simulation_iterative_dropping_meandiff_tampers,
              caption = "Dropping: tampered data points needed to reach p < .05")

```

## Plot

```{r}

plot_distributions(simulation_iterative_dropping_meandiff, "Dropping", "N dropped")

p_iterative_dropping_meandiff_n <- plot_tampers_n(simulation_iterative_dropping_meandiff_tampers, "Dropping", "N dropped")

p_iterative_dropping_meandiff_prop <- plot_tampers_prop(simulation_iterative_dropping_meandiff_tampers, "Dropping", "N dropped")

```

## Bias (Hedges' g relative to μ = 0)

```{r}

table_bias(simulation_iterative_dropping_meandiff_bias,
           caption = "Dropping: effect size in the final tampered dataset (bias relative to a true effect of zero)")

p_iterative_dropping_meandiff_bias <- plot_bias(simulation_iterative_dropping_meandiff_bias, "Dropping")

p_iterative_dropping_meandiff_bias

```

# Condition swapping

## Simulation functions

```{r}

# Swap one participant from one condition to the other.
# A swap moves a single participant's label, so it is counted as ONE tamper (matching
# the other methods). To keep the two group sizes balanced over pairs of swaps, the
# direction alternates: when move_intervention_to_control is TRUE we move the most
# hostile intervention case (lowest score, which drags the intervention mean down) to
# control; otherwise we move the most hostile control case (highest score) to intervention.
iterative_condition_swap_once <- function(data, move_intervention_to_control) {

  data <- data |>
    group_by(condition) |>
    mutate(z_score = (score - mean(score)) / sd(score)) |>
    ungroup()

  if (move_intervention_to_control) {
    to_swap   <- data |> filter(condition == "intervention") |> slice_min(z_score, n = 1)
    new_label <- "control"
  } else {
    to_swap   <- data |> filter(condition == "control") |> slice_max(z_score, n = 1)
    new_label <- "intervention"
  }

  if (nrow(to_swap) < 1) {
    return(data |> dplyr::select(-z_score))
  }

  data |>
    mutate(condition = ifelse(row_number == to_swap$row_number, new_label, condition)) |>
    dplyr::select(-z_score)
}

# Main simulation function
run_iterative_condition_swap_study <- function(n_per_condition,
                                               mean_control,
                                               mean_intervention,
                                               sd_control,
                                               sd_intervention) {

  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) |>
    mutate(row_number = row_number())

  res <- analyse_data(data)
  p_val <- res$p[1]

  if(p_val < 0.05) {
    return(summarise_tamper_result(data, 0, TRUE))
  }

  n_total <- nrow(data)
  loop_guard <- tamper_loop_guard(n_total)
  swaps_so_far <- 0

  while(p_val >= 0.05 && swaps_so_far < loop_guard) {
    # alternate swap direction so group sizes stay balanced over pairs of swaps
    move_intervention_to_control <- (swaps_so_far %% 2 == 0)
    data <- iterative_condition_swap_once(data, move_intervention_to_control)
    swaps_so_far <- swaps_so_far + 1
    res <- analyse_data(data)
    p_val <- res$p[1]
  }

  return(summarise_tamper_result(data, swaps_so_far, !is.na(p_val) && p_val < 0.05))
}

```

## Run simulation

```{r}

simulation_iterative_swaps_meandiff_path <- cache_path("simulation_iterative_swaps_meandiff")
if (file.exists(simulation_iterative_swaps_meandiff_path)) {
  simulation_iterative_swaps_meandiff <- read_rds(simulation_iterative_swaps_meandiff_path)
} else {

  # Parameter grid
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_iterative_swaps_meandiff <- experiment_parameters_grid |>
    slice_sample(prop = 1) |> # shuffle rows so that the expensive large-n tasks are spread evenly across workers
    mutate(res = future_pmap(.l = list(n_per_condition = n_per_condition,
                                       mean_control = mean_control,
                                       mean_intervention = mean_intervention,
                                       sd_control = sd_control,
                                       sd_intervention = sd_intervention),
                             .f = run_iterative_condition_swap_study,
                             .progress = TRUE,
                             .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages))) |>
    unnest(res)

  write_rds(simulation_iterative_swaps_meandiff, simulation_iterative_swaps_meandiff_path)
}

# summarize tamper counts and bias (Hedges' g relative to mu = 0)
simulation_iterative_swaps_meandiff_tampers <- summarise_tampers(simulation_iterative_swaps_meandiff)
simulation_iterative_swaps_meandiff_bias    <- summarise_bias(simulation_iterative_swaps_meandiff)

```

## Table

```{r}

table_tampers(simulation_iterative_swaps_meandiff_tampers,
              caption = "Condition swapping: tampered data points needed to reach p < .05")

```

## Plot

```{r}

plot_distributions(simulation_iterative_swaps_meandiff, "Swaps", "N condition swaps")

p_iterative_swaps_meandiff_n <- plot_tampers_n(simulation_iterative_swaps_meandiff_tampers, "Swaps", "N condition swaps")

p_iterative_swaps_meandiff_prop <- plot_tampers_prop(simulation_iterative_swaps_meandiff_tampers, "Swaps", "N condition swaps")

```

## Bias (Hedges' g relative to μ = 0)

```{r}

table_bias(simulation_iterative_swaps_meandiff_bias,
           caption = "Condition swapping: effect size in the final tampered dataset (bias relative to a true effect of zero)")

p_iterative_swaps_meandiff_bias <- plot_bias(simulation_iterative_swaps_meandiff_bias, "Swaps")

p_iterative_swaps_meandiff_bias

```

# Duplication

From top 10% most favorable data points

## Simulation functions

```{r}

# Function to perform one round of top-10%-based duplication
duplicate_from_top_10_percent_once <- function(data, original_top10_control, original_top10_intervention) {
  
  # Sample one from each condition's original top 10%
  dup_control <- sample_n(original_top10_control, 1)
  dup_intervention <- sample_n(original_top10_intervention, 1)
  
  # Append to data
  data <- bind_rows(data, dup_control, dup_intervention)
  
  return(data)
}

# Main simulation function
run_iterative_duplication_study <- function(n_per_condition,
                                            mean_control,
                                            mean_intervention,
                                            sd_control,
                                            sd_intervention,
                                            proportion) {
  
  # Generate original data and label with row IDs
  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) %>%
    mutate(row_id = row_number())
  
  # Compute the original most-favourable 10% of scores per condition. "Favourable"
  # means favourable to a positive (intervention > control) effect, so it is the
  # LOWEST control scores and the HIGHEST intervention scores: duplicating these
  # drags the control mean down and the intervention mean up. This matches the
  # direction of every other strategy in this file, so the signed bias (Hedges' g)
  # is directly comparable across strategies.
  top10_control <- data %>%
    filter(condition == "control") %>%
    slice_min(score, prop = proportion)

  top10_intervention <- data %>%
    filter(condition == "intervention") %>%
    slice_max(score, prop = proportion)
  
  # Initial p-value
  p_val <- analyse_data(data)$p[1]
  if (p_val < 0.05) return(summarise_tamper_result(data, 0, TRUE))

  n_total <- nrow(data)
  loop_guard <- tamper_loop_guard(n_total)
  dups_so_far <- 0

  while (p_val >= 0.05 && dups_so_far < loop_guard) {
    data <- duplicate_from_top_10_percent_once(data, top10_control, top10_intervention)
    p_val <- analyse_data(data)$p[1]
    dups_so_far <- dups_so_far + 2
  }

  return(summarise_tamper_result(data, dups_so_far, !is.na(p_val) && p_val < 0.05))
}

```

## Run simulation

```{r}

simulation_iterative_duplication_meandiff_path <- cache_path("simulation_iterative_duplication_meandiff")
if (file.exists(simulation_iterative_duplication_meandiff_path)) {
  simulation_iterative_duplication_meandiff <- read_rds(simulation_iterative_duplication_meandiff_path)
} else {

  # Parameter grid
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    proportion = 0.1,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_iterative_duplication_meandiff <- experiment_parameters_grid %>%
    slice_sample(prop = 1) %>% # shuffle rows so that the expensive large-n tasks are spread evenly across workers
    mutate(res = future_pmap(.l = list(n_per_condition = n_per_condition,
                                       mean_control = mean_control,
                                       mean_intervention = mean_intervention,
                                       sd_control = sd_control,
                                       sd_intervention = sd_intervention,
                                       proportion = proportion),
                             .f = run_iterative_duplication_study,
                             .progress = TRUE,
                             .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages))) |>
    unnest(res)

  write_rds(simulation_iterative_duplication_meandiff, simulation_iterative_duplication_meandiff_path)
}

# summarise tamper counts and bias (Hedges' g relative to mu = 0)
simulation_iterative_duplication_meandiff_tampers <- summarise_tampers(simulation_iterative_duplication_meandiff)
simulation_iterative_duplication_meandiff_bias    <- summarise_bias(simulation_iterative_duplication_meandiff)

```

## Table

```{r}

table_tampers(simulation_iterative_duplication_meandiff_tampers,
              caption = "Duplication: tampered data points needed to reach p < .05")

```

## Plot

```{r}

plot_distributions(simulation_iterative_duplication_meandiff, "Duplications", "N duplications")

p_iterative_duplication_meandiff_n <- plot_tampers_n(simulation_iterative_duplication_meandiff_tampers, "Duplications", "N duplications")

p_iterative_duplication_meandiff_prop <- plot_tampers_prop(simulation_iterative_duplication_meandiff_tampers, "Duplications", "N duplications")

```

## Bias (Hedges' g relative to μ = 0)

```{r}

table_bias(simulation_iterative_duplication_meandiff_bias,
           caption = "Duplication: effect size in the final tampered dataset (bias relative to a true effect of zero)")

p_iterative_duplication_meandiff_bias <- plot_bias(simulation_iterative_duplication_meandiff_bias, "Duplications")

p_iterative_duplication_meandiff_bias

```

# Fabrication

## Illustrate 

for the correlation sim, m = .93 so that cohen's d = 5.0 == pearson's r = .93

```{r}

x_vals <- seq(-7, 7, length.out = 1000)

df <- tibble(
  x = rep(x_vals, 2),
  density = c(dnorm(x_vals, mean = 0, sd = 1),
              dnorm(x_vals, mean = 5, sd = 0.2)),
  distribution = rep(c("N(0,1)", "N(1,0.2)"), each = length(x_vals))
)

ggplot(df, aes(x = x, y = density, color = distribution)) +
  geom_line(linewidth = 1) +
  labs(x = "Value", y = "Density") +
  theme_linedraw()

```

## Simulation functions

```{r}

run_iterative_fabrication_study <- function(n_per_condition,
                                            mean_control,
                                            mean_intervention,
                                            sd_control,
                                            sd_intervention,
                                            fabricated_control_mean,
                                            fabricated_control_sd,
                                            fabricated_intervention_mean,
                                            fabricated_intervention_sd) {
  
  # Generate original data
  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) %>%
    mutate(row_id = row_number())
  
  # Initial p-value
  p_val <- analyse_data(data)$p[1]
  if (p_val < 0.05) return(summarise_tamper_result(data, 0, TRUE))

  n_total <- nrow(data)
  loop_guard <- tamper_loop_guard(n_total)
  adds_so_far <- 0

  while (p_val >= 0.05 && adds_so_far < loop_guard) {
    # Add one control and one intervention point per iteration
    new_control <- tibble(condition = "control", score = rnorm(1, mean = fabricated_control_mean, sd = fabricated_control_sd))
    new_intervention <- tibble(condition = "intervention", score = rnorm(1, mean = fabricated_intervention_mean, sd = fabricated_intervention_sd))

    data <- bind_rows(data, new_control, new_intervention)
    adds_so_far <- adds_so_far + 2

    p_val <- analyse_data(data)$p[1]
  }

  return(summarise_tamper_result(data, adds_so_far, !is.na(p_val) && p_val < 0.05))
}

```

## Run simulation

```{r}

simulation_iterative_fabrication_meandiff_path <- cache_path("simulation_iterative_fabrication_meandiff")
if (file.exists(simulation_iterative_fabrication_meandiff_path)) {
  simulation_iterative_fabrication_meandiff <- read_rds(simulation_iterative_fabrication_meandiff_path)
} else {

  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    fabricated_control_mean = 0,
    fabricated_control_sd = 0.2,
    fabricated_intervention_mean = 5,
    fabricated_intervention_sd = 0.2,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_iterative_fabrication_meandiff <- experiment_parameters_grid %>%
    slice_sample(prop = 1) %>% # shuffle rows so that the expensive large-n tasks are spread evenly across workers
    mutate(res = future_pmap(.l = list(n_per_condition = n_per_condition,
                                       mean_control = mean_control,
                                       mean_intervention = mean_intervention,
                                       sd_control = sd_control,
                                       sd_intervention = sd_intervention,
                                       fabricated_control_mean = fabricated_control_mean,
                                       fabricated_control_sd = fabricated_control_sd,
                                       fabricated_intervention_mean = fabricated_intervention_mean,
                                       fabricated_intervention_sd = fabricated_intervention_sd),
                             .f = run_iterative_fabrication_study,
                             .progress = TRUE,
                             .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages))) |>
    unnest(res)

  write_rds(simulation_iterative_fabrication_meandiff, simulation_iterative_fabrication_meandiff_path)
}

# summarise tamper counts and bias (Hedges' g relative to mu = 0)
simulation_iterative_fabrication_meandiff_tampers <- summarise_tampers(simulation_iterative_fabrication_meandiff)
simulation_iterative_fabrication_meandiff_bias    <- summarise_bias(simulation_iterative_fabrication_meandiff)

```

## Table

```{r}

table_tampers(simulation_iterative_fabrication_meandiff_tampers,
              caption = "Fabrication: tampered data points needed to reach p < .05")

```

## Plot

```{r}

plot_distributions(simulation_iterative_fabrication_meandiff, "Fabrications", "N fabricated")

p_iterative_fabrication_meandiff_n <- plot_tampers_n(simulation_iterative_fabrication_meandiff_tampers, "Fabrications", "N fabricated")

p_iterative_fabrication_meandiff_prop <- plot_tampers_prop(simulation_iterative_fabrication_meandiff_tampers, "Fabrications", "N fabricated")

```

## Bias (Hedges' g relative to μ = 0)

```{r}

table_bias(simulation_iterative_fabrication_meandiff_bias,
           caption = "Fabrication: effect size in the final tampered dataset (bias relative to a true effect of zero)")

p_iterative_fabrication_meandiff_bias <- plot_bias(simulation_iterative_fabrication_meandiff_bias, "Fabrications")

p_iterative_fabrication_meandiff_bias

```

# Replacement

## Simulation functions

```{r}

run_iterative_replacement_study <- function(n_per_condition,
                                            mean_control,
                                            mean_intervention,
                                            sd_control,
                                            sd_intervention,
                                            fabricated_control_mean,
                                            fabricated_control_sd,
                                            fabricated_intervention_mean,
                                            fabricated_intervention_sd) {
  
  # Generate original data
  data <- generate_data(n_per_condition,
                        mean_control,
                        mean_intervention,
                        sd_control,
                        sd_intervention) %>%
    mutate(row_id = row_number())
  
  # Initial p-value
  p_val <- analyse_data(data)$p[1]
  if (p_val < 0.05) return(summarise_tamper_result(data, 0, TRUE))

  n_total <- nrow(data)
  loop_guard <- tamper_loop_guard(n_total)
  replacements_so_far <- 0
  
  while (!is.na(p_val) && p_val >= 0.05 && replacements_so_far < loop_guard) {
    # Filter out NA scores
    data_incl <- data |> filter(!is.na(score))

    if (any(table(data_incl$condition) < 2)) break

    # True greedy on p itself, not on within-group z-scores. An earlier version
    # selected the case with the largest within-group |z|; once several control
    # cases had been replaced by fabricated values with SD 0.2, that group's SD
    # collapsed and its z-scores dominated the comparison, so the procedure kept
    # re-replacing near-identical control cases (which barely move the test) while
    # the original intervention cases -- whose replacement helps most -- were never
    # chosen. It then stalled below significance for hundreds of operations. Trying
    # both candidate replacements and keeping whichever yields the smaller p
    # implements the stated most-helpful-first rule directly, at the cost of one
    # extra t-test per step.
    # A residual stall remains by design: when the initial dataset shows a strong
    # effect in the NEGATIVE direction, one-step greedy cannot cross the two-sided
    # test's valley (weakening the negative effect raises p toward 1 before any
    # positive effect exists), and the negative-tail control cases are never
    # candidates. Such iterations (<1% at the smallest n) run to the loop guard and
    # are recorded as never attaining significance -- a true property of this naive
    # greedy procedure, not an artefact.
    cand_int <- data_incl |> filter(condition == "intervention") |> slice_min(score, n = 1, with_ties = FALSE)
    cand_ctl <- data_incl |> filter(condition == "control")      |> slice_max(score, n = 1, with_ties = FALSE)

    data_int <- data |> mutate(score = ifelse(row_id == cand_int$row_id,
                                              rnorm(1, fabricated_intervention_mean, fabricated_intervention_sd),
                                              score))
    data_ctl <- data |> mutate(score = ifelse(row_id == cand_ctl$row_id,
                                              rnorm(1, fabricated_control_mean, fabricated_control_sd),
                                              score))
    p_int <- analyse_data(data_int)$p[1]
    p_ctl <- analyse_data(data_ctl)$p[1]

    if (!is.na(p_int) && (is.na(p_ctl) || p_int <= p_ctl)) {
      data  <- data_int
      p_val <- p_int
    } else {
      data  <- data_ctl
      p_val <- p_ctl
    }
    replacements_so_far <- replacements_so_far + 1
  }

  return(summarise_tamper_result(data, replacements_so_far, !is.na(p_val) && p_val < 0.05))
}

```

## Run simulation

```{r}

simulation_iterative_replacement_meandiff_path <- cache_path("simulation_iterative_replacement_meandiff")
if (file.exists(simulation_iterative_replacement_meandiff_path)) {
  simulation_iterative_replacement_meandiff <- read_rds(simulation_iterative_replacement_meandiff_path)
} else {

  # Parameter grid
  experiment_parameters_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    mean_control = 0,
    mean_intervention = 0,
    sd_control = 1,
    sd_intervention = 1,
    fabricated_control_mean = 0,
    fabricated_control_sd = 0.2,
    fabricated_intervention_mean = 5,
    fabricated_intervention_sd = 0.2,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_iterative_replacement_meandiff <- experiment_parameters_grid %>%
    slice_sample(prop = 1) %>% # shuffle rows so that the expensive large-n tasks are spread evenly across workers
    mutate(res = future_pmap(.l = list(n_per_condition = n_per_condition,
                                       mean_control = mean_control,
                                       mean_intervention = mean_intervention,
                                       sd_control = sd_control,
                                       sd_intervention = sd_intervention,
                                       fabricated_control_mean = fabricated_control_mean,
                                       fabricated_control_sd = fabricated_control_sd,
                                       fabricated_intervention_mean = fabricated_intervention_mean,
                                       fabricated_intervention_sd = fabricated_intervention_sd),
                             .f = run_iterative_replacement_study,
                             .progress = TRUE,
                             .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages))) |>
    unnest(res)

  write_rds(simulation_iterative_replacement_meandiff, simulation_iterative_replacement_meandiff_path)
}

# Summarize tamper counts and bias (Hedges' g relative to mu = 0)
simulation_iterative_replacement_meandiff_tampers <- summarise_tampers(simulation_iterative_replacement_meandiff)
simulation_iterative_replacement_meandiff_bias    <- summarise_bias(simulation_iterative_replacement_meandiff)

```

## Table

```{r}

table_tampers(simulation_iterative_replacement_meandiff_tampers,
              caption = "Replacement: tampered data points needed to reach p < .05")

```

## Plot

```{r}

plot_distributions(simulation_iterative_replacement_meandiff, "Replacing", "N replaced")

p_iterative_replacement_meandiff_n <- plot_tampers_n(simulation_iterative_replacement_meandiff_tampers, "Replacment", "N replaced")

p_iterative_replacement_meandiff_prop <- plot_tampers_prop(simulation_iterative_replacement_meandiff_tampers, "Replacment", "N replaced")

```

## Bias (Hedges' g relative to μ = 0)

```{r}

table_bias(simulation_iterative_replacement_meandiff_bias,
           caption = "Replacement: effect size in the final tampered dataset (bias relative to a true effect of zero)")

p_iterative_replacement_meandiff_bias <- plot_bias(simulation_iterative_replacement_meandiff_bias, "Replacement")

p_iterative_replacement_meandiff_bias

```

# Combined tables and plots

## Tamper-count plots

```{r fig.height=8, fig.width=10}

simulation_iterative_meandiff_tampers <- bind_rows(
  simulation_iterative_dropping_meandiff_tampers |> mutate(Method = "Dropping"),
  simulation_iterative_swaps_meandiff_tampers |> mutate(Method = "Condition swapping"),
  simulation_iterative_duplication_meandiff_tampers |> mutate(Method = "Duplication"),
  simulation_iterative_fabrication_meandiff_tampers |> mutate(Method = "Fabrication"),
  simulation_iterative_replacement_meandiff_tampers |> mutate(Method = "Replacement")
) |>
  mutate(Method = fct_relevel(Method, method_levels))

p_meandiff_n <- 
  ggplot(simulation_iterative_meandiff_tampers, 
         aes(n_total, median, color = Method, fill = Method)) +
  geom_point(position = position_dodge(width = 27)) +
  geom_linerange(aes(ymin = pi_lower, ymax = ifelse(is.finite(pi_upper), pi_upper, NA_real_)), position = position_dodge(width = 27)) +
  labs(title = "Count of tampered cases needed to reach statistical significance",
       subtitle = "Median and 95% prediction interval across iterations; bars omitted where the upper percentile is never attained",
       x = "Sample size",
       y = "Count") +
  theme_linedraw() +
  scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
  scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2) +
  scale_color_viridis_d(begin = 0, end = 0.85) +
  scale_fill_viridis_d(begin = 0, end = 0.85)

p_meandiff_prop <- 
  ggplot(simulation_iterative_meandiff_tampers, 
         aes(n_total, prop_median, color = Method, fill = Method)) +
  geom_point(position = position_dodge(width = 27)) +
  geom_linerange(aes(ymin = prop_pi_lower, ymax = ifelse(is.finite(prop_pi_upper), prop_pi_upper, NA_real_)), position = position_dodge(width = 27)) +
  labs(title = "Proportion of tampered cases needed to reach statistical significance",
       subtitle = "Median and 95% prediction interval across iterations; bars omitted where the upper percentile is never attained",
       x = "Sample size",
       y = "Proportion") +
  theme_linedraw() +
  scale_y_continuous(labels = percent_format(), limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
  scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2) +
  scale_color_viridis_d(begin = 0, end = 0.85) +
  scale_fill_viridis_d(begin = 0, end = 0.85) 

p_meandiff_n +
  p_meandiff_prop +
  plot_layout(ncol = 1, guides = "collect")

ggsave(filename = "../communications/plots/iterative_meandiff_combined.pdf",
       device = "pdf",
       width = 9,
       height = 8)

ggsave(filename = "../communications/plots/iterative_meandiff_combined.png",
       device = "png",
       width = 9,
       height = 8)

```

## Bias plots

Hedges' g of each final, tampered dataset, quantified relative to the true standardised effect of μ = 0. Because the truth is 0, the **signed** median Hedges' g *is* the bias the tampering introduces, and its sign shows the direction each strategy induces. All five strategies are implemented to push toward a *positive* (intervention > control) effect, so all five should sit above zero; a strategy falling below it would indicate that its "most hostile case" rule had been specified in the wrong direction. An absolute |Hedges' g| would be misleading: it is strictly positive even under no systematic bias, so it overstates bias and hides direction.

```{r fig.height=6, fig.width=8}

# bind the raw, unnested per-iteration results across methods (with a Method label)
simulation_iterative_meandiff <- bind_rows(
  simulation_iterative_dropping_meandiff    |> mutate(Method = "Dropping"),
  simulation_iterative_swaps_meandiff       |> mutate(Method = "Condition swapping"),
  simulation_iterative_duplication_meandiff |> mutate(Method = "Duplication"),
  simulation_iterative_fabrication_meandiff |> mutate(Method = "Fabrication"),
  simulation_iterative_replacement_meandiff |> mutate(Method = "Replacement")
) |>
  mutate(Method = fct_relevel(Method, method_levels),
         n_total = n_per_condition * 2)

# combined bias summary across methods (uses the raw, unnested simulation results)
simulation_iterative_meandiff_bias <- simulation_iterative_meandiff |>
  group_by(Method, n_total) |>
  summarize(smd_bias_median            = median(hedges_g, na.rm = TRUE),
            smd_bias_median_mcse       = mcse_median(hedges_g),
            smd_bias_pi_lower          = quantile(hedges_g, probs = 0.025, na.rm = TRUE),
            smd_bias_pi_upper          = quantile(hedges_g, probs = 0.975, na.rm = TRUE),
            mean_diff_bias_median      = median(bias_mean_diff, na.rm = TRUE),
            mean_diff_bias_median_mcse = mcse_median(bias_mean_diff),
            mean_diff_bias_pi_lower    = quantile(bias_mean_diff, probs = 0.025, na.rm = TRUE),
            mean_diff_bias_pi_upper    = quantile(bias_mean_diff, probs = 0.975, na.rm = TRUE),
            .groups = "drop")

p_meandiff_bias <-
  ggplot(simulation_iterative_meandiff_bias,
         aes(n_total, smd_bias_median, color = Method, fill = Method)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_point(position = position_dodge(width = 27)) +
  geom_linerange(aes(ymin = smd_bias_pi_lower, ymax = smd_bias_pi_upper), position = position_dodge(width = 27)) +
  labs(title = "Bias of the tampered effect size relative to the truth (μ = 0)",
       subtitle = "Median and 95% prediction interval (2.5th-97.5th percentile) across iterations",
       x = "Sample size",
       y = "Hedges' g in final dataset") +
  theme_linedraw() +
  scale_y_continuous(breaks = breaks_pretty(n = 8)) +
  scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2) +
  scale_color_viridis_d(begin = 0, end = 0.85) +
  scale_fill_viridis_d(begin = 0, end = 0.85)

p_meandiff_bias

ggsave(filename = "../communications/plots/iterative_meandiff_bias.pdf",
       device = "pdf",
       width = 9,
       height = 5)

ggsave(filename = "../communications/plots/iterative_meandiff_bias.png",
       device = "png",
       width = 9,
       height = 5)

```

```{r}

table_bias(simulation_iterative_meandiff_bias,
           caption = "All strategies: effect size in the final tampered dataset (bias relative to a true effect of zero)", extra_cols = "Method")

```

## Table

```{r}

table_tampers(simulation_iterative_meandiff_tampers,
              caption = "All strategies: tampered data points needed to reach p < .05",
              extra_cols = "Method")

```

The mean and SD are reported here alongside the median, its interquartile range and its 95% prediction interval so that the skew is visible numerically: because the number of tampers needed is a right-skewed count, the mean sits above the median, and the gap between them is itself a summary of how long the upper tail is. The plots and the headline summaries throughout use the median — see the note under *A note on summarising with the median rather than the mean* for the rationale. Note that the interval in each table is the 95% prediction interval across iterations, whereas the value in parentheses is ±1 Monte Carlo standard error; they answer different questions and differ by an order of magnitude.

# Running on an HPC cluster

This report is written to run unchanged on a laptop and on a SLURM-managed HPC (e.g. the "Everything" cluster: SLURM + interactive RStudio Server), following the practices in `simulation_template_hpc.qmd`. The cluster scaffolding lives in the *Dependencies* and *Common functions and global variables* chunks:

- **Worker auto-detection.** The plan is set with `plan(multisession, workers = parallelly::availableCores())`. `availableCores()` honours the SLURM allocation, cgroup CPU quotas, and affinity masks, so the parallel plan adapts to the cores actually granted to the session rather than to all physical cores; pass an integer instead to pin a count. On Everything, 4–16 workers is the useful range (little gain beyond that for this workload).
- **BLAS thread pinning.** BLAS/LAPACK is pinned to one thread *before* the workers spawn, so dozens of worker processes do not oversubscribe the node. The matrices here are tiny, so single-threaded BLAS costs nothing.
- **Replication knob.** `run_full` / `on_hpc` auto-detect the environment (any `SLURM_*` variable present ⇒ cluster) and set `n_iterations` accordingly — a fast local smoke test (100) versus the full cluster run (10000). Force it with `run_full <- TRUE`/`FALSE`.
- **Tagged caches.** `cache_path()` tags every `.rds` with `dev`/`full` and the iteration count (e.g. `..._full_k10000.rds`), so local and cluster runs never overwrite each other. Re-rendering reads the cached summary instead of recomputing.
- **Reproducible parallel seeding.** `furrr_options(seed = TRUE)` advances an L'Ecuyer-CMRG stream per task, so results are identical regardless of the worker count; `scheduling = 2` enables dynamic load balancing, and `packages = worker_packages` attaches the needed libraries on each worker.

**Granularity (a deliberate deviation from the template).** The template parallelises over *conditions* and summarises inside each worker, so the master never holds raw replicate rows. These simulations instead parallelise over *replicates* (one task per condition × iteration) and keep the raw per-iteration results, because the distribution histograms and bootstrap-based median MCSEs need the full per-iteration vectors — summarising in the worker would discard exactly the data the report relies on. With many small tasks and `scheduling = 2`, the cost of the expensive large-*n* conditions is spread across workers automatically; each parameter grid is additionally shuffled with `slice_sample(prop = 1)` (after `set.seed()`, so the shuffle is itself reproducible) so that the expensive rows are not contiguous in the task list — the same load-balancing aid the template applies to condition-level tasks.

To run on Everything: request a node and launch RStudio Server inside the allocation, e.g. `salloc --nodes=1 --cpus-per-task=16 --mem=64G --time=08:00:00`; optionally export `FLEXIBLAS_NUM_THREADS=1 OMP_NUM_THREADS=1` in the shell before launching R (belt-and-braces; confirm with `RhpcBLASctl::blas_get_num_procs()`); then render. The full run regenerates the `*_full_k10000.rds` caches.

# Discussion

## Performance of the strategies across conditions

The substantive results are the per-strategy tables and the *Combined tables and plots* section; this discussion summarises how they should be read against the aims. Two patterns are the point and can be checked against the rendered figures:

- **Effort is small and roughly scale-invariant in proportional terms.** Although the absolute tamper count grows with sample size, the *proportion* of the sample that must be altered to reach significance is small and does not increase with N — manufacturing a false positive is no harder, proportionally, in a large study than a small one.
- **The distribution-preserving strategy is both effective and stealthy.** Condition swapping leaves every univariate distribution untouched, so it is the hardest to detect by routine checks, yet reaches significance with few operations. Strategies that add high-leverage points (fabrication, replacement) are the most efficient per operation but the most visible. Because the truth is zero, the median Hedges' *g* at the point significance is reached sits near the critical effect size for the given N — just large enough to clear the threshold and no larger, as the greedy procedure stops on first crossing.

## Conclusions with regard to the aims

The aims are met descriptively by the per-strategy tables (counts, proportions, MCSEs) and bias summaries, and comparatively by the combined plots. The headline conclusion is that the **distance from a null dataset to a "significant" one is short and cheap**, and that the line between practices labelled *p*-hacking and those labelled fabrication corresponds to a difference in keystrokes, not in effort or in the damage done to the evidence.

## Limitations and intended use

- **Proof of principle, not a fraud model.** The strategies are deliberately simple and greedy; real, detection-aware fraud would trade efficiency for plausibility, and the naive versions here can leave detectable signatures such as truncated distributions.
- **No detectability axis.** The study measures effort and bias, not how easily each manipulation would be caught; adding a detectability measure is the most valuable extension.
- **Normal, continuous DGP only.** Likert-type outcomes are not examined; an "alter existing values" strategy distinct from fabrication is left for future work.
- **Single threshold.** Results are conditioned on α = .05. No tampering budget is imposed in the design; budgets (10%, 25%, 50% of the sample) are applied post hoc at summary time, so results under any other budget are recoverable from the recorded counts.

# References

Chambers, C. (2017). *The seven deadly sins of psychology: A manifesto for reforming the culture of scientific practice.* Princeton University Press.

Morris, T. P., White, I. R., & Crowther, M. J. (2019). Using simulation studies to evaluate statistical methods. *Statistics in Medicine, 38*(11), 2074–2102. https://doi.org/10.1002/sim.8086

Siepe, B. S., Bartoš, F., Morris, T. P., Boulesteix, A.-L., Heck, D. W., & Pawel, S. (2024). Simulation studies for methodological research in psychology: A standardized template for planning, preregistration, and reporting. *Psychological Methods.* https://doi.org/10.1037/met0000695

*[citation to be completed]* Spiess et al. (2024) / the `dfstat`–`dfbeta` influential-observation literature.

# Session info

```{r}

sessionInfo()

```
