---
title: "Little big lies: iterative data tampering in 2x2 association (chi-square) designs"
author: "Ian Hussey"
date: today
format:
  html:
    embed-resources: true # ignored when rendering book
    toc: true
    toc_float: true
    code-fold: true
    code-overflow: wrap
    code-tools: true
    warning: false
    df-print: kable
    theme:
      light: cosmo
      dark: darkly
    highlight-style:
      light: breeze
      dark: dracula
    include-in-header:
      text: |
        <style>
          .kable-table table { width: 70%; margin: 0 auto; }
          .quarto-color-scheme-toggle::after { content: " Dark mode"; font-family: sans-serif; font-size: 0.85em; }
        </style>
---

# Introduction

This introduction is structured following the ADEMP framework (Aims, Data-generating process, Estimands, Methods, Performance) as described by Morris, White, and Crowther (2019) and operationalised for psychology by Siepe et al. (2024). Working through the framework explicitly makes the design decisions behind the simulations visible and auditable.

This file covers the **2×2 association (chi-square)** setting. The companion files `compendium_iterative_ttest.Rmd` and `compendium_iterative_correlation.Rmd` apply the same logic to a two-group mean difference and a Pearson correlation respectively. Because the outcome is binary the strategy menu is smaller than for continuous data — duplication, fabrication, and replacement collapse onto a single operation — so three distinct strategies are examined: the duplication/fabrication/replacement family, dropping, and condition switching.

## Project description

This is a simulation study of **data tampering and fabrication**. It asks: starting from a 2×2 table generated with no true association between group and outcome, how much must a dishonest analyst tamper with the data before a chi-square test crosses *p* < .05? Because the outcome is binary, the tampering strategies collapse: adding a favourable cell, duplicating one, or replacing an unfavourable one all reduce to **adding 1s to the exposed group or 0s to the unexposed group**. Quantifying the *effort* (the count of altered/added cells) makes concrete how little manipulation can be required.

## Prior related work

There is no closed-form benchmark for the minimal number of tampered cells needed to manufacture significance, so the study is **descriptive** rather than a calibration exercise. It connects to work on the fragility of findings to a few influential observations (the `dfstat`/`dfbeta` tradition; see also Spiess et al., 2024 *[citation to be completed]*) and to "how to cheat" framings such as Chambers' (2017) "dirty dozen".

## Aims

The task is the **characterisation of a tampering procedure** for a 2×2 table. For each strategy, sample size, and prevalence the aims are:

1. To estimate **how many tampered cells are required to reach *p* < .05** from null data, and how this scales with sample size and outcome prevalence, both absolutely and as a proportion of the sample.
2. To estimate the **proportion of datasets that never reach significance** within a pre-set tampering budget (cap).
3. To **compare** the effect of sample size and prevalence.

(No standardised effect-size bias is computed in this binary setting; the count of added cells is the primary quantity.)

## Data-Generating Process

### DGP specification approach

The DGP is **fully parametric** and generated under the **true null** (group independent of outcome), so any significant result is a false positive produced entirely by the tampering. Two groups (`exposed`, `not_exposed`) of `n_per_condition` observations each are drawn, with a binary outcome from the same Bernoulli prevalence in both groups. Produced by `generate_null_2x2_data()` via `rbinom()`.

### DGP factors

Three factors are relevant: the **tampering strategy**, the **per-group sample size** `n_per_condition`, and the **outcome prevalence**.

### Factor values and settings

- `n_per_condition` ∈ {10, 25, 50, 75, 100, 125, 150, 175, 200}.
- Prevalence ∈ {0.20, 0.05}, to contrast a common with a rare outcome.
- Tampering strategy ∈ {duplication/fabrication/replacement (a single operation for binary data), dropping, condition switching}.
- Held constant: equal prevalence in both groups (the null); the chi-square test without continuity correction; α = .05; a tampering budget of at most 50% of the sample.

### Factor combination and number of conditions

The grid crosses 3 strategies × 9 sample sizes × 2 prevalences = **54 conditions**, each replicated `n_iterations` times.

## Estimands and Targets

- **Primary target — tamper count.** `tampers_needed`, the number of fabricated cells required to first reach *p* < .05, summarised by its **median** (with IQR; rationale in *A note on summarising with the median rather than the mean*) and the **proportion never significant** within the cap.

There is no analytic ground truth; the Monte Carlo estimates are the result.

## Methods and extracted quantities

Each strategy generates a null 2×2 table, tests it, and while non-significant (and under the cap) applies one tampering operation that moves the table toward an exposed-high / not-exposed-low association, then re-tests with `chisq.test(correct = FALSE)`. The three operations (detailed in *Simulation functions*) are: **duplication/fabrication/replacement** — add a 1 to the exposed group or a 0 to the not-exposed group (a single operation for binary data); **dropping** — remove a 1 from the not-exposed group or a 0 from the exposed group (reduces N); and **condition switching** — move a not-exposed 1 into the exposed group or an exposed 0 into the not-exposed group, alternating direction to keep group sizes balanced (the only strategy that preserves total N). From each replication the worker returns `tampers_needed` and `reached_sig` (whether *p* < .05 was actually reached, so no replication is silently dropped). Note that none of these strategies preserves the marginal outcome prevalence. (There is no 2×2 analogue of the continuous "sign change" strategy.)

## Performance and Uncertainty

### Performance measures

The tamper-count distribution is summarised by its **median**, its interquartile range, and its **95% prediction interval** (the 2.5th and 97.5th percentiles, i.e. the range within which a single new study's tamper count would fall 95% of the time), with the mean and SD also tabulated to expose the skew, and as a **proportion of the sample**; the **empirical detection rate** within the tampering budget — equivalently its complement, the **proportion never significant** — is reported separately. Full rationale for the median is in *A note on summarising with the median rather than the mean*.

### Monte Carlo uncertainty

Every summary carries a Monte Carlo standard error: the binomial SE for the never-significant proportion, the SE of the mean for the mean count, and a nonparametric bootstrap (Morris et al., 2019) for the median count. See the *Monte Carlo standard errors* subsection below.

### Number of simulation repetitions

Each condition is replicated `n_iterations` times: **10000** for a full run and **100** for a fast local/development render (auto-selected by the `on_hpc` knob; see *Running on an HPC cluster*).

### Non-convergence and missing values

The chi-square test is closed-form, but a table with an empty margin (e.g. at very low prevalence and small N) yields an undefined statistic; this returns a missing *p*-value and is classified as never-significant via `reached_sig`; summaries use `na.rm`. Degenerate cells (no iteration reaching significance, or an undefined effect size for every iteration) are guarded in the summarising helpers, which return `NA` for that cell rather than aborting the render.

### Interpretation of performance measures *(optional)*

There is no pass/fail criterion; the comparison is relative. The expectation is that fewer fabricated cells are needed at larger N, and that a rare outcome (prevalence 0.05) is especially easy to tip, because a handful of fabricated cases moves the table disproportionately.

## Ethics, optics, and the *p*-hacking / fabrication continuum

This report documents weaponisable knowledge; the simulations make a continuum visible rather than providing a recipe. For a 2×2 table the manipulations are unusually stark: a single fabricated or relabelled cell can move the result, and the line between "excluding a participant" and "inventing one" is a single keystroke. The point is a trolley problem: is deleting data really ethically different from making it up, when both serve the individual's advancement at the expense of the evidence?

## Realism and scope

These are **proofs of principle**, not forensic models of real fraud. None of the strategies preserves the marginal outcome prevalence, which would be a giveaway; a detection-aware fraudster might constrain the operations to maintain plausibility. Out of scope: prevalence-preserving variants, an effect-size (φ / Cramér's V) bias measure, and a detectability axis.

# Dependencies

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

```{r}

library(tidyr)
library(dplyr)
library(readr)
library(forcats)
library(tibble)
library(purrr) 
library(furrr)
library(MASS)
library(effectsize)
library(ggplot2)
library(scales)
library(knitr)
library(kableExtra)
library(ggtext)
library(patchwork)
library(wesanderson)

options(scipen=999)

dir.create("../data", showWarnings = FALSE)
dir.create("../communications/plots", showWarnings = FALSE)

# --- Pin BLAS/LAPACK to one thread BEFORE workers spawn ---------------------
# With dozens of worker processes, multi-threaded BLAS would oversubscribe the
# node (workers x BLAS threads all fighting for cores). The operations here use
# tiny matrices, so single-threaded BLAS loses nothing. Setting these in the
# master means the multisession (PSOCK) workers inherit them at spawn.
Sys.setenv(FLEXIBLAS_NUM_THREADS = "1", OMP_NUM_THREADS = "1",
           OPENBLAS_NUM_THREADS = "1", MKL_NUM_THREADS = "1", BLIS_NUM_THREADS = "1",
           VECLIB_MAXIMUM_THREADS = "1") # VECLIB covers macOS Accelerate (local testing)
if (requireNamespace("flexiblas", quietly = TRUE)) flexiblas::flexiblas_set_num_threads(1L)
if (requireNamespace("RhpcBLASctl", quietly = TRUE)) {
  RhpcBLASctl::blas_set_num_threads(1)
  RhpcBLASctl::omp_set_num_threads(1)
}

# --- Parallel plan ----------------------------------------------------------
# parallelly::availableCores() honours a SLURM allocation, cgroup CPU quotas,
# and affinity masks, so it returns the cores granted to THIS session rather than
# all physical cores. RStudio Server is not fork-safe, so use multisession
# (separate R processes), not multicore. furrr_options(seed = TRUE) advances an
# L'Ecuyer-CMRG stream per task, giving reproducibility independent of the
# worker count.
plan(multisession, workers = parallelly::availableCores())

```

# Common functions and global variables

```{r}

# --- Replication knob (local vs cluster) ------------------------------------
# RStudio Server does not pass the job's SLURM_* allocation variables into the R
# session, but the HPC sets some SLURM_* global that survives and is absent on a
# laptop, so "any SLURM_* variable present" is a reliable cluster detector.
# run_full forces the choice: NULL = auto-detect, TRUE = full run, FALSE = quick
# local smoke test.
run_full <- TRUE
on_hpc <- if (!is.null(run_full)) isTRUE(run_full) else any(grepl("^SLURM_", names(Sys.getenv())))

# n iterations for all sims: full cluster run vs fast local smoke test
n_iterations <- if (on_hpc) 10000L else 100L

# Cache tag: keeps the full and dev runs in separate .rds files so they never
# collide. cache_path() builds a tagged path under ../data for each simulation.
cache_tag <- paste0(if (on_hpc) "full" else "dev", "_k", n_iterations)
cache_path <- function(stem) paste0("../data/", stem, "_", cache_tag, ".rds")

# Packages each worker must attach (belt-and-braces; future also auto-detects
# the packages of exported globals). MASS is only needed by the correlation sims.
worker_packages <- c("dplyr", "tidyr", "tibble", "purrr", "effectsize", "MASS")

# Display order of the tampering strategies ----
# As in the mean-difference and correlation reports, strategies are ordered by their
# impact: from the one needing the SMALLEST proportion of the sample to be tampered
# with to reach p < .05 to the one needing the largest (median proportion tampered,
# averaged over sample sizes and both prevalences). Condition switching precedes
# dropping here as it does in the other two reports.
method_levels <- c("Condition switching", "Dropping", "Duplication/fabrication/replacement")

# Monte Carlo standard errors (MCSEs) ----
# The mean and proportion MCSEs use the closed-form expressions from Siepe et al. (2024)
# / Morris et al. (2019), matching simulation_template_estimation.qmd:
#   MCSE(mean)       = SD / sqrt(K)
#   MCSE(proportion) = sqrt(p (1 - p) / K)
# The median has no neat closed form under skew, so its MCSE is obtained by the
# nonparametric bootstrap recommended by Morris et al. (2019). The RNG state is saved
# and restored so that computing summaries never perturbs the simulation seeds.
mcse_mean <- function(x){
  x <- x[!is.na(x)]
  if (length(x) < 2) return(NA_real_)
  sd(x) / sqrt(length(x))
}

mcse_proportion <- function(p, k){
  if (is.na(p) || k < 1) return(NA_real_)
  sqrt(p * (1 - p) / k)
}

mcse_median <- function(x, n_boot = 1000, seed = 42){
  x <- x[!is.na(x)]
  n <- length(x)
  if (n < 2) return(NA_real_)
  old_seed <- if (exists(".Random.seed", envir = .GlobalEnv)) get(".Random.seed", envir = .GlobalEnv) else NULL
  set.seed(seed)
  boot_medians <- replicate(n_boot, median(x[sample.int(n, n, replace = TRUE)]))
  if (!is.null(old_seed)) assign(".Random.seed", old_seed, envir = .GlobalEnv)
  # a censored median (Inf) has no estimable MCSE
  if (!all(is.finite(boot_medians))) return(NA_real_)
  sd(boot_medians)
}

# Guards and censoring-aware summaries ----
# Two things can go wrong in a single condition cell and take a whole render with them.
# (a) A degenerate cell: no iteration reached significance, or the effect size is
#     undefined for every iteration (an empty 2x2 margin at very low prevalence, a
#     constant variable, a group reduced below two cases). base R then warns and returns
#     +/-Inf from min()/max() and NaN from mean(). The safe_* wrappers return NA instead,
#     so the cell is reported as missing and every other cell still renders.
# (b) Right censoring: tampering stops at the budget (floor(0.5 * n_total)), so an
#     iteration that never reached p < .05 does not have a tamper count of "the cap", it
#     has a count of "at least the cap". Dropping those iterations and summarising the
#     survivors answers a different question ("given that tampering worked, how many
#     edits?") from the one this report asks ("starting from null data, how many edits?"),
#     and biases the answer downwards. Representing them as +Inf puts them in their
#     correct place at the top of the ordering: a quantile below the censoring point is
#     then estimated exactly, and one above it returns Inf -- reported as "> budget"
#     rather than silently reported as the survivors' quantile.
safe_min  <- function(x){ x <- x[is.finite(x)]; if (length(x) == 0) NA_real_ else min(x) }
safe_max  <- function(x){ x <- x[is.finite(x)]; if (length(x) == 0) NA_real_ else max(x) }
safe_mean <- function(x){ x <- x[is.finite(x)]; if (length(x) == 0) NA_real_ else mean(x) }
safe_sd   <- function(x){ x <- x[is.finite(x)]; if (length(x) < 2)  NA_real_ else sd(x) }

# quantile of a possibly right-censored vector (censored observations coded as Inf)
q_censored <- function(x, probs){
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_real_)
  unname(quantile(x, probs = probs, names = FALSE, type = 7))
}

median_censored <- function(x) q_censored(x, 0.5)

# median absolute deviation of a censored vector; NA when the median itself is censored
mad_censored <- function(x){
  x <- x[!is.na(x)]
  m <- q_censored(x, 0.5)
  if (length(x) == 0 || !is.finite(m)) return(NA_real_)
  1.4826 * q_censored(abs(x - m), 0.5)
}

# Table formatting ----
# Following simulation_template_lego.qmd, a point estimate is printed with its Monte
# Carlo standard error in parentheses and an interval in square brackets. The number of
# decimal places is driven by the MCSE (Morris et al., 2019): the MCSE is shown to two
# significant figures and the estimate to the same number of places. Rounding the whole
# table to a fixed 2 dp -- as an earlier version did -- erased exactly the resolution the
# substantive claims rest on: median proportions of .035, .028 and .023 all printed as
# "0.04"/"0.03"/"0.02", and every proportion MCSE printed as "0.00".
dp_from_mcse <- function(mcse, min_dp = 0, max_dp = 4){
  mcse <- mcse[is.finite(mcse) & mcse > 0]
  if (length(mcse) == 0) return(min_dp)
  min(max_dp, max(min_dp, max(ceiling(-log10(mcse)) + 1)))
}

# A censored quantile is Inf; it is printed as "> budget" rather than as a number,
# because all that is known is that it lies beyond the 50% tampering budget.
fmt_num <- function(x, dp){
  ifelse(is.na(x), "NA",
         ifelse(is.finite(x), formatC(x, format = "f", digits = dp), "> budget"))
}

mcse_cell <- function(estimate, mcse, min_dp = 0, max_dp = 4){
  dp <- dp_from_mcse(mcse, min_dp = min_dp, max_dp = max_dp)
  ifelse(is.finite(estimate) & is.finite(mcse),
         paste0(fmt_num(estimate, dp), " (", fmt_num(mcse, dp), ")"),
         fmt_num(estimate, dp))
}

interval_cell <- function(lower, upper, dp = 0){
  paste0("[", fmt_num(lower, dp), ", ", fmt_num(upper, dp), "]")
}

tampers_note <- paste(
  "Values in parentheses are +/-1 Monte Carlo standard error.",
  "IQR is the 25th-75th percentile and PI the 95% prediction interval (2.5th-97.5th",
  "percentile) of the results across iterations; neither is an MCSE, and both describe",
  "the spread of the simulated studies rather than the precision of the summary.",
  "The median, IQR and PI are computed over all iterations, with those that never reached",
  "significance ranked above the tampering budget; '> budget' means that percentile lies",
  "beyond the budget and so is not attainable within it.",
  "The mean and SD cannot be computed that way (the censored counts are unbounded above),",
  "so they are conditional on reaching significance: they are downward-biased for the",
  "unconditional mean whenever 'Never sig.' exceeds 0, and should be read as a within-cell",
  "skew diagnostic rather than compared across cells.",
  "The 0th/100th percentiles (min/max) are held in the summary object but not printed:",
  "they drift with the number of iterations and so are not stable quantities."
)

bias_note <- paste(
  "Values in parentheses are +/-1 Monte Carlo standard error (nonparametric bootstrap",
  "for medians). PI is the 95% prediction interval (2.5th-97.5th percentile) of the",
  "effect sizes across iterations. Because the true effect is zero, the signed median",
  "effect size in the final tampered dataset is itself the bias."
)

# Tamper-count table. extra_cols names any grouping columns (e.g. "Method") that should
# be printed before the sample size.
table_tampers <- function(data, caption, extra_cols = NULL){
  data |>
    transmute(across(all_of(extra_cols)),
              `N total`               = n_total,
              `Never sig.`            = mcse_cell(proportion_never_sig, proportion_never_sig_mcse, min_dp = 3),
              `Median tampers`        = mcse_cell(median, median_mcse),
              `IQR`                   = interval_cell(iqr_lower, iqr_upper),
              `95% PI`                = interval_cell(pi_lower, pi_upper),
              `Mean (reached only)`   = mcse_cell(mean_reached, mean_reached_mcse),
              `SD (reached only)`     = fmt_num(sd_reached, 2),
              `Median prop. tampered` = mcse_cell(prop_median, prop_median_mcse, min_dp = 3),
              `95% PI (proportion)`   = interval_cell(prop_pi_lower, prop_pi_upper, dp = 3)) |>
    kable(caption = caption) |>
    kable_classic(full_width = FALSE) |>
    footnote(general = tampers_note, general_title = "Note.", footnote_as_chunk = TRUE)
}

summarise_tampers <- function(data){
  data |>
    # right-censor: an iteration that hit the budget without reaching p < .05 has a
    # tamper count of "at least the cap", coded as Inf (see the guards above)
    mutate(tampers_censored = ifelse(reached_sig, tampers_needed, Inf)) |>
    group_by(n_per_condition) |>
    summarize(k = n(), # total iterations in this cell
              k_reached = sum(reached_sig), # iterations that reached significance
              # Empirical detection rate: the proportion of null datasets pushed over
              # p < .05 within the tampering budget. This is the standard Monte Carlo
              # name for the quantity; proportion_never_sig is its complement (the two
              # sum to 1 and share the same binomial MCSE) and is kept because it is the
              # form named as an estimand in the ADEMP sections above. It is also the
              # censoring fraction, and so says how far the summaries below can be trusted.
              detection_rate = mean(reached_sig),
              detection_rate_mcse = mcse_proportion(detection_rate, k),
              proportion_never_sig = 1 - detection_rate,
              proportion_never_sig_mcse = detection_rate_mcse,
              # --- headline summaries: computed over ALL k iterations, censoring-aware
              median = median_censored(tampers_censored),
              median_mcse = mcse_median(tampers_censored),
              mad = mad_censored(tampers_censored),
              # the interquartile range (a 50% interval) and the 95% prediction interval
              # (2.5th-97.5th percentile), i.e. the range within which a single new study's
              # tamper count would fall 95% of the time. Named so that neither can be
              # mistaken for the other or for the +/-1 MCSE of the median. Either bound is
              # Inf if that percentile lies beyond the tampering budget.
              iqr_lower = q_censored(tampers_censored, 0.25),
              iqr_upper = q_censored(tampers_censored, 0.75),
              pi_lower  = q_censored(tampers_censored, 0.025),
              pi_upper  = q_censored(tampers_censored, 0.975),
              # --- conditional summaries: over the iterations that DID reach significance.
              # The mean is not estimable over all k under censoring (the censored values
              # are unbounded above), so the mean and SD here are conditional on reaching
              # significance and are downward-biased for the unconditional mean whenever
              # any censoring occurred. They are reported as a within-cell skew diagnostic
              # against median_reached, not as a headline summary, and must not be compared
              # across cells with different censoring fractions. min/max are the 0th and
              # 100th percentiles: unstable (the max is the largest of k draws, so it drifts
              # upward as k grows), so retained for diagnostics but neither tabled nor plotted.
              median_reached = median_censored(tampers_needed[reached_sig]),
              mean_reached = safe_mean(tampers_needed[reached_sig]),
              mean_reached_mcse = mcse_mean(tampers_needed[reached_sig]),
              sd_reached = safe_sd(tampers_needed[reached_sig]),
              min_reached = safe_min(tampers_needed[reached_sig]),
              max_reached = safe_max(tampers_needed[reached_sig])) |>
    ungroup() |>
    mutate(n_total = n_per_condition*2,
           # the tampering budget itself, used to truncate the plotted intervals
           cap = floor(0.5 * n_total),
           prop_median = median/n_total,
           prop_median_mcse = median_mcse/n_total,
           prop_mad = mad/n_total,
           prop_iqr_lower = iqr_lower/n_total,
           prop_iqr_upper = iqr_upper/n_total,
           prop_pi_lower  = pi_lower/n_total,
           prop_pi_upper  = pi_upper/n_total,
           prop_mean_reached = mean_reached/n_total,
           prop_mean_reached_mcse = mean_reached_mcse/n_total,
           prop_sd_reached = sd_reached/n_total,
           prop_min_reached = min_reached/n_total,
           prop_max_reached = max_reached/n_total) |>
    relocate(n_total, .before = n_per_condition) |>
    dplyr::select(-n_per_condition)
}

plot_distributions <- function(data, title, x_axis){
  ggplot(data, aes(x = tampers_needed)) +
    geom_histogram(binwidth = 1) +
    labs(title = title, 
         x = x_axis,
         y = "Count") +
    ylim(0, NA) +
    theme_linedraw() +
    scale_x_continuous(breaks = breaks_pretty(n = 5)) +
    facet_wrap(~ n_per_condition) # N PER CONDITION NOT TOTAL N
}

plot_tampers_n <- function(data, title, y_axis){
  ggplot(data, aes(n_total, median)) +
    geom_point() +
    # a censored bound (Inf) is drawn at the budget, which is the largest attainable
    # value; the true bound lies at or beyond it
    geom_linerange(aes(ymin = pi_lower, ymax = pmin(pi_upper, cap))) +
    labs(title = title,
         subtitle = "Median and 95% prediction interval across iterations, truncated at the 50% tampering budget",
         x = "Sample size",
         y = y_axis) +
    theme_linedraw() +
    scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
    scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2)
}

plot_tampers_prop <- function(data, title, y_axis){
  ggplot(data, aes(n_total, prop_median)) +
    geom_point() +
    geom_linerange(aes(ymin = prop_pi_lower, ymax = pmin(prop_pi_upper, 0.5))) +
    labs(title = title,
         subtitle = "Median and 95% prediction interval across iterations, truncated at the 50% tampering budget",
         x = "Sample size",
         y = y_axis) +
    theme_linedraw() +
    scale_y_continuous(labels = percent_format(), limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
    scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2)
}

```

# A note on summarising with the median rather than the mean

Each simulation records `tampers_needed`: the number of tampering operations applied before the result crossed *p* < .05. This quantity is summarised throughout with the **median** (with the interquartile range and the 95% prediction interval) rather than the mean. There are three reasons.

1. **The distribution is right-skewed.** Most datasets reach significance after only a few tampers, but a long tail require many. Part of the point of these simulations is to *show* that skew, so the central-tendency measure should not hide it. The summary tables report `mean_reached` and `sd_reached` alongside the median, its interquartile range and its 95% prediction interval, precisely so the skew is visible: the mean sits above the median, and the size of that gap indexes the length of the tail. Read that gap only within a condition, and only where no censoring has occurred — where `proportion_never_sig` is above zero the two statistics are computed over different sets of iterations (all of them for the median, the uncensored ones for the mean), so the comparison no longer isolates skew. `median_reached` is carried in the summary object for anyone who wants the like-for-like comparison.

2. **The distribution is censored, and this is the decisive reason.** Tampering stops at a budget (`floor(0.5 × n_total)`), and in some conditions a substantial proportion of datasets never reach significance within it (reported as `proportion_never_sig`, which is also the censoring fraction). Those iterations do not have a tamper count of "the budget" — they have a count of *at least* the budget, with no upper bound. Under such right censoring:

   - The **median, IQR and prediction interval are computed over all iterations**, with the censored ones ranked above the budget (coded as `Inf`). Any percentile that falls below the censoring point is then estimated exactly and is completely unaffected by where the censored values would have landed; any percentile above it is reported as `> budget` rather than being silently replaced by the corresponding percentile of the survivors. This matters: summarising only the iterations that reached significance answers "*given* that tampering worked within the budget, how many edits?", whereas the question this report asks is "starting from null data, how many edits?".
   - The **mean cannot be computed that way at all**, because the censored counts are unbounded above. The mean and SD reported here are therefore *conditional on reaching significance* (`mean_reached`, `sd_reached`), and are downward-biased estimates of the unconditional mean by construction, since they discard precisely the largest values. The size of that bias grows with the censoring fraction, so they must not be compared across conditions with different `proportion_never_sig`.

3. **The median refers to an actual case.** For an integer count, a mean such as "2.7 tampers" corresponds to no real study, whereas a median of 2 describes a realisable scenario ("half the time, two edits were enough"). Because the substantive claim concerns the realistic effort an individual would expend, the statistic that points at an actual case is the more faithful one.

Note that the central-tendency statistics **include** the datasets that were significant from the outset and so required zero tampers (the ~5% expected under the null at α = .05). The median therefore answers "starting from null data, how many edits to significance?" rather than "*given* that tampering was needed, how many?". This framing was chosen deliberately because it matches the question of how readily null data can be pushed to significance.

## Monte Carlo standard errors

Because each summary statistic is itself estimated from a finite number of simulation iterations (`n_iterations`), it carries Monte Carlo error. Following the approach in `simulation_template_estimation.qmd` (and Siepe et al., 2024; Morris et al., 2019), every table reports a Monte Carlo standard error (MCSE) next to each estimated quantity, so that apparent differences between conditions can be judged against simulation noise. The MCSEs are computed as:

- **Proportion never significant** (`proportion_never_sig_mcse`): the binomial standard error $\sqrt{\hat p\,(1-\hat p)/K}$, where *K* = `n_iterations`.
- **Mean** (`mean_mcse`): the standard error of the mean $\text{SD}/\sqrt{K_\text{reached}}$, where $K_\text{reached}$ is the number of iterations that reached significance before the cap (i.e. the iterations the mean is computed over).
- **Median** (`median_mcse`): obtained by nonparametric bootstrap (1,000 resamples), because the count distribution is skewed and has no convenient closed-form quantile MCSE. The same bootstrap is used for the median effect-size (bias) columns.

Two things are deliberately *not* MCSEs and are never plotted as such. The **interquartile range** (`iqr_lower`/`iqr_upper`) and the **95% prediction interval** (`pi_lower`/`pi_upper`, the 2.5th and 97.5th percentiles) describe the spread of the simulated studies themselves, not the precision with which the summary is estimated; the prediction interval is the wider and is what the plots show. The 0th and 100th percentiles (`min`/`max`) are computed and retained in the summary objects but are neither tabled nor plotted, because the maximum of *K* draws drifts upward as *K* grows and so is not a stable quantity.

A caveat specific to the median of an integer count: at *K* = 10,000 the bootstrap MCSE is frequently exactly 0, because every resample returns the same integer. That is honest rather than an error — the median is pinned — but it means the MCSE column is uninformative for those cells, and it is one respect in which the mean (whose MCSE is smooth and always positive) is better behaved. It is not a reason to prefer the mean as the headline summary, for the censoring reason given above. A censored median has no estimable MCSE at all and is reported as `NA`.

Tables print each estimate with its MCSE in parentheses, following `simulation_template_lego.qmd`. The number of decimal places is set from the MCSE rather than fixed for the whole table: the MCSE is shown to two significant figures and the estimate to the same number of places (Morris et al., 2019). This matters here because the headline quantity — the proportion of the sample that must be tampered with — falls between .02 and .04 across the larger sample sizes, and a fixed 2 dp collapsed those distinct values onto the same printed number.

The `median_mcse` is the relevant uncertainty for the headline statistic; with `n_iterations` in the low thousands it is typically a small fraction of a tamper, confirming the medians are estimated precisely. Because the tamper counts are discrete, the bootstrap median MCSE is sometimes exactly **0**: this is not an error but indicates that the median is invariant to resampling — its sampling variability is below the one-unit resolution of the data, so the median is estimated to sub-count precision. (The bootstrap saves and restores the RNG state, so it never disturbs the simulation seeds.) The `mean_mcse`, which is always non-zero, provides a complementary, continuous-scale check on Monte Carlo precision in these cases.

# Simulation functions

For binary outcomes the menu of tampering strategies is smaller than for continuous data, because several labels collapse onto the same operation:

- **Duplication / fabrication / replacement.** With a 0/1 outcome there is no value to perturb, only cells to add: duplicating a favourable case, fabricating a new favourable case, and replacing an unfavourable case with a favourable one are all the same atomic move — **add a 1 to the exposed group or a 0 to the not-exposed group**. They are therefore treated as a single strategy here, labelled *Duplication/fabrication/replacement*. (For the same reason the continuous "sign change" strategy has no 2×2 analogue.)
- **Dropping.** The mirror of the above: **drop a 1 from the not-exposed group or a 0 from the exposed group** — i.e. remove the cases that work against the association. This reduces N.
- **Condition switching.** **Move a case between groups** (a not-exposed 1 into exposed, or an exposed 0 into not-exposed), keeping the outcome fixed. This is equivalent to one fabrication plus one drop, and is the only 2×2 strategy that preserves the total N (the binary analogue of the *t*-test condition swap and the correlation re-pairing).

All three push the table toward an exposed-high / not-exposed-low association. None of them preserves the marginal outcome prevalence, which a detection-aware fraudster might wish to maintain.

```{r}

# Generate initial 2x2 data under the null (group independent of outcome) ----
generate_null_2x2_data <- function(n_per_condition, prevalence) {
  tibble(
    group = rep(c("exposed", "not_exposed"), each = n_per_condition),
    outcome = c(rbinom(n_per_condition, 1, prevalence), rbinom(n_per_condition, 1, prevalence))
  )
}

# Chi-square test for a 2x2 table (no Yates correction) ----
# Robust to degenerate intermediate tables: factors fix the 2x2 shape, and a table with
# an empty row/column margin (undefined statistic) returns NA rather than erroring.
analyse_chisq <- function(data) {
  tbl <- table(factor(data$group,   levels = c("exposed", "not_exposed")),
               factor(data$outcome, levels = c(0, 1)))
  if (any(rowSums(tbl) == 0) || any(colSums(tbl) == 0)) return(tibble(p = NA_real_))
  res <- tryCatch(suppressWarnings(chisq.test(tbl, correct = FALSE)), error = function(e) NULL)
  if (is.null(res)) return(tibble(p = NA_real_))
  tibble(p = res$p.value)
}

```

# Duplication/fabrication/replacement

For binary data these three strategies coincide (see *Simulation functions* above): each round adds one favourable cell — a 1 to the exposed group or a 0 to the not-exposed group.

## Simulation functions

```{r}

# add a 1 to exposed or a 0 to not_exposed
run_iterative_fabrication_2x2 <- function(n_per_condition, prevalence, max_add_prop) {
  data <- generate_null_2x2_data(n_per_condition, prevalence)

  p_val <- analyse_chisq(data)$p[1]
  if (!is.na(p_val) && p_val < 0.05) return(tibble(tampers_needed = 0, reached_sig = TRUE))

  n_total <- nrow(data)
  max_adds <- floor(max_add_prop * n_total)
  adds_so_far <- 0

  while (!is.na(p_val) && p_val >= 0.05 && adds_so_far < max_adds) {
    fabricated_case <- tibble(group = "exposed", outcome = 1)
    fabricated_noncase <- tibble(group = "not_exposed", outcome = 0)

    # Randomly choose which to add
    new_row <- if (runif(1) < 0.5) fabricated_case else fabricated_noncase

    data <- bind_rows(data, new_row)
    adds_so_far <- adds_so_far + 1

    p_val <- analyse_chisq(data)$p[1]
  }

  return(tibble(tampers_needed = adds_so_far, reached_sig = !is.na(p_val) && p_val < 0.05))
}

```

## Prevalence = .20

### Run simulation

```{r}

simulation_iterative_fabrication_2x2_path <- cache_path("simulation_iterative_fabrication_2x2")
if (file.exists(simulation_iterative_fabrication_2x2_path)) {
  simulation_iterative_fabrication_2x2 <- read_rds(simulation_iterative_fabrication_2x2_path)
} else {

  experiment_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    prevalence = 0.20,
    max_add_prop = 0.5,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_iterative_fabrication_2x2 <- experiment_grid %>%
    slice_sample(prop = 1) %>% # shuffle rows so that the expensive large-n tasks are spread evenly across workers
    mutate(res = future_pmap(
      .l = list(n_per_condition = n_per_condition,
                prevalence = prevalence,
                max_add_prop = max_add_prop),
      .f = run_iterative_fabrication_2x2,
      .progress = TRUE,
      .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages)
    )) |>
    unnest(res)

  write_rds(simulation_iterative_fabrication_2x2, simulation_iterative_fabrication_2x2_path)
}

simulation_fabrication_2x2_tampers <- summarise_tampers(simulation_iterative_fabrication_2x2)

```

### Table

```{r}

table_tampers(simulation_fabrication_2x2_tampers,
              caption = "Duplication/fabrication/replacement, prevalence = 0.20: tampered data points needed to reach p < .05")

```

### Plot

```{r}

plot_tampers_n(simulation_fabrication_2x2_tampers, "Duplication/fabrication/replacement", "N added")

plot_tampers_prop(simulation_fabrication_2x2_tampers, "Duplication/fabrication/replacement", "Proportion added")

```

## Prevalence = .05

### Run simulation

```{r}

simulation_fabrication_2x2_lowprev_path <- cache_path("simulation_iterative_fabrication_2x2_lowprev")
if (file.exists(simulation_fabrication_2x2_lowprev_path)) {
  simulation_fabrication_2x2_lowprev <- read_rds(simulation_fabrication_2x2_lowprev_path)
} else {

  experiment_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    prevalence = 0.05,
    max_add_prop = 0.5,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_fabrication_2x2_lowprev <- experiment_grid %>%
    slice_sample(prop = 1) %>% # shuffle rows so that the expensive large-n tasks are spread evenly across workers
    mutate(res = future_pmap(
      .l = list(n_per_condition = n_per_condition,
                prevalence = prevalence,
                max_add_prop = max_add_prop),
      .f = run_iterative_fabrication_2x2,
      .progress = TRUE,
      .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages)
    )) |>
    unnest(res)

  write_rds(simulation_fabrication_2x2_lowprev, simulation_fabrication_2x2_lowprev_path)
}

simulation_fabrication_2x2_lowprev_tampers <- summarise_tampers(simulation_fabrication_2x2_lowprev)

```

### Table

```{r}

table_tampers(simulation_fabrication_2x2_lowprev_tampers,
              caption = "Duplication/fabrication/replacement, prevalence = 0.05: tampered data points needed to reach p < .05")

```

### Plot

```{r}

plot_tampers_n(simulation_fabrication_2x2_lowprev_tampers, "Duplication/fabrication/replacement", "N added")

plot_tampers_prop(simulation_fabrication_2x2_lowprev_tampers, "Duplication/fabrication/replacement", "Proportion added")

```

# Dropping

The mirror of fabrication: each round drops one case that works against the association — a 1 from the not-exposed group or a 0 from the exposed group. This reduces N.

## Simulation functions

```{r}

# drop a 1 from not_exposed or a 0 from exposed
run_iterative_dropping_2x2 <- function(n_per_condition, prevalence, max_drop_prop) {
  data <- generate_null_2x2_data(n_per_condition, prevalence)

  p_val <- analyse_chisq(data)$p[1]
  if (!is.na(p_val) && p_val < 0.05) return(tibble(tampers_needed = 0, reached_sig = TRUE))

  n_total <- nrow(data)
  max_drops <- floor(max_drop_prop * n_total)
  drops_so_far <- 0

  while (!is.na(p_val) && p_val >= 0.05 && drops_so_far < max_drops) {
    # hostile cells: a 1 in not_exposed, or a 0 in exposed
    drop_from_notexposed <- runif(1) < 0.5
    idx <- if (drop_from_notexposed) which(data$group == "not_exposed" & data$outcome == 1) else which(data$group == "exposed" & data$outcome == 0)

    if (length(idx) == 0) {
      # the chosen hostile cell is empty; fall back to the other one
      idx <- if (drop_from_notexposed) which(data$group == "exposed" & data$outcome == 0) else which(data$group == "not_exposed" & data$outcome == 1)
      if (length(idx) == 0) break
    }

    data <- data[-idx[1], ]
    drops_so_far <- drops_so_far + 1

    # stop if a whole group has been emptied
    if (any(table(factor(data$group, levels = c("exposed", "not_exposed"))) < 1)) break

    p_val <- analyse_chisq(data)$p[1]
  }

  return(tibble(tampers_needed = drops_so_far, reached_sig = !is.na(p_val) && p_val < 0.05))
}

```

## Prevalence = .20

### Run simulation

```{r}

simulation_iterative_dropping_2x2_path <- cache_path("simulation_iterative_dropping_2x2")
if (file.exists(simulation_iterative_dropping_2x2_path)) {
  simulation_iterative_dropping_2x2 <- read_rds(simulation_iterative_dropping_2x2_path)
} else {

  experiment_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    prevalence = 0.20,
    max_drop_prop = 0.5,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_iterative_dropping_2x2 <- experiment_grid %>%
    slice_sample(prop = 1) %>% # shuffle rows so that the expensive large-n tasks are spread evenly across workers
    mutate(res = future_pmap(
      .l = list(n_per_condition = n_per_condition,
                prevalence = prevalence,
                max_drop_prop = max_drop_prop),
      .f = run_iterative_dropping_2x2,
      .progress = TRUE,
      .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages)
    )) |>
    unnest(res)

  write_rds(simulation_iterative_dropping_2x2, simulation_iterative_dropping_2x2_path)
}

simulation_dropping_2x2_tampers <- summarise_tampers(simulation_iterative_dropping_2x2)

```

### Table

```{r}

table_tampers(simulation_dropping_2x2_tampers,
              caption = "Dropping, prevalence = 0.20: tampered data points needed to reach p < .05")

```

### Plot

```{r}

plot_tampers_n(simulation_dropping_2x2_tampers, "Dropping", "N dropped")

plot_tampers_prop(simulation_dropping_2x2_tampers, "Dropping", "Proportion dropped")

```

## Prevalence = .05

### Run simulation

```{r}

simulation_dropping_2x2_lowprev_path <- cache_path("simulation_iterative_dropping_2x2_lowprev")
if (file.exists(simulation_dropping_2x2_lowprev_path)) {
  simulation_dropping_2x2_lowprev <- read_rds(simulation_dropping_2x2_lowprev_path)
} else {

  experiment_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    prevalence = 0.05,
    max_drop_prop = 0.5,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_dropping_2x2_lowprev <- experiment_grid %>%
    slice_sample(prop = 1) %>% # shuffle rows so that the expensive large-n tasks are spread evenly across workers
    mutate(res = future_pmap(
      .l = list(n_per_condition = n_per_condition,
                prevalence = prevalence,
                max_drop_prop = max_drop_prop),
      .f = run_iterative_dropping_2x2,
      .progress = TRUE,
      .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages)
    )) |>
    unnest(res)

  write_rds(simulation_dropping_2x2_lowprev, simulation_dropping_2x2_lowprev_path)
}

simulation_dropping_2x2_lowprev_tampers <- summarise_tampers(simulation_dropping_2x2_lowprev)

```

### Table

```{r}

table_tampers(simulation_dropping_2x2_lowprev_tampers,
              caption = "Dropping, prevalence = 0.05: tampered data points needed to reach p < .05")

```

### Plot

```{r}

plot_tampers_n(simulation_dropping_2x2_lowprev_tampers, "Dropping", "N dropped")

plot_tampers_prop(simulation_dropping_2x2_lowprev_tampers, "Dropping", "Proportion dropped")

```

# Condition switching

Each round moves one case between groups, keeping its outcome — a not-exposed 1 into the exposed group, or an exposed 0 into the not-exposed group. The direction alternates so the two group sizes stay balanced over pairs of switches; this is the only 2×2 strategy that preserves the total N.

## Simulation functions

```{r}

# move a not_exposed 1 into exposed (even step) or an exposed 0 into not_exposed (odd step)
run_iterative_switching_2x2 <- function(n_per_condition, prevalence, max_switch_prop) {
  data <- generate_null_2x2_data(n_per_condition, prevalence)

  p_val <- analyse_chisq(data)$p[1]
  if (!is.na(p_val) && p_val < 0.05) return(tibble(tampers_needed = 0, reached_sig = TRUE))

  n_total <- nrow(data)
  max_switches <- floor(max_switch_prop * n_total)
  switches_so_far <- 0

  while (!is.na(p_val) && p_val >= 0.05 && switches_so_far < max_switches) {
    # alternate direction to keep group sizes balanced over pairs of switches
    move_notexposed_case <- (switches_so_far %% 2 == 0)
    if (move_notexposed_case) {
      idx <- which(data$group == "not_exposed" & data$outcome == 1); new_group <- "exposed"
    } else {
      idx <- which(data$group == "exposed" & data$outcome == 0); new_group <- "not_exposed"
    }

    if (length(idx) == 0) {
      # preferred donor cell empty; try the other direction
      if (move_notexposed_case) {
        idx <- which(data$group == "exposed" & data$outcome == 0); new_group <- "not_exposed"
      } else {
        idx <- which(data$group == "not_exposed" & data$outcome == 1); new_group <- "exposed"
      }
      if (length(idx) == 0) break
    }

    data$group[idx[1]] <- new_group
    switches_so_far <- switches_so_far + 1

    p_val <- analyse_chisq(data)$p[1]
  }

  return(tibble(tampers_needed = switches_so_far, reached_sig = !is.na(p_val) && p_val < 0.05))
}

```

## Prevalence = .20

### Run simulation

```{r}

simulation_iterative_switching_2x2_path <- cache_path("simulation_iterative_switching_2x2")
if (file.exists(simulation_iterative_switching_2x2_path)) {
  simulation_iterative_switching_2x2 <- read_rds(simulation_iterative_switching_2x2_path)
} else {

  experiment_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    prevalence = 0.20,
    max_switch_prop = 0.5,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_iterative_switching_2x2 <- experiment_grid %>%
    slice_sample(prop = 1) %>% # shuffle rows so that the expensive large-n tasks are spread evenly across workers
    mutate(res = future_pmap(
      .l = list(n_per_condition = n_per_condition,
                prevalence = prevalence,
                max_switch_prop = max_switch_prop),
      .f = run_iterative_switching_2x2,
      .progress = TRUE,
      .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages)
    )) |>
    unnest(res)

  write_rds(simulation_iterative_switching_2x2, simulation_iterative_switching_2x2_path)
}

simulation_switching_2x2_tampers <- summarise_tampers(simulation_iterative_switching_2x2)

```

### Table

```{r}

table_tampers(simulation_switching_2x2_tampers,
              caption = "Condition switching, prevalence = 0.20: tampered data points needed to reach p < .05")

```

### Plot

```{r}

plot_tampers_n(simulation_switching_2x2_tampers, "Condition switching", "N switched")

plot_tampers_prop(simulation_switching_2x2_tampers, "Condition switching", "Proportion switched")

```

## Prevalence = .05

### Run simulation

```{r}

simulation_switching_2x2_lowprev_path <- cache_path("simulation_iterative_switching_2x2_lowprev")
if (file.exists(simulation_switching_2x2_lowprev_path)) {
  simulation_switching_2x2_lowprev <- read_rds(simulation_switching_2x2_lowprev_path)
} else {

  experiment_grid <- expand_grid(
    n_per_condition = c(10, seq(25, 200, by = 25)),
    prevalence = 0.05,
    max_switch_prop = 0.5,
    iteration = 1:n_iterations
  )

  set.seed(123)

  simulation_switching_2x2_lowprev <- experiment_grid %>%
    slice_sample(prop = 1) %>% # shuffle rows so that the expensive large-n tasks are spread evenly across workers
    mutate(res = future_pmap(
      .l = list(n_per_condition = n_per_condition,
                prevalence = prevalence,
                max_switch_prop = max_switch_prop),
      .f = run_iterative_switching_2x2,
      .progress = TRUE,
      .options = furrr_options(seed = TRUE, scheduling = 2, packages = worker_packages)
    )) |>
    unnest(res)

  write_rds(simulation_switching_2x2_lowprev, simulation_switching_2x2_lowprev_path)
}

simulation_switching_2x2_lowprev_tampers <- summarise_tampers(simulation_switching_2x2_lowprev)

```

### Table

```{r}

table_tampers(simulation_switching_2x2_lowprev_tampers,
              caption = "Condition switching, prevalence = 0.05: tampered data points needed to reach p < .05")

```

### Plot

```{r}

plot_tampers_n(simulation_switching_2x2_lowprev_tampers, "Condition switching", "N switched")

plot_tampers_prop(simulation_switching_2x2_lowprev_tampers, "Condition switching", "Proportion switched")

```

# Combined tables and plots

```{r fig.height=8, fig.width=10}

simulation_2x2_combined <- bind_rows(
  simulation_fabrication_2x2_tampers         |> mutate(Method = "Duplication/fabrication/replacement", prevalence = 0.20),
  simulation_fabrication_2x2_lowprev_tampers |> mutate(Method = "Duplication/fabrication/replacement", prevalence = 0.05),
  simulation_dropping_2x2_tampers            |> mutate(Method = "Dropping", prevalence = 0.20),
  simulation_dropping_2x2_lowprev_tampers    |> mutate(Method = "Dropping", prevalence = 0.05),
  simulation_switching_2x2_tampers           |> mutate(Method = "Condition switching", prevalence = 0.20),
  simulation_switching_2x2_lowprev_tampers   |> mutate(Method = "Condition switching", prevalence = 0.05)
) |>
  mutate(Method = fct_relevel(Method, method_levels),
         prevalence_label = paste0("prevalence = ", prevalence))

p_2x2_n <-
  ggplot(simulation_2x2_combined, aes(n_total, median, color = Method, fill = Method)) +
  geom_point(position = position_dodge(width = 27)) +
  geom_linerange(aes(ymin = pi_lower, ymax = pmin(pi_upper, cap)), position = position_dodge(width = 27)) +
  facet_wrap(~ prevalence_label) +
  labs(title = "Count of tampered cells needed to reach a significant 2×2 association",
       subtitle = "Median and 95% prediction interval across iterations, truncated at the 50% tampering budget",
       x = "Sample size",
       y = "Count") +
  theme_linedraw() +
  scale_y_continuous(limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
  scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2) +
  scale_color_manual(values = wes_palette("Cavalcanti1")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1"))

p_2x2_prop <-
  ggplot(simulation_2x2_combined, aes(n_total, prop_median, color = Method, fill = Method)) +
  geom_point(position = position_dodge(width = 27)) +
  geom_linerange(aes(ymin = prop_pi_lower, ymax = pmin(prop_pi_upper, 0.5)), position = position_dodge(width = 27)) +
  facet_wrap(~ prevalence_label) +
  labs(title = "Proportion of the sample tampered to reach a significant 2×2 association",
       subtitle = "Median and 95% prediction interval across iterations, truncated at the 50% tampering budget",
       x = "Sample size",
       y = "Proportion") +
  theme_linedraw() +
  scale_y_continuous(labels = percent_format(), limits = c(0, NA), breaks = breaks_pretty(n = 8)) +
  scale_x_continuous(breaks = c(10, seq(25, 200, by = 25))*2) +
  scale_color_manual(values = wes_palette("Cavalcanti1")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1"))

p_2x2_n +
  p_2x2_prop +
  plot_layout(ncol = 1, guides = "collect")

ggsave(filename = "../communications/plots/iterative_2x2_combined.pdf", device = "pdf", width = 10, height = 8)
ggsave(filename = "../communications/plots/iterative_2x2_combined.png", device = "png", width = 10, height = 8)

```

```{r}

table_tampers(simulation_2x2_combined,
              caption = "All strategies and prevalences: tampered cells needed to reach p < .05",
              extra_cols = c("Method", "prevalence"))

```


# Running on an HPC cluster

This report is written to run unchanged on a laptop and on a SLURM-managed HPC (e.g. the "Everything" cluster: SLURM + interactive RStudio Server), following the practices in `simulation_template_hpc.qmd`. The cluster scaffolding lives in the *Dependencies* and *Common functions and global variables* chunks:

- **Worker auto-detection.** The plan is set with `plan(multisession, workers = parallelly::availableCores())`. `availableCores()` honours the SLURM allocation, cgroup CPU quotas, and affinity masks, so the parallel plan adapts to the cores actually granted to the session rather than to all physical cores; pass an integer instead to pin a count. On Everything, 4–16 workers is the useful range (little gain beyond that for this workload).
- **BLAS thread pinning.** BLAS/LAPACK is pinned to one thread *before* the workers spawn, so dozens of worker processes do not oversubscribe the node. The matrices here are tiny, so single-threaded BLAS costs nothing.
- **Replication knob.** `run_full` / `on_hpc` auto-detect the environment (any `SLURM_*` variable present ⇒ cluster) and set `n_iterations` accordingly — a fast local smoke test (100) versus the full cluster run (10000). Force it with `run_full <- TRUE`/`FALSE`.
- **Tagged caches.** `cache_path()` tags every `.rds` with `dev`/`full` and the iteration count (e.g. `..._full_k10000.rds`), so local and cluster runs never overwrite each other. Re-rendering reads the cached summary instead of recomputing.
- **Reproducible parallel seeding.** `furrr_options(seed = TRUE)` advances an L'Ecuyer-CMRG stream per task, so results are identical regardless of the worker count; `scheduling = 2` enables dynamic load balancing, and `packages = worker_packages` attaches the needed libraries on each worker.

**Granularity (a deliberate deviation from the template).** The template parallelises over *conditions* and summarises inside each worker, so the master never holds raw replicate rows. These simulations instead parallelise over *replicates* (one task per condition × iteration) and keep the raw per-iteration results, because the distribution histograms and bootstrap-based median MCSEs need the full per-iteration vectors — summarising in the worker would discard exactly the data the report relies on. With many small tasks and `scheduling = 2`, the cost of the expensive large-*n* conditions is spread across workers automatically; each parameter grid is additionally shuffled with `slice_sample(prop = 1)` (after `set.seed()`, so the shuffle is itself reproducible) so that the expensive rows are not contiguous in the task list — the same load-balancing aid the template applies to condition-level tasks.

To run on Everything: request a node and launch RStudio Server inside the allocation, e.g. `salloc --nodes=1 --cpus-per-task=16 --mem=64G --time=08:00:00`; optionally export `FLEXIBLAS_NUM_THREADS=1 OMP_NUM_THREADS=1` in the shell before launching R (belt-and-braces; confirm with `RhpcBLASctl::blas_get_num_procs()`); then render. The full run regenerates the `*_full_k10000.rds` caches.

# Discussion

## Performance of the strategies across conditions

The substantive results are in the per-strategy / per-prevalence tables and the *Combined tables and plots* section; this discussion summarises how they should be read against the aims. Three patterns are checkable against the rendered figures:

- **Effort is small in proportional terms.** Across all three strategies the proportion of cells that must be tampered to reach significance is small and does not grow with N.
- **Rare outcomes are especially fragile.** At prevalence 0.05 a handful of tampered cells moves the table disproportionately, so significance is reached with fewer operations than at prevalence 0.20.
- **The strategies behave similarly, with switching preserving N.** Because for binary data the operations are near-equivalent, the duplication/fabrication/replacement family, dropping, and condition switching require comparable numbers of tampered cells; condition switching is the only one that leaves the total N unchanged, which makes it the least conspicuous on the sample size alone.

## Conclusions with regard to the aims

The aims — to estimate the number of tampered cells needed and how it varies with strategy, sample size, and prevalence — are met by the per-strategy sections and the combined comparison. The headline conclusion mirrors the other settings: the distance from a null 2×2 table to a "significant" one is short, and is shortest when the outcome is rare.

## Limitations and intended use

- **Proof of principle, not a fraud model.** The strategies are deliberately simple and greedy; a detection-aware fraudster would trade efficiency for plausibility.
- **Prevalence not preserved.** None of the strategies preserves the marginal outcome prevalence, which would be a giveaway; prevalence-preserving variants are a natural extension.
- **No effect-size bias.** Only the tamper count is reported in this binary setting; a φ / Cramér's V bias measure (the analogue of the Hedges' g / Pearson r work in the companion files) is left for future work.
- **Single threshold and budget.** Results are conditioned on α = .05 and a 50% tampering cap, both reported transparently.

# References

Chambers, C. (2017). *The seven deadly sins of psychology: A manifesto for reforming the culture of scientific practice.* Princeton University Press.

Morris, T. P., White, I. R., & Crowther, M. J. (2019). Using simulation studies to evaluate statistical methods. *Statistics in Medicine, 38*(11), 2074–2102. https://doi.org/10.1002/sim.8086

Siepe, B. S., Bartoš, F., Morris, T. P., Boulesteix, A.-L., Heck, D. W., & Pawel, S. (2024). Simulation studies for methodological research in psychology: A standardized template for planning, preregistration, and reporting. *Psychological Methods.* https://doi.org/10.1037/met0000695

*[citation to be completed]* Spiess et al. (2024) / the `dfstat`–`dfbeta` influential-observation literature.

# Session info

```{r}

sessionInfo()

```
